---
title: "TFM_CODE"
author: "Noah Emanuel Fürup"
format: html
editor: visual
---

##Libraries

```{r LIBRARIES}
library(tidyverse)
library(readxl)
library(sf)
library(rnaturalearth)
library(httr2)
library(countrycode)
library(mapSpain)
library(stargazer)
library(rvest)

# tables are printed as text here; use type = "html" (or "latex") for the written document
table_type <- "text"
```

##Charts of the Theoretical Framework

ARREGLAR FRANCIA, NORUEGA que faltan


```{r ELECTORAL SYSTEM MAP}
electoral_system_database <-
  read_excel("idea_export_electoral_system_design_database.xlsx") |> 
  group_by(ISO3) |> 
   slice_max(Year) |> 
  ungroup() |> 
  select(ISO3, Country, Year,
         family = `Electoral system family`,
         system = `Electoral system for national legislature`)
 
europe <- 
  ne_countries(scale = "medium", continent = "Europe", returnclass = "sf") |>
  left_join(electoral_system_database, by = c("iso_a3" = "ISO3"))

ggplot(europe) +
  geom_sf(aes(fill = family), colour = "white", linewidth = 0.2) +
  coord_sf(xlim = c(-25, 45), ylim = c(34, 72)) +
  scale_fill_brewer(palette = "Set2", na.value = "grey90", name = "Electoral system family") +
  labs(title = "Electoral systems in Europe",
       caption = "Source: IDEA Electoral System Design Database") +
  theme_void()

```

```{r share of each electoral family}
europe |>
  st_drop_geometry() |>
  count(family) |>
  mutate(pct = round(100 * n / sum(n), 1)) |>
  arrange(desc(n))
```
Maldita from Wayback Machine 
https://maldita.es/malditodato/20240229/partidos-politicos-registro-ministerio-interior/
https://files.maldita.es/uploads/2024/02/65df376828af1informefirma-1706550494955-pdf.pdf#toolbar=0

```{r PARTY REGISTRY}
party_registry_maldita <- read_csv("PARTY_REGISTRY_MALDITA_2024.csv", show_col_types = FALSE)

party_registry_2026 <- read_excel("PARTY_REGISTRY_2026_09.xlsx", col_types = "text") |>
  mutate(new_party = !is.na(registration_date),
         party_id  = cumsum(new_party)) |>
  group_by(party_id) |>
  summarise(party             = str_squish(str_c(party, collapse = " ")),
            address           = first(adress),
            province          = first(province),
            registration_date = first(registration_date),
            .groups = "drop") |>
  mutate(
    party_name        = str_squish(str_remove(party, "\\(([^()]*)\\)$")),
    abbreviation      = str_match(party, "\\(([^()]*)\\)$")[, 2],
    town              = str_squish(str_remove(province, "\\(([^()]*)\\)$")),
    province          = str_match(province, "\\(([^()]*)\\)$")[, 2],
    registration_date = dmy(registration_date),
    status            = "registered"
  ) |>
  select(party_name, abbreviation, registration_date, town, province, address, status)

# parties registered after the 2024 list; some of them were already in it as "not registered"
new_registrations <- party_registry_2026 |>
  filter(registration_date > max(party_registry_maldita$registration_date, na.rm = TRUE))

all_parties <- party_registry_maldita |>
  left_join(new_registrations |>
              select(party_name, new_date = registration_date, province, address),
            by = "party_name") |>
  mutate(registration_date = coalesce(registration_date, new_date),
         status = if_else(!is.na(new_date) & status == "not registered", "registered", status)) |>
  select(-new_date) |>
  bind_rows(new_registrations |> anti_join(party_registry_maldita, by = "party_name")) |>
  arrange(registration_date)
```

FE de las JONS aparece dos veces sin fecha de inscripción: protocolo 21 "(Novillo)" y 22 "(Zulueta)", que son las dos facciones que pidieron la inscripción en 1976. En cambio, party_register la da inscrita el 04/10/1976. Para la fecha de FE de las JONS usa esa.
Alianza Popular (protocolo 961) sale sin fecha de inscripción. Es por cómo lo registra el Ministerio: el PP heredó la inscripción de AP.


```{r TABLE OF FAR RIGHT PARTIES BEFORE VOX}
casals_parties <- c(
  "FUERZA NUEVA", "CENTRO DEMOCRATICO Y SOCIAL", "JUNTAS ESPAÑOLAS", "FRENTE NACIONAL",
  "DEMOCRACIA NACIONAL", "MOVIMIENTO SOCIAL REPUBLICANO",
  "PARTIDO DEL TRABAJO Y EMPLEO-AGRUPACION RUIZ MATEOS \"TRABAJO Y EMPLEO\"",
  "GRUPO INDEPENDIENTE LIBERAL", "SOCIEDAD CIVIL Y DEMOCRACIA", "PLATAFORMA PER CATALUNYA", "ESPAÑA 2000",
  "PARTIDO POR LA LIBERTAD", "ADÑ", "FALANGE ESPAÑOLA DE LAS J.O.N.S.", "LA FALANGE", "FALANGE ESPAÑOLA AUTENTICA",
  "FALANGE AUTENTICA", "FALANGE 2.000", "\"NOSOTROS\" PARTIDO DE LA REGENERACIÓN SOCIAL",
  "ALTERNATIVA ESPAÑOLA", "ALTERNATIVA DE GOVERN", "ALIANZA NACIONAL",
  "SOBERANÍA Y LIBERTAD", "PUEBLO", "FRENTE: NACIONAL IDENTITARIO, PARTIDO NACIONALSOCIALISTA OBRERO ESPAÑOL",
  "SOM IDENTITARIS",
  # Via Democratica, split from PxC in 2010 (Casals); registered under its later name
  "JUNTS PER TABARNIA - VIA DEMOCRÁTICA"
)

casals_registry <- bind_rows(all_parties, party_registry_2026) |>
  filter(party_name %in% casals_parties, !is.na(registration_date)) |>
  group_by(party_name) |>
  slice_min(registration_date, n = 1, with_ties = FALSE) |>
  ungroup() |>
  select(party_name, abbreviation, registration_date, cancellation_date) |>
  arrange(registration_date) |> 
  mutate(party = party_name)

far_right_table <- casals_registry |>
  mutate(across(c(registration_date, cancellation_date), \(x) replace_na(format(x, "%d/%m/%Y"), "")),
         abbreviation = replace_na(abbreviation, "")) |>
  select(Party = party_name, Abbreviation = abbreviation,
         Registered = registration_date, Cancelled = cancellation_date)

stargazer(as.data.frame(far_right_table), summary = FALSE, rownames = FALSE, type = table_type,
          title = "Far-right and right-wing populist parties since 1976",
          notes = c("Source: Ministry of the Interior registry of political parties (Maldita 2024; update of September 2026).",
                    "Parties selected following Casals (2020). Cancelled: date the party was struck off the registry."))
```


```{r FAMILY TREE FAR PARTIES 1977-2026}
library(tidygraph)
library(ggraph)

# ---- 1. connections: `from` gave rise to `to` ------------------------------------------------------
# type: "split / successor" = the new organisation comes out of the old one
#       "coalition / merger" = the old one is part of a coalition, platform or merger
#       "members / founders" = founders or members of the old one created or joined the new one
far_right_links <- tribble(
  ~from,                             ~to,                                ~type,                 ~year, ~source,
  # Francoist dictatorship, mainstream right and Vox (Casals 2020)
  "Francoist dictatorship",          "Alianza Popular",                  "members / founders",  1977L, "Casals (2020: 368)",
  "Alianza Popular",                 "Partido Popular",                  "split / successor",   1989L, "Casals (2020: 368)",
  "Partido Popular",                 "Vox",                              "split / successor",   2013L, "Casals (2020: 374, 379)",
  "Sociedad Civil y Democracia",     "Vox",                              "members / founders",  2014L, "Casals (2020: 371)",
  "Francoist dictatorship",          "Fuerza Nueva",                     "members / founders",  1976L, "Casals (2020: 366-367)",
  "Fuerza Nueva",                    "Frente Nacional (1986)",           "split / successor",   1986L, "Casals (2020: 370)",
  "Juntas Españolas",                "Alternativa Demócrata Nacional",   "coalition / merger",  1994L, "Casals (2020: 370)",
  "Fuerza Nueva",                    "Alternativa Española",             "members / founders",  2003L, "Casals (2020: 379)",
  "Partido Popular",                 "NOSOTROS",                         "members / founders",  2017L, "Ramos (2022: 165)",
  # Falangism (Wikipedia: FE de las JONS (1976), FE Independiente, FE Auténtica, La Falange, Falange Auténtica)
  "Francoist dictatorship",          "FE de las JONS",                   "members / founders",  1976L, "Wikipedia",
  "Frente de Estudiantes Sindicalistas", "FE Independiente",             "split / successor",   1977L, "Wikipedia",
  "FE de las JONS (Auténtica)",      "FE Auténtica",                     "split / successor",   1979L, "Wikipedia",
  "FE Auténtica",                    "FE de las JONS",                   "members / founders",  1994L, "Wikipedia",
  "FE de las JONS",                  "La Falange",                       "split / successor",   1999L, "Wikipedia",
  "La Falange",                      "Falange Auténtica",                "split / successor",   2002L, "Wikipedia",
  "FE Independiente",                "FE de las JONS",                   "coalition / merger",  2004L, "Wikipedia",
  "Falange 2000",                    "FE de las JONS",                   "coalition / merger",  2004L, "Wikipedia",
  "La Falange",                      "Frente Nacional (2006)",           "split / successor",   2005L, "Wikipedia",
  "La Falange",                      "FE de las JONS",                   "coalition / merger",  2024L, "Wikipedia",
  "FE de las JONS",                  "Comunidad Política Vértice",       "members / founders",  2018L, "Ramos (2022: 210)",
  "Fuerza Nueva",                    "Frente de la Juventud",            "split / successor",   NA,    "Ramos (2022: 199)",
  "Frente de la Juventud",           "ACIMJI",                           "members / founders",  2011L, "Ramos (2022: 199)",
  # neo-Nazi and national-revolutionary (Ramos 2022)
  "CEDADE",                          "Democracia Nacional",              "members / founders",  1995L, "Ramos (2022: 126)",
  "Acción Radical",                  "Democracia Nacional",              "members / founders",  1995L, "Ramos (2022: 126)",
  "Juntas Españolas",                "Democracia Nacional",              "members / founders",  1995L, "Ramos (2022: 126)",
  "Alternativa Demócrata Nacional",  "Democracia Nacional",              "split / successor",   1995L, "Ramos (2022: 126)",
  "Democracia Nacional",             "Plataforma España 2000",           "coalition / merger",  2000L, "Ramos (2022: 127)",
  "Plataforma España 2000",          "España 2000",                      "split / successor",   2002L, "Ramos (2022: 130)",
  "FE-Frente Nacional Sindicalista", "España 2000",                      "members / founders",  2002L, "Ramos (2022: 130)",
  "Alianza por la Unidad Nacional",  "Alianza Nacional",                 "members / founders",  2005L, "Ramos (2022: 148)",
  "AC Rey Sisebuto",                 "Alianza Nacional",                 "members / founders",  2005L, "Ramos (2022: 148)",
  "CEDADE",                          "Hermandad Aria",                   "split / successor",   1995L, "Ramos (2022: 106)",
  "CEDADE",                          "Devenir Europeo",                  "members / founders",  2007L, "Ramos (2022: 212)",
  "Alternativa Europea",             "Movimiento Social Republicano",    "split / successor",   1999L, "Ramos (2022: 237)",
  "Movimiento Social Republicano",   "Soberanía y Libertad",             "split / successor",   2014L, "Ramos (2022: 240)",
  "Soberanía y Libertad",            "Pueblo",                           "split / successor",   2018L, "Ramos (2022: 241)",
  "Movimiento Social Republicano",   "Hogar Social Madrid",              "members / founders",  2014L, "Ramos (2022: 168)",
  "REFNI",                           "FNI-PNSOE",                        "split / successor",   2020L, "Ramos (2022: 242)",
  # coalitions and platforms (Ramos 2022)
  "Democracia Nacional",             "La España en Marcha",              "coalition / merger",  2013L, "Ramos (2022: 235)",
  "Alianza Nacional",                "La España en Marcha",              "coalition / merger",  2013L, "Ramos (2022: 235)",
  "Nudo Patriota Español",           "La España en Marcha",              "coalition / merger",  2013L, "Ramos (2022: 235)",
  "La Falange",                      "La España en Marcha",              "coalition / merger",  2013L, "Ramos (2022: 235)",
  "Movimiento Católico Español",     "La España en Marcha",              "coalition / merger",  2013L, "Ramos (2022: 235)",
  "FE de las JONS",                  "ADÑ",                              "coalition / merger",  2018L, "Ramos (2022: 157)",
  "La Falange",                      "ADÑ",                              "coalition / merger",  2018L, "Ramos (2022: 157)",
  "Democracia Nacional",             "ADÑ",                              "coalition / merger",  2018L, "Ramos (2022: 157)",
  "Alternativa Española",            "ADÑ",                              "coalition / merger",  2018L, "Ramos (2022: 157)",
  "España 2000",                     "Respeto",                          "coalition / merger",  2016L, "Ramos (2022: 145, 243)",
  "Plataforma per Catalunya",        "Respeto",                          "coalition / merger",  2016L, "Ramos (2022: 145, 243)",
  "Partido por la Libertad",         "Respeto",                          "coalition / merger",  2016L, "Ramos (2022: 145, 243)",
  "Iniciativa por Albacete",         "Respeto",                          "coalition / merger",  2016L, "Ramos (2022: 145)",
  # populist radical right in Catalonia (Ramos 2022)
  "Fuerza Nueva",                    "Plataforma per Catalunya",         "members / founders",  2002L, "Ramos (2022: 238)",
  "Plataforma Vigatana",             "Plataforma per Catalunya",         "split / successor",   2003L, "Ramos (2022: 238)",
  "Plataforma per Catalunya",        "Vox",                              "members / founders",  2019L, "Ramos (2022: 243); Casals (2020: 373)",
  "Plataforma per Catalunya",        "Som Catalans",                     "split / successor",   2014L, "Ramos (2022: 257)",
  "Plataforma per Catalunya",        "Frente Nacional Identitario (Anglada)", "split / successor", 2015L, "Ramos (2022: 244)",
  "Frente Nacional Identitario (Anglada)", "Som Identitaris",            "split / successor",   2016L, "Ramos (2022: 244)",
  "Plataforma per Catalunya",        "Generación Identitaria",           "members / founders",  2016L, "Ramos (2022: 233)",
  "Plataforma per Catalunya",        "Casal Tramuntana",                 "members / founders",  2012L, "Ramos (2022: 226)",
  "Movimiento Social Republicano",   "Casal Tramuntana",                 "members / founders",  2012L, "Ramos (2022: 226)",
  "Movimiento Patriótico Catalán",   "Casal Tramuntana",                 "members / founders",  2012L, "Ramos (2022: 226)",
  "Democracia Nacional",             "Por España Me Atrevo",             "coalition / merger",  2016L, "Ramos (2022: 239)",
  "Somatemps",                       "Por España Me Atrevo",             "coalition / merger",  2016L, "Ramos (2022: 239)",
  "Generación Identitaria",          "Por España Me Atrevo",             "coalition / merger",  2016L, "Ramos (2022: 239)",
  "Legión Urbana",                   "Por España Me Atrevo",             "coalition / merger",  2016L, "Ramos (2022: 239)",
  "Hermandad de Antiguos Caballeros Legionarios", "Legión Urbana",       "members / founders",  NA,    "Ramos (2022: 236)",
  "Somatemps",                       "Tradició Catalana",                "members / founders",  2015L, "Ramos (2022: 247)",
  "CEDADE",                          "Unitat Nacional Catalana",         "members / founders",  2000L, "Ramos (2022: 256)",
  "Unitat Nacional Catalana",        "Moviment Identitari Català",       "members / founders",  2015L, "Ramos (2022: 255)",
  # identitarian and neo-fascist groups (Ramos 2022)
  "Iberia Cruor",                    "Hacer Nación",                     "coalition / merger",  2020L, "Ramos (2022: 180)",
  "Acción Social Cádiz",             "Hacer Nación",                     "coalition / merger",  2020L, "Ramos (2022: 180)",
  "El Galeón",                       "Hacer Nación",                     "coalition / merger",  2020L, "Ramos (2022: 180)",
  "Málaga 1487",                     "Hacer Nación",                     "coalition / merger",  2020L, "Ramos (2022: 180)",
  "España 2000",                     "Hacer Nación",                     "members / founders",  2020L, "Ramos (2022: 180)",
  "Vox",                             "Hacer Nación",                     "members / founders",  2020L, "Ramos (2022: 180)",
  "Lo Nuestro",                      "El Galeón",                        "split / successor",   2018L, "Ramos (2022: 188)",
  # people who later joined Vox (Ramos 2022)
  "España 2000",                     "Vox",                              "members / founders",  2019L, "Ramos (2022: 106, 159)",
  "Hermandad Aria",                  "Vox",                              "members / founders",  2019L, "Ramos (2022: 106)",
  "Acción Radical",                  "Vox",                              "members / founders",  2019L, "Ramos (2022: 106, 159)",
  "Partido por la Libertad",         "Vox",                              "members / founders",  2019L, "Ramos (2022: 106)",
  "Vox",                             "Regeneración",                     "split / successor",   2019L, "Ramos (2022: 159)",
  "Regeneración",                    "Identitarios",                     "split / successor",   2019L, "Ramos (2022: 159)"
)

# ---- 2. nodes: family and founding year --------------------------------------------------------------------
# documented = FALSE: the sources give no founding year, so `founded` is an approximate year of activity
far_right_nodes <- tribble(
  ~name,                                          ~family,                                ~founded, ~documented,
  "Francoist dictatorship",                       "Francoist dictatorship",               1939L,     TRUE,
  "Alianza Popular",                              "Mainstream right",                     1976L,     TRUE,
  "Partido Popular",                              "Mainstream right",                     1989L,     TRUE,
  "Sociedad Civil y Democracia",                  "Mainstream right",                     2011L,     TRUE,
  "Vox",                                          "Vox",                                  2013L,     TRUE,
  "Regeneración",                                 "Vox",                                  2019L,     TRUE,
  "Identitarios",                                 "Vox",                                  2019L,     TRUE,
  "NOSOTROS",                                     "Vox",                                  2017L,     TRUE,
  "Fuerza Nueva",                                 "Neo-Francoism / national-Catholicism", 1976L,     TRUE,
  "Frente Nacional (1986)",                       "Neo-Francoism / national-Catholicism", 1986L,     TRUE,
  "Juntas Españolas",                             "Neo-Francoism / national-Catholicism", 1984L,     TRUE,
  "Alternativa Española",                         "Neo-Francoism / national-Catholicism", 2003L,     TRUE,
  "Alianza por la Unidad Nacional",               "Neo-Francoism / national-Catholicism", 1995L,     TRUE,
  "Movimiento Católico Español",                  "Neo-Francoism / national-Catholicism", 1982L,     TRUE,
  "Frente de la Juventud",                        "Neo-Francoism / national-Catholicism", 1980L,     FALSE,   # active in 1980-1981 (Ramos 2022: 199)
  "ACIMJI",                                       "Neo-Francoism / national-Catholicism", 2011L,     TRUE,
  "Nudo Patriota Español",                        "Neo-Francoism / national-Catholicism", 2011L,     FALSE,   # before La España en Marcha (2013)
  "Somatemps",                                    "Neo-Francoism / national-Catholicism", 2013L,     TRUE,
  "Tradició Catalana",                            "Neo-Francoism / national-Catholicism", 2015L,     TRUE,
  "Legión Urbana",                                "Neo-Francoism / national-Catholicism", 2014L,     FALSE,   # before Por España Me Atrevo (2016)
  "Hermandad de Antiguos Caballeros Legionarios", "Neo-Francoism / national-Catholicism", 1985L,     FALSE,   # active "for decades" (Ramos 2022: 239)
  "Movimiento Patriótico Catalán",                "Neo-Francoism / national-Catholicism", 1994L,     FALSE,   # Wikidata
  "FE de las JONS",                               "Falangism",                            1976L,     TRUE,
  "FE Independiente",                             "Falangism",                            1977L,     TRUE,
  "FE de las JONS (Auténtica)",                   "Falangism",                            1976L,     TRUE,
  "FE Auténtica",                                 "Falangism",                            1979L,     TRUE,
  "La Falange",                                   "Falangism",                            1999L,     TRUE,
  "Falange Auténtica",                            "Falangism",                            2002L,     TRUE,
  "Falange 2000",                                 "Falangism",                            2000L,     TRUE,
  "Frente Nacional (2006)",                       "Falangism",                            2006L,     TRUE,
  "Frente de Estudiantes Sindicalistas",          "Falangism",                            1963L,     TRUE,
  "FE-Frente Nacional Sindicalista",              "Falangism",                            1996L,     TRUE,
  "Comunidad Política Vértice",                   "Falangism",                            2018L,     TRUE,
  "CEDADE",                                       "Neo-Nazi / national-revolutionary",    1966L,     TRUE,
  "Acción Radical",                               "Neo-Nazi / national-revolutionary",    1990L,     FALSE,   # dismantled in 1993 (Ramos 2022: 229)
  "Democracia Nacional",                          "Neo-Nazi / national-revolutionary",    1995L,     TRUE,
  "Alternativa Demócrata Nacional",               "Neo-Nazi / national-revolutionary",    1994L,     TRUE,
  "Alianza Nacional",                             "Neo-Nazi / national-revolutionary",    2005L,     TRUE,
  "AC Rey Sisebuto",                              "Neo-Nazi / national-revolutionary",    2003L,     FALSE,   # before Alianza Nacional (2005)
  "Hermandad Aria",                               "Neo-Nazi / national-revolutionary",    1995L,     TRUE,
  "Devenir Europeo",                              "Neo-Nazi / national-revolutionary",    2007L,     TRUE,
  "Alternativa Europea",                          "Neo-Nazi / national-revolutionary",    1994L,     FALSE,   # active in 1994 (Ramos 2022: 451)
  "Movimiento Social Republicano",                "Neo-Nazi / national-revolutionary",    1999L,     TRUE,
  "Soberanía y Libertad",                         "Neo-Nazi / national-revolutionary",    2014L,     TRUE,
  "Pueblo",                                       "Neo-Nazi / national-revolutionary",    2018L,     TRUE,
  "REFNI",                                        "Neo-Nazi / national-revolutionary",    2017L,     TRUE,
  "FNI-PNSOE",                                    "Neo-Nazi / national-revolutionary",    2020L,     TRUE,
  "Plataforma España 2000",                       "Populist radical right",               2000L,     TRUE,
  "España 2000",                                  "Populist radical right",               2002L,     TRUE,
  "Plataforma per Catalunya",                     "Populist radical right",               2002L,     TRUE,
  "Plataforma Vigatana",                          "Populist radical right",               2002L,     TRUE,
  "Partido por la Libertad",                      "Populist radical right",               2013L,     TRUE,
  "Iniciativa por Albacete",                      "Populist radical right",               2015L,     FALSE,   # before Respeto (2016)
  "Frente Nacional Identitario (Anglada)",        "Populist radical right",               2015L,     TRUE,
  "Som Identitaris",                              "Populist radical right",               2016L,     TRUE,
  "Casal Tramuntana",                             "Populist radical right",               2012L,     TRUE,
  "Hacer Nación",                                 "Identitarian / neo-fascist groups",    2020L,     TRUE,
  "Iberia Cruor",                                 "Identitarian / neo-fascist groups",    2008L,     TRUE,
  "Acción Social Cádiz",                          "Identitarian / neo-fascist groups",    2016L,     TRUE,
  "El Galeón",                                    "Identitarian / neo-fascist groups",    2018L,     TRUE,
  "Málaga 1487",                                  "Identitarian / neo-fascist groups",    2013L,     TRUE,
  "Lo Nuestro",                                   "Identitarian / neo-fascist groups",    2016L,     FALSE,   # before El Galeón (2018)
  "Hogar Social Madrid",                          "Identitarian / neo-fascist groups",    2014L,     TRUE,
  "Generación Identitaria",                       "Identitarian / neo-fascist groups",    2016L,     TRUE,
  "Por España Me Atrevo",                         "Identitarian / neo-fascist groups",    2016L,     TRUE,
  "Som Catalans",                                 "Catalan independentist far right",     2014L,     TRUE,
  "Unitat Nacional Catalana",                     "Catalan independentist far right",     2000L,     TRUE,
  "Moviment Identitari Català",                   "Catalan independentist far right",     2015L,     TRUE,
  "La España en Marcha",                          "Coalitions",                           2013L,     TRUE,
  "ADÑ",                                          "Coalitions",                           2018L,     TRUE,
  "Respeto",                                      "Coalitions",                           2016L,     TRUE
)

# every organisation in the links needs a row in the nodes table (must return 0 rows)
tibble(name = unique(c(far_right_links$from, far_right_links$to))) |>
  anti_join(far_right_nodes, by = "name")

# ---- 3. position of each node ----------------------------------------------------------------------------
# x = founding year; y = one horizontal band per family
family_order <- c("Francoist dictatorship", "Mainstream right", "Vox", "Coalitions", "Populist radical right",
                  "Identitarian / neo-fascist groups", "Catalan independentist far right",
                  "Neo-Nazi / national-revolutionary", "Neo-Francoism / national-Catholicism", "Falangism")

far_right_nodes <- far_right_nodes |>
  mutate(x    = pmax(founded, 1965L),               # the dictatorship (1939) is drawn at the left edge
         band = match(family, family_order)) |>
  group_by(family) |>
  arrange(x, .by_group = TRUE) |>
  mutate(y = band * 9 + (row_number() %% 5) * 1.6) |>   # stagger names inside each band
  ungroup() |>
  select(-band)

far_right_graph <- tbl_graph(nodes = far_right_nodes, edges = far_right_links, directed = TRUE)

# ---- 4. plot -----------------------------------------------------------------------------------------------
family_colours <- c(
  "Francoist dictatorship"               = "#3B3B3B",
  "Mainstream right"                     = "#1F77B4",
  "Vox"                                  = "#2CA02C",
  "Coalitions"                           = "#7F7F7F",
  "Populist radical right"               = "#FF7F0E",
  "Identitarian / neo-fascist groups"    = "#9467BD",
  "Catalan independentist far right"     = "#E6AB02",
  "Neo-Nazi / national-revolutionary"    = "#000000",
  "Neo-Francoism / national-Catholicism" = "#8C564B",
  "Falangism"                            = "#D62728"
)

ggraph(far_right_graph, layout = "manual", x = x, y = y) +
  geom_edge_arc(aes(colour = type), strength = 0.1,
                arrow = arrow(length = unit(1.5, "mm"), type = "closed"),
                end_cap = circle(2.2, "mm"), edge_width = 0.35, alpha = 0.55) +
  geom_node_point(aes(colour = family, fill = if_else(documented, family, "approximate")),
                  shape = 21, size = 3, stroke = 0.9) +
  geom_node_text(aes(label = name, colour = family), size = 2.4, repel = TRUE, max.overlaps = Inf,
                 segment.colour = "grey85", segment.size = 0.2, show.legend = FALSE) +
  scale_colour_manual(values = family_colours, breaks = family_order, name = NULL) +
  scale_fill_manual(values = c(family_colours, approximate = "white"), guide = "none") +
  scale_edge_colour_manual(values = c("split / successor"  = "#D62728",
                                      "coalition / merger" = "#1F77B4",
                                      "members / founders" = "grey60"),
                           name = NULL) +
  scale_x_continuous(breaks = c(1965, seq(1975, 2025, 5)), labels = c("<1975", seq(1975, 2025, 5)),
                     expand = expansion(add = 2)) +
  scale_y_reverse() +
  labs(title = "Family tree of the Spanish far right, 1975-2026",
       subtitle = "Each arrow goes from an organisation to the one that came out of it, joined a coalition with it or supplied its members. Hollow dots: approximate year",
       caption = "Sources: Ramos (2022), Casals (2020), Wikipedia (Falangist parties)",
       x = NULL, y = NULL) +
  theme_minimal(base_size = 11) +
  theme(panel.grid.major.y = element_blank(), panel.grid.minor = element_blank(),
        panel.grid.major.x = element_line(colour = "grey93"),
        axis.text.y = element_blank(),
        legend.position = "bottom", legend.box = "vertical",
        plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(size = 9, colour = "grey35"))
```
HACERLO FACETED, FALANGE, 



```{r}
library(tidyverse)
library(grid)


```

https://web.archive.org/web/20140209110219/http://www.reconversion.es/

```{r TABLE SIGNEES RECONVERSION.ES}

signees_reconversion <- read_html("https://web.archive.org/web/20140202204018/http://www.reconversion.es/firmantes/") |>
  html_text2() |>
  str_split("\n") |>
  unlist() |>
  str_squish() |> 
  keep(~ str_detect(.x, ",")) |>
  keep(~ str_count(.x, ",") >= 1) |>
  keep(~ str_length(.x) > 3) |> 
  str_split_fixed(",", 2) |> 
  as_tibble(.name_repair = "minimal") |> 
  set_names(c("surname", "given_name")) |>
  mutate(
    surname = str_squish(surname),
    given_name = str_squish(given_name),
    name = paste(given_name, surname)
  ) |>
  select(name) |> 
  slice(-c(1, n()))

signees_reconversion
```


```{r TABLE OF VOX'S FOUNDERS}
founders <- tribble(
  ~name,                              ~relationship_with_vox,                                      ~origin,

  "Alejo Vidal-Quadras",
  "First party President and Candidate for the 2014 European elections",
  "President of the Catalan branch of the PP, member of its national executive; Director of FAES, DENAES",

  "Ana Velasco Vidal-Abarca",
  "Militant",
  "Villasneros Foundation",

  "Cristina Seguí",
  "Ex-militant",
  "-",

  "Enrique Álvarez López",
  "Militant",
  "-",

  "Ignacio Camuñas Solís",
  "Ex-Vice-President and ex-President",
  "Co-founder and ex Minister of UCD",

  "Iván Espinosa de los Monteros",
  "Spokesman, former Vice-President, ex-General Secretariat",
  "Ex-Secretary of DENAES",

  "José Antonio Ortega Lara",
  "Militant",
  "Ex-militant of the PP",

  "José Luis González Quirós",
  "Second President (ad interim), ex-militant",
  "UCD, CDS, ex-militant of PP, FAES",

  "María Jesús Prieto-Laffargue",
  "Ex-militant",
  "General Direction in the first Aznar Government (PP)",

  "Santiago Abascal",
  "President",
  "PP, President of DENAES"
)

founders_table <- founders |>
  select(Name = name, `Relationship with Vox` = relationship_with_vox, Origin = origin)

stargazer(as.data.frame(founders_table), summary = FALSE, rownames = FALSE, type = table_type,
          title = "Founders of Vox",
          notes = "Founding members of Vox and their relationship with the party.")
```


# Data Extraction

## General elections

Extraction of general elections candidates: european and general elections via github.com/JaimeObregon/infoelectoral



```{r GITHUB AND INTERIOR MINISTERY PORTAL}
GITHUB <- "https://raw.githubusercontent.com/JaimeObregon/infoelectoral/main/files"
PORTAL <- "https://infoelectoral.interior.gob.es/estaticos/docxl/apliextr"
```

The github has the files in .DAT, a file type that doesn't separate in ";" or "\|". In order to separate each column, we need to map where the separation between each variable occurs. I asked Claude to give me the exact character positions where the separation is. I used fwf_positions of ReadR to write down the column names, given the positions where the variables lie in the .DAT file.

EXPLICAR COMO FUNCIONAN LO DE 02197706_TOTA y decir que el github copia la estructura del ministerio

```{r FIXED POSITIONS}
# positions of each variable in the .DAT files
cols_parties <- fwf_cols(
  party_code   = c(9, 14),
  abbreviation = c(15, 64),
  party_name   = c(65, 214)
)

cols_candidates <- fwf_cols(
  province_code  = c(10, 11),
  party_code     = c(16, 21),
  list_position  = c(22, 24),
  candidate_type = c(25, 25),
  candidate      = c(26, 100),
  sex            = c(101, 101),
  elected        = c(120, 120)
)
```

```{r GENERAL AND EUROPEAN ELECTIONS TIBBLE (2014-2024)}
elections <- tribble(
  ~process, ~year,  ~month, ~chamber,              ~mirror_folder,
  "02",     "1977", "06",   "Congress",            "congreso",
  "02",     "1979", "03",   "Congress",            "congreso",
  "02",     "1982", "10",   "Congress",            "congreso",
  "02",     "1986", "06",   "Congress",            "congreso",
  "02",     "1989", "10",   "Congress",            "congreso",
  "02",     "1993", "06",   "Congress",            "congreso",
  "02",     "1996", "03",   "Congress",            "congreso",
  "02",     "2000", "03",   "Congress",            "congreso",
  "02",     "2004", "03",   "Congress",            "congreso",
  "02",     "2008", "03",   "Congress",            "congreso",
  "02",     "2011", "11",   "Congress",            "congreso",
  "02",     "2015", "12",   "Congress",            "congreso",
  "02",     "2016", "06",   "Congress",            "congreso",
  "02",     "2019", "04",   "Congress",            "congreso",
  "02",     "2019", "11",   "Congress",            "congreso",
  "02",     "2023", "07",   "Congress",            "congreso",
  "03",     "1977", "06",   "Senate",              "senado",
  "03",     "1979", "03",   "Senate",              "senado",
  "03",     "1982", "10",   "Senate",              "senado",
  "03",     "1986", "06",   "Senate",              "senado",
  "03",     "1989", "10",   "Senate",              "senado",
  "03",     "1993", "06",   "Senate",              "senado",
  "03",     "1996", "03",   "Senate",              "senado",
  "03",     "2000", "03",   "Senate",              "senado",
  "03",     "2004", "03",   "Senate",              "senado",
  "03",     "2008", "03",   "Senate",              "senado",
  "03",     "2011", "11",   "Senate",              "senado",
  "03",     "2015", "12",   "Senate",              "senado",
  "03",     "2016", "06",   "Senate",              "senado",
  "03",     "2019", "04",   "Senate",              "senado",
  "03",     "2019", "11",   "Senate",              "senado",
  "03",     "2023", "07",   "Senate",              "senado",
  "07",     "2014", "05",   "European Parliament", "europeas",
  "07",     "2019", "05",   "European Parliament", "europeas",
  "07",     "2024", "06",   "European Parliament", "europeas"
)
```

CHEQUEAR SI SE PUEDE MEJORAR CODIGO DE DESCARGA DE DATOS GENERALS

types of processes: 02 = congress, 03 = Senate, 07 = European Parliament

```{r ADD SOURCE AND PATH OF EACH ELECTION}
elections <-
  elections |>
  mutate(
    folder    = str_c(process, year, month, "_TOTA"),
    code      = str_c(process, str_sub(year, 3, 4), month),
    on_github = !(year %in% c("2023", "2024") |
                    (chamber == "Congress" & year %in% c("1989", "1993"))),
    path      = if_else(on_github,
                        str_c(GITHUB, "/", mirror_folder, "/", folder, "/"),
                        str_c(tempdir(), "/", folder, "/"))
  )

```

```{r DOWNLOAD NON-GITHUB DATA TO TEMPORARY FOLDER}
for (folder in elections$folder[!elections$on_github]) {
  zip_file <- tempfile(fileext = ".zip")
  download.file(str_c(PORTAL, "/", folder, ".zip"), zip_file, mode = "wb",
                method = "libcurl", headers = c("User-Agent" = "Mozilla/5.0"))
  unzip(zip_file, exdir = str_c(tempdir(), "/", folder), junkpaths = TRUE)
}
```

```{r READ CANDIDATES FROM PATH}

# candidates of every election (file 04)
candidates <- elections |>
  rowwise() |>
  reframe(
    folder, chamber, year, month,
    read_fwf(str_c(path, "04", code, ".DAT"), cols_candidates,
             locale = locale(encoding = "ISO-8859-1"), col_types = cols(.default = "c"))
  )

candidates
```

```{r JOIN PARTIES TO CANDIDATES}
# parties of every election (file 03)
parties <- elections |>
  rowwise() |>
  reframe(
    folder,
    read_fwf(str_c(path, "03", code, ".DAT"), cols_parties,
             locale = locale(encoding = "ISO-8859-1"), col_types = cols(.default = "c"))
  )

all_candidates <-
  candidates |> 
  left_join(parties,   by = c("folder", "party_code"))

all_candidates
```

```{r JOIN PROVINCE TO CANDIDATES}
provinces <- tribble(
  ~province_code, ~province,
  "01", "Araba/Alava",    "02", "Albacete",      "03", "Alicante",
  "04", "Almeria",        "05", "Avila",         "06", "Badajoz",
  "07", "Illes Balears",  "08", "Barcelona",     "09", "Burgos",
  "10", "Caceres",        "11", "Cadiz",         "12", "Castellon",
  "13", "Ciudad Real",    "14", "Cordoba",       "15", "A Coruna",
  "16", "Cuenca",         "17", "Girona",        "18", "Granada",
  "19", "Guadalajara",    "20", "Gipuzkoa",      "21", "Huelva",
  "22", "Huesca",         "23", "Jaen",          "24", "Leon",
  "25", "Lleida",         "26", "La Rioja",      "27", "Lugo",
  "28", "Madrid",         "29", "Malaga",        "30", "Murcia",
  "31", "Navarra",        "32", "Ourense",       "33", "Asturias",
  "34", "Palencia",       "35", "Las Palmas",    "36", "Pontevedra",
  "37", "Salamanca",      "38", "S.C. Tenerife", "39", "Cantabria",
  "40", "Segovia",        "41", "Sevilla",       "42", "Soria",
  "43", "Tarragona",      "44", "Teruel",        "45", "Toledo",
  "46", "Valencia",       "47", "Valladolid",    "48", "Bizkaia",
  "49", "Zamora",         "50", "Zaragoza",      "51", "Ceuta",
  "52", "Melilla",        "99", "Single national constituency"
)

 all_candidates <-
  all_candidates |>
  left_join(provinces, by = "province_code")
 
 all_candidates 
```

```{r MUTATE AND SELECT VARIABLES OF ALL_CANDIDATES}

all_candidates <-
  all_candidates |> 
  mutate(
    election      = str_c(if_else(chamber == "European Parliament", "european", "general"),
                          "_", year, "_", month),
    year          = as.integer(year),
    chamber,
    province,
    list_position = as.integer(list_position),
    candidate     = candidate |>
           str_to_upper() |>
           stringi::stri_trans_general("Latin-ASCII") |>
           str_remove_all("[^A-Z ]") |>
           str_squish(),
    sex           = case_match(sex, "M" ~ "Male", "F" ~ "Female"),
    is_titular    = candidate_type == "T",
    is_elected    = elected == "S",
    abbreviation,
    level         = "National",
    .keep = "none"
  )
all_candidates
```




```{r TAKE DON/ DOÑA OUT}
all_candidates <- 
  all_candidates |>
  mutate(candidate = str_remove(candidate,
                                regex("^(don|do[ñn]a|d[.ª])\\s+", ignore_case = TRUE)))
```

```{r FILTER PP, CS AND VOX}
pp_candidates <- all_candidates |> 
  filter(str_detect(abbreviation, "PP"))

vox_candidates <- all_candidates |> 
  filter(str_detect(abbreviation, "VOX"))

cs_candidates <- all_candidates |> 
  filter(str_detect(abbreviation, "C's"))
```


```{r FILTER HISTORICAL RADICAL RIGHT PARTIES}
radical_right_candidates <- bind_rows(
  read_excel("CANDIDATOS_PARTIDOS_HISTORICOS.xlsx", sheet = "01 Generales"),
  read_excel("CANDIDATOS_PARTIDOS_HISTORICOS.xlsx", sheet = "02 Europeas"))

```

## Autonomical Elections

Claude Code For Autonomical elections: two attempts to confirm it is reproducible.

Prompt 1:

"Quiero que busques todas las listas de todas las elecciones autonómicas que han habido en España desde 2015 hasta 2026 y que pongas en el archivo excel en el excel de CANDIDATOS_AUTONOMICAS. Tienes que ir a la web para recopilar las siguientes variables de cada eleccion: CANDIDATO, CCAA, ELECCION, CIRCUNSCRIPCION, ORDEN, ELEGIDO (TRUE/FALSE)"

Prompt 2:

"You're a Computational Social Scientist. I want you to search for every candidate list of all autonomical elections in Spain from 2015 to 2026 of the following parties: PP and VOX. You have to get into the web and recopilate the following variables for each election:

election: election (write it in like this: "autonomous_2015_05", where 2015 would be the year and 05 the month where the election took place) year: year of the election chamber: the chamber (such as Parlament of Andalucia, Parlament of Catalonia...) province: the province that was being represented by the candidate list_position: the position of the candidate within that list candidate: the candidate's full name is_titular: whether he was the titular of that seat or he was a substitution candidate (TRUE/FALSE), is_elected: whether he was elected (TRUE/FALSE) abbreviation: the abbreviation of the party's name

the final file should be a csv file."




```{r AUTONOMICAL CANDIDATES 2000-2012 PP}
autonomical_pre2015 <-
  read_csv("autonomicas_scrape/out_pre2015/pp_autonomicas_2000_2012.csv") |> 
  mutate(
    election,
    year          = as.integer(year),
    chamber,
    province,
    list_position = as.integer(list_position),
    candidate     = candidate |>
           str_to_upper() |>
           stringi::stri_trans_general("Latin-ASCII") |>
           str_remove_all("[^A-Z ]") |>
           str_squish(),
    sex,
    is_titular,
    is_elected,   
    abbreviation,
    level         = "Autonomical",
    .keep = "none")

autonomical_pre2015
```

```{r AUTONOMICAL CANDIDATES 2015-2026 PP AND VOX}
autonomical_2015_2026 <- 
  read_csv("candidatos_pp_vox_autonomicas_2015_2026.csv") |>
  mutate(
    election,
    year          = as.integer(year),
    chamber,
    province,
    list_position = as.integer(list_position),
    candidate     = candidate |>
           str_to_upper() |>
           stringi::stri_trans_general("Latin-ASCII") |>
           str_remove_all("[^A-Z ]") |>
           str_squish(),
    sex,
    is_titular,
    is_elected,
    abbreviation,
    level         = "Autonomical",
    .keep = "none")
```

```{r FILTER PP AND VOX}
autonomical_candidates <- bind_rows(autonomical_pre2015, autonomical_2015_2026)

pp_candidates__autonomical <- autonomical_candidates |> 
  filter(str_detect(abbreviation, "PP"))

vox_candidates_autonomical <- autonomical_candidates |> 
  filter(str_detect(abbreviation, "VOX"))

vox_candidates_autonomical
```

```{r READ CIUDADANOS CANDIDATES AUTONOMICAL 2015-2026}
cs_candidates_autonomical <- 
  read_csv("candidatos_cs_autonomicas_2015_2026.csv") |> 
  select(election, year, chamber, province, list_position, candidate, sex, is_titular, is_elected, abbreviation) |> 
  mutate(sex = if_else(str_detect(sex, "Mujer"), "Female", "Male"))

cs_candidates_autonomical
```


```{r RADICAL RIGHT CANDIDATES}
radical_right_candidates <- bind_rows(radical_right_candidates,
  read_excel("CANDIDATOS_PARTIDOS_HISTORICOS.xlsx", sheet = "03 Autonomicas")
) |>
  mutate(candidate = candidate |>
           str_to_upper() |>
           stringi::stri_trans_general("Latin-ASCII") |>
           str_remove_all("[^A-Z ]") |>
           str_squish())

```

```{r JOIN GENERAL ELECTIONS AND AUTONOMICAL}
pp_candidates <-
  bind_rows(pp_candidates, pp_candidates__autonomical)

vox_candidates <-
  bind_rows(vox_candidates, vox_candidates_autonomical)

cs_candidates <-
  bind_rows(cs_candidates, cs_candidates_autonomical)
```


# Municipal Elections


```{r READ ALL VOX MUNICIPAL CANDIDATES}
# municipality -> province, from the INE list (one sheet per province). The key
# ignores accents, case and the order of the bilingual halves, because the file
# writes "Alicante/Alacant" and the INE "Alacant/Alicante".
municipalities <- list()
for (s in excel_sheets("26codmun.xlsx")) {
  municipalities[[s]] <- read_excel("26codmun.xlsx", sheet = s, skip = 2,
                                    col_types = "text") |>
    filter(!is.na(CMUN)) |>
    transmute(province_code = CPRO, municipality = NOMBRE)
}
municipalities <- bind_rows(municipalities) |>
  left_join(provinces, by = "province_code") |>
  mutate(muni_key = municipality |>
           str_to_upper() |>
           stringi::stri_trans_general("Latin-ASCII") |>
           str_split("/") |>
           map_chr(~ str_c(sort(str_squish(.x)), collapse = "/"))) |>
  distinct(muni_key, province)

municipal_vox <- readxl::read_xlsx("CANDIDATOS_MUNICIPALES.xlsx") |>
  mutate(muni_key = CIRCUNSCRIPCION |>
           str_to_upper() |>
           stringi::stri_trans_general("Latin-ASCII") |>
           str_split("/") |>
           map_chr(~ str_c(sort(str_squish(.x)), collapse = "/"))) |>
  left_join(municipalities, by = "muni_key", relationship = "many-to-many")

# a municipality name shared by several provinces ("Villanueva", "Castrillo"...)
# duplicates the rows: keep the province that belongs to the list's region, taken
# from the names that are not shared
region_provinces <- municipal_vox |>
  group_by(ELECCION, CIRCUNSCRIPCION, ORDEN) |>
  filter(n() == 1) |>
  ungroup() |>
  distinct(CCAA, province)

municipal_vox <- municipal_vox |>
  group_by(ELECCION, CIRCUNSCRIPCION, ORDEN) |>
  filter(n() == 1 | province %in% region_provinces$province[region_provinces$CCAA == first(CCAA)]) |>
  slice_head(n = 1) |>
  ungroup() |>
  transmute(
    election      = ELECCION,
    year          = as.integer(str_extract(ELECCION, "\\d{4}")),
    chamber       = CIRCUNSCRIPCION,   # the municipality
    province,                          # its province, from the INE list
    list_position = as.integer(ORDEN),
    candidate     = CANDIDATO,
    is_elected    = ELEGIDO,
    source        = FUENTE,
    level         = "Municipal"
  )
# municipalities the INE list no longer carries under that name (it is dated
# 1 January 2026 and these data reach 2015), assigned by hand
manual_provinces <- tribble(
  ~chamber,                       ~province_manual,
  "Castell-Platja d'Aro",         "Girona",
  "Sant Carles de la Ràpita",     "Tarragona",
  "Moncada",                      "Valencia",
  "Alfarp",                       "Valencia",
  "Castelló de la Plana",         "Castellon",
  "Guadiana del Caudillo",        "Badajoz",
  "Maó-Mahón",                    "Illes Balears"
)

municipal_vox <- municipal_vox |>
  left_join(manual_provinces, by = "chamber") |>
  mutate(province = coalesce(province, province_manual)) |>
  select(-province_manual)
```

VER SI SOLO CONCURRIO EN 646 municipios de 8116 en 2019

```{r READ ALL PP MUNICIPAL CANDIDATES}
# PP candidates in the municipal elections. The ministry's file 04 for municipal
# elections carries the municipality, and up to 2003 the name is in one field, so
# read 26-100 as a single block: in the split layout the three fields are adjacent
# and str_squish collapses the padding to the same result.
PORTAL <- "https://infoelectoral.interior.gob.es/estaticos/docxl/apliextr"
dir.create("datos_pp_locales", showWarnings = FALSE)

cols_04_local <- fwf_cols(
  province_code  = c( 10,  11),
  muni_code      = c( 13,  15),   # 12 is a control digit
  party_code     = c( 16,  21),
  list_position  = c( 22,  24),
  candidate_type = c( 25,  25),
  full_name_cand = c( 26, 100),
  sex            = c(101, 101),
  elected        = c(120, 120)
)
latin1 <- locale(encoding = "ISO-8859-1")

# file 03: candidacies (parties)
cols_03 <- fwf_cols(
  party_code   = c(  9,  14),
  abbreviation = c( 15,  64),
  full_name    = c( 65, 214)
)

# municipality names, one sheet per province
municipalities <- list()
for (s in excel_sheets("26codmun.xlsx")) {
  municipalities[[s]] <- read_excel("26codmun.xlsx", sheet = s, skip = 2,
                                    col_types = "text")
}
municipalities <- bind_rows(municipalities) |>
  filter(!is.na(CMUN)) |>
  transmute(province_code = CPRO, muni_code = CMUN, municipality = NOMBRE)

local_candidates <- list()
years <- c("2003", "2007", "2011", "2015", "2019", "2023")

for (i in seq_along(years)) {

  folder   <- str_c("04", years[i], "05_TOTA")
  zip_file <- str_c("datos_pp_locales/", folder, ".zip")
  unzip_to <- tempfile()

  if (!file.exists(zip_file))
    download.file(str_c(PORTAL, "/", folder, ".zip"), zip_file, mode = "wb", quiet = TRUE)
  unzip(zip_file, exdir = unzip_to)

  path_03 <- list.files(unzip_to, "^03.*[.]DAT$", full.names = TRUE,
                        recursive = TRUE, ignore.case = TRUE)[1]
  path_04 <- list.files(unzip_to, "^04.*[.]DAT$", full.names = TRUE,
                        recursive = TRUE, ignore.case = TRUE)[1]

  # the full name is the reliable filter: many local parties are abbreviated PPx
  # ("PPN" is Partido Progresista Nerjeño, "P.P.SO." Plataforma del Pueblo Soriano)
  parties <- read_fwf(path_03, cols_03, locale = latin1,
                      col_types = cols(.default = "c")) |>
    mutate(across(everything(), str_trim)) |>
    filter(str_detect(stringi::stri_trans_general(str_to_upper(full_name),
                                                  "Latin-ASCII"),
                      "PARTIDO POPULAR|PARTIT POPULAR"))

  message("Reading municipal ", years[i], " ...")

  local_candidates[[i]] <- read_fwf(path_04, cols_04_local, locale = latin1,
                                    col_types = cols(.default = "c")) |>
    mutate(across(everything(), str_trim)) |>
    inner_join(parties, by = "party_code") |>
    mutate(year = years[i])
}

municipal_pp <- bind_rows(local_candidates) |>
  left_join(municipalities, by = c("province_code", "muni_code")) |>
  left_join(provinces, by = "province_code") |>
  transmute(
    election      = str_c("municipal_", year, "_05"),
    year          = as.integer(year),
    chamber       = municipality,
    province,
    list_position = as.integer(list_position),
    candidate     = str_squish(full_name_cand),
    sex           = recode(sex, "M" = "Male", "F" = "Female", .default = NA_character_),
    is_titular    = candidate_type == "T",
    is_elected    = elected == "S",
    abbreviation  = "PP",
    level         = "Municipal"
  ) |>
  arrange(year, province, chamber, list_position)
```

```{r READ ALL CS MUNICIPAL CANDIDATES}
municipal_cs <- read_excel("CANDIDATOS_MUNICIPALES_CS.xlsx")
```

```{r READ ALL RADICAL RIGHT MUNICIPAL CANDIDATES}
radical_right_candidates <- bind_rows(radical_right_candidates,
    read_excel("CANDIDATOS_PARTIDOS_HISTORICOS.xlsx", sheet = "04 Municipales"),
  read_excel("CANDIDATOS_PARTIDOS_HISTORICOS.xlsx", sheet = "05 Cabildos insulares")
) |>
  mutate(person_id = candidate |>
           str_to_upper() |>
           stringi::stri_trans_general("Latin-ASCII") |>
           str_remove_all("[^A-Z ]") |>
           str_squish())
```

```{r JOIN GENERAL/AUTONOMICAL ELECTIONS AND AUTONOMICAL}
pp_candidates <-
  bind_rows(pp_candidates, municipal_pp)

vox_candidates <-
  bind_rows(vox_candidates, municipal_vox)

cs_candidates <-
  bind_rows(cs_candidates, municipal_cs)

```

```{r NORMALIZE NAMES AND ELECTION DATE}
vox_candidates <- vox_candidates |>
  mutate(
    # crude person id: uppercase, no accents, no punctuation
    person_id = candidate |>
      str_to_upper() |>
      stringi::stri_trans_general("Latin-ASCII") |>
      str_remove_all("[^A-Z ]") |>
      str_squish()
    # Date?
  ) 

pp_candidates <- pp_candidates |>
  mutate(
    # crude person id: uppercase, no accents, no punctuation
    person_id = candidate |>
      str_to_upper() |>
      stringi::stri_trans_general("Latin-ASCII") |>
      str_remove_all("[^A-Z ]") |>
      str_squish()
    # Date?
  ) 

cs_candidates <- cs_candidates |>
  mutate(
    # crude person id: uppercase, no accents, no punctuation
    person_id = candidate |>
      str_to_upper() |>
      stringi::stri_trans_general("Latin-ASCII") |>
      str_remove_all("[^A-Z ]") |>
      str_squish()
    # Date?
  ) 

radical_right_candidates <- radical_right_candidates |>
  mutate(
    # crude person id: uppercase, no accents, no punctuation
    person_id = candidate |>
      str_to_upper() |>
      stringi::stri_trans_general("Latin-ASCII") |>
      str_remove_all("[^A-Z ]") |>
      str_squish()
    # Date?
  ) 

```

QUITAR ESTO PARA QUE YA ESTÉ ASI EN TODOS.

```{r ABBREVIATIONS OF RADICAL RIGHT PARTIES}
# one abbreviation per party (the printed ones vary: "DN", "D.N.", ...)
party_abbreviations <- tribble(
  ~party,                              ~prev_party,
  "ADN (coalicion 2019)",              "ADÑ",
  "Agrupacion Ruiz-Mateos",            "ARM",
  "Alianza Nacional",                  "AN",
  "Alternativa de Govern",             "AG",
  "Alternativa Democrata Nacional",    "ADN",
  "Alternativa Espanola",              "AES",
  "Centro Democratico y Social",       "CDS",
  "Comunion Tradicionalista Carlista", "CTC",
  "Democracia Nacional",               "DN",
  "Espana 2000",                       "E2000",
  "FE de las JONS",                    "FE-JONS",
  "Falange 2000",                      "FE-2000",
  "Falange Autentica",                 "FA",
  "Frente Nacional (2006)",            "FN",
  "Fuerza Nueva",                      "FN",
  "Grupo Independiente Liberal",       "GIL",
  "Hacer Nacion",                      "HN",
  "Juntas Espanolas",                  "JJEE",
  "La Falange",                        "LF",
  "Movimiento Social Republicano",     "MSR",
  "Nosotros",                          "NOSOTROS",
  "PADE",                              "PADE",
  "Partido por la Libertad",           "PxL",
  "Plataforma per Catalunya",          "PxC",
  "Sociedad Civil y Democracia",       "SCD",
  "Som Identitaris",                   "SOMI",
  "Union del Pueblo Navarro",          "UPN",
  "Via Democratica",                   "VD"
)
```


## Vox's Satellite organizations 

```{r VOX'S SATELLITE ORGANIZATION MEMBERS}

pages <- tribble(
  ~organisation, ~prefix,
  "DENAES",      "denaes.es/quienes-somos",
  "Disenso",     "fundaciondisenso.org/nosotros",
  "ISSEP",       "issep.es/nosotros",
  "ISSEP",       "issep.es/profesores",
  "HazteOir",    "hazteoir.org/conocenos",
  "HazteOir",    "www.hazteoir.org/conocenos",
  "CitizenGO",   "citizengo.org/es/conocenos",
  "CitizenGO",   "www.citizengo.org/es/conocenos",
  "CitizenGO",   "citizengo.org/en/about-us",
  "Solidaridad", "sindicatosolidaridad.es/contacto"
)

# --- 1. what the Archive has --------------------------------------------------
captures <- pages |>
  mutate(cdx = map(prefix, \(p)
    request("http://web.archive.org/cdx/search/cdx") |>
      req_url_query(url = p, matchType = "prefix", output = "json",
                    fl = "timestamp,original", collapse = "digest",
                    filter = c("statuscode:200", "mimetype:text/html"),
                    limit = 2000, .multi = "explode") |>
      req_retry(max_tries = 3) |>
      req_perform() |> resp_body_json() |>
      (\(x) tibble(timestamp = map_chr(x[-1], 1), original = map_chr(x[-1], 2)))())) |>
  unnest(cdx) |>
  mutate(date = ymd(str_sub(timestamp, 1, 8)), year = year(date),
         snapshot = str_c("https://web.archive.org/web/", timestamp, "id_/", original)) |>
  # one capture per page and year is enough to follow who is on a board
  slice_max(date, n = 1, by = c(original, year), with_ties = FALSE)

# --- 2. download them in parallel ---------------------------------------------
responses <- captures$snapshot |>
  map(\(u) request(u) |> req_timeout(30) |> req_error(is_error = \(r) FALSE)) |>
  req_perform_parallel(max_active = 5, on_error = "continue", progress = TRUE)

lines <- captures |>
  mutate(html = map_chr(responses, possibly(\(r) resp_body_string(r), NA_character_))) |>
  filter(!is.na(html)) |>
  # only the code is dropped. These sites wrap their content in <header> and
  # <aside>, so removing those takes the board away with the menu; the census
  # test below is what actually separates people from navigation.
  mutate(text = map(html, \(h) {
    page <- read_html(h)
    page |> html_elements("script, style, noscript") |> xml2::xml_remove()
    page |> html_elements("body") |> html_text2() |> str_split_1("\n") |> str_squish()
  }), .keep = "unused") |>
  unnest(text) |>
  filter(text != "")

# --- 3. which of those lines is a person --------------------------------------
PARTICLES <- c("de", "del", "la", "las", "los", "y", "e", "da", "do", "van", "von", "i")

# labels, menus and headlines the capitalisation rule would otherwise accept
NOT_A_PERSON <- str_c(
  "EQUIPO|CONSEJO|PATRONATO|DIRECCION|COMITE|JUNTA|FUNDACION|INSTITUTO|SINDICATO|",
  "ASOCIACION|UNIVERSIT|SCIENCES PO|COLLEGE|SCHOOL|REVISTA|DIARIO|DIGEST|OBJECTIVE|",
  "TELEVISION|NEWSLETTER|GACETA|BLOG|ESCUELA|CENTRO|OBSERVATORIO|PREMIO|PROGRAMA|",
  "CURSO|JORNADA|CONGRESO|ESPANA|NACIONAL|INTERNACIONAL|MOVIMIENTO|INICIO|CONTACTO|",
  "PRENSA|AGENDA|BOLETIN|TIENDA|AVISO|LEGAL|PRIVACIDAD|COOKIES|DISCURSO|ENTREVISTA|",
  "ARTICULO|CONFERENCIA|PONENCIA|VIDEO|FOUNDATION|CAMPAIGN|MANAGER|IP ADDRESS|",
  "SOBRE NOSOTROS|^LEER|^VER |^FISCAL|ENGLISH|ESPANOL|FRANCAIS|DEUTSCH|",
  "ISSEP|DISENSO|CITIZENGO|HAZTEOIR|DENAES")

# what may follow a name and be kept. The column is "role or description" on
# purpose: these sites print a post ("Presidente") and a profession
# ("Filosofo", "Pediatra") in exactly the same place
ROLES <- str_c(
  "PRESIDENT|VICEPRESIDENT|DIRECTOR|SECRETARI|TESORER|VOCAL|PATRON|COORDINADOR|",
  "RESPONSABLE|GERENTE|PORTAVOZ|FUNDADOR|MIEMBRO|ASESOR|PROFESOR|CATEDRATIC|",
  "DOCENTE|INVESTIGADOR|ANALISTA|CONSEJER|DELEGAD|ESCRITOR|PERIODISTA|FILOSOF|",
  "HISTORIADOR|ABOGAD|ECONOMISTA|MILITAR|EMPRESARI|MAGISTRAD|SOCIOLOG|POLITOLOG|",
  "JURISTA|DECANO|RECTOR|EDITOR|COLUMNISTA|DIPUTAD|SENADOR|EURODIPUTAD")

flat <- function(x) str_to_upper(stringi::stri_trans_general(x, "Latin-ASCII"))

# The real test of whether a line is a person: the INE census lists, which the
# project already uses to assign sex to candidates. A word list of things that
# are NOT names never ends; a list of the names that exist does.
ine_forenames <- c("Hombres", "Mujeres") |>
  map(\(s) read_excel("nombres_por_edad_media.xlsx", sheet = s, skip = 5)$Nombre) |>
  list_c() |> flat() |> str_squish()

# nicknames and rare forenames the census does not carry
ALSO_PEOPLE <- c("CAKE", "EDMALY")

# 2 to 6 words, all of them capitalised except the particles
is_person <- function(x) {
  capitalised <- map_lgl(str_split(x, " "), \(w) {
    w <- w[!str_to_lower(flat(w)) %in% PARTICLES]
    length(w) >= 2 && all(str_detect(w, "^\\p{Lu}\\p{L}"))
  })
  str_length(x) >= 5 & str_length(x) <= 60 &
    str_count(x, "\\S+") >= 2 & str_count(x, "\\S+") <= 6 &
    !str_detect(x, "[0-9@#/|]") & capitalised &
    flat(word(x, 1)) %in% c(ine_forenames, ALSO_PEOPLE) &
    !str_detect(flat(x), NOT_A_PERSON) & !str_detect(flat(x), ROLES)
}

people <- lines |>
  # "Ignacio Arsuaga Rato, Presidente" is a person and their role on one line;
  # without splitting it, the line is thrown away for containing "Presidente"
  separate_wider_delim(text, ", ", names = c("person", "inline_role"),
                       too_many = "merge", too_few = "align_start") |>
  mutate(inline_role = replace_na(inline_role, ""),
         # the other layout puts the role on the line below the name
         next_line = lead(person), .by = c(organisation, original, timestamp)) |>
  filter(is_person(person)) |>
  mutate(role = case_when(
           str_detect(flat(inline_role), ROLES) ~ inline_role,
           str_detect(flat(next_line), ROLES) & str_length(next_line) <= 90 ~ next_line,
           .default = ""),
         # Disenso prints the same board in caps on one page and in title case
         # on another: "AMANDO DE MIGUEL" and "Amando de Miguel" are one person
         person = if_else(person == str_to_upper(person), str_to_title(person), person),
         person = str_replace_all(person, str_c("(?<= )(", str_c(str_to_title(PARTICLES),
                                                                 collapse = "|"), ")(?= )"),
                                  str_to_lower))

# the same person is listed with and without the second surname ("Ana Velasco",
# "Ana Velasco Vidal-Abarca"): the longest form wins
variants <- people |>
  distinct(person) |>
  # compound surnames are opened at the hyphen, so that "Alejo Vidal Quadras"
  # falls inside "Alejo Vidal-Quadras Roca"
  mutate(tokens = str_split(flat(person), "[ -]"), first = map_chr(tokens, 1))

people <- variants |>
  rename(long = person, long_tokens = tokens) |>
  inner_join(rename(variants, short = person, short_tokens = tokens), by = "first",
             relationship = "many-to-many") |>
  filter(map2_lgl(short_tokens, long_tokens, \(s, l) all(s %in% l)),
         lengths(short_tokens) <= lengths(long_tokens)) |>
  slice_max(lengths(long_tokens), n = 1, by = short, with_ties = FALSE) |>
  select(person = short, canonical = long) |>
  right_join(people, by = "person") |>
  mutate(person = coalesce(canonical, person)) |>
  summarise(.by = c(organisation, person),
            roles      = str_c(sort(unique(role[role != ""])), collapse = " | "),
            first_seen = min(date),
            last_seen  = max(date),
            n_captures = n_distinct(timestamp),
            source     = first(str_c("https://web.archive.org/web/", timestamp, "/", original))) |>
  arrange(organisation, person)

# --- 4. Solidaridad, from the BOE ---------------------------------------------
# The union's website archives no team page, so its officers are taken from the
# four announcements of the deposit of its statutes (Direccion General de
# Trabajo, union SPDSTE, deposit number 99105954). Searched at boe.es for
# SPDSTE in the title. Rafael Martinez de la Gandara, who signs all four, is
# the civil servant issuing the resolution and is NOT part of the union.
solidaridad_boe <- tribble(
  ~person,                              ~roles,             ~date,        ~boe,
  "Santiago Alarcón Gordon",            "Solicitante del depósito", "2020-08-12", "BOE-B-2020-26040",
  "Paula Bernardina Hidalgo de la Cruz","Promotora",        "2020-08-12", "BOE-B-2020-26040",
  "Raquel Moreno Barba",                "Promotora",        "2020-08-12", "BOE-B-2020-26040",
  "Ángel García Ochoa",                 "Promotor",         "2020-08-12", "BOE-B-2020-26040",
  "Santiago Alarcón Gordon",            "Solicitante del depósito", "2021-02-15", "BOE-B-2021-7089",
  "Rodrigo Alonso Fernández",           "Secretario General", "2021-02-15", "BOE-B-2021-7089",
  "Javier Tomás de la Cruz Bazo",       "Solicitante del depósito", "2025-05-23", "BOE-B-2025-18656",
  "Jordi Albert de la Fuente Miró",     "Secretario General", "2025-05-23", "BOE-B-2025-18656",
  "Javier Tomás de la Cruz Bazo",       "Solicitante del depósito", "2026-07-03", "BOE-B-2026-22639",
  "Jordi Albert de la Fuente Miró",     "Secretario General", "2026-07-03", "BOE-B-2026-22639"
) |>
  summarise(.by = person,
            organisation = "Solidaridad",
            roles      = str_c(sort(unique(roles)), collapse = " | "),
            first_seen = min(ymd(date)),
            last_seen  = max(ymd(date)),
            n_captures = n(),
            source     = str_c("https://www.boe.es/diario_boe/txt.php?id=", first(boe)))

people <- 
  bind_rows(people, solidaridad_boe) |> 
  arrange(organisation, person) |> 
   mutate(person = person |>
           str_to_upper() |>
           stringi::stri_trans_general("Latin-ASCII") |>
           str_remove_all("[^A-Z ]") |>
           str_squish()) |> 
  filter(!person %in% c("ADOLFO SUAREZ", "YA SOMOS", "SEGUNDO APELLIDO"))

# --- 5. other organisations found in the press and Wikipedia ------------------
# organisations the candidates within the expected seats belonged to before
# (or besides) Vox: federations, foundations, associations, platforms
press_organisations <- read_excel("ORGANIZACIONES_CANDIDATOS_VOX.xlsx") |>
  select(person = PERSON_ID, organisation = ORGANIZACION, roles = ROL, source = FUENTE)

people <- people |>
  bind_rows(press_organisations) |>
  arrange(organisation, person)

# --- 6. Revuelta and Homo Legens, from the press ------------------------------
# Revuelta: youth association launched in October 2023 by five Vox affiliates and
# party employees, who formed its board; the board is the one reported when two of
# them took the association to the prosecutor over the DANA donations.
# Homo Legens: publishing house of the Intereconomia group, founded in 2017 by the
# former PP deputy Julio Ariza and run by his son Gabriel; it publishes the books of
# Vox leaders and has employed Lidia Bedman, Santiago Abascal's wife, since 2019
REVUELTA_SOURCE    <- "https://www.eldebate.com/espana/20251211/revuelta-asociacion-juvenil-denunciado-vox-claves-polemica_364072.html"
HOMO_LEGENS_SOURCE <- "https://es.wikipedia.org/wiki/Homo_Legens"

press_satellites <- tribble(
  ~person,                      ~organisation, ~roles,
  "Jaime Hernández Zúñiga",     "Revuelta",    "Presidente",
  "Arturo Villarroya González", "Revuelta",    "Vicepresidente",
  "Pablo González Gasca",       "Revuelta",    "Secretario",
  "Javier Esteban Bejarano",    "Revuelta",    "Vocal",
  "Santiago Aneiros",           "Revuelta",    "Vocal | Tesorero de facto",
  "Julio Ariza",                "Homo Legens", "Fundador",
  "Gabriel Ariza",              "Homo Legens", "Director",
  "Kiko Méndez-Monasterio",     "Homo Legens", "Figura clave de la editorial",
  "Lidia Bedman",               "Homo Legens", "Asesora de redes sociales (desde 2019)"
) |>
  mutate(source = if_else(organisation == "Revuelta", REVUELTA_SOURCE, HOMO_LEGENS_SOURCE),
         person = person |>
           str_to_upper() |>
           stringi::stri_trans_general("Latin-ASCII") |>
           str_remove_all("[^A-Z ]") |>
           str_squish())

people <- people |>
  bind_rows(press_satellites) |>
  arrange(organisation, person)
```


```{r TABLE OF VOX'S SATELLITE ORGANIZATIONS}
# the organisations around Vox whose members were collected above, with how many of
# them could be identified and how many of those have stood on a Vox list. The
# organisations of step 5 (federations, foundations, associations the candidates
# belonged to before Vox) are not satellites and stay out of the table
satellite_organisations <- tribble(
  ~organisation, ~type,                     ~founded, ~link_with_vox,
  "DENAES",      "Foundation",              2006,     "Founded and chaired by Santiago Abascal before Vox",
  "HazteOir",    "Association (lobby)",     2001,     "Ultra-Catholic lobby, ally of Vox",
  "CitizenGO",   "Online platform",         2013,     "International platform of HazteOir",
  "Homo Legens", "Publishing house",        2017,     "Publisher of Vox leaders (Intereconomia group)",
  "Disenso",     "Foundation (think tank)", 2020,     "Think tank of Vox, chaired by Santiago Abascal",
  "ISSEP",       "Training institute",      2020,     "Madrid branch of Marion Marechal's institute; trains Vox cadres",
  "Solidaridad", "Trade union",             2020,     "Trade union promoted by Vox",
  "Revuelta",    "Youth association",       2023,     "Launched by Vox affiliates and employees"
)

satellite_members <- people |>
  filter(organisation %in% satellite_organisations$organisation) |>
  summarise(members        = n_distinct(person),
            vox_candidates = n_distinct(person[person %in% vox_candidates$person_id]),
            .by = organisation)

satellite_table <- satellite_organisations |>
  left_join(satellite_members, by = "organisation") |>
  mutate(across(c(members, vox_candidates), \(x) replace_na(x, 0L))) |>
  arrange(founded) |>
  # as text: stargazer prints numbers with a thousands separator ("2,006")
  mutate(founded = as.character(founded)) |>
  select(Organisation = organisation, Type = type, Founded = founded, `Link with Vox` = link_with_vox,
         `Members identified` = members, `Vox candidates` = vox_candidates)

stargazer(as.data.frame(satellite_table), summary = FALSE, rownames = FALSE, type = table_type,
          title = "Vox's satellite organisations",
          notes = c("Members identified: board members and staff found in the archived websites (Wayback Machine),",
                    "the BOE (Solidaridad) and the press (Revuelta, Homo Legens). Vox candidates: members who stood on a Vox list."))

# who they are
people |>
  filter(organisation %in% satellite_organisations$organisation,
         person %in% vox_candidates$person_id) |>
  distinct(organisation, person, roles) |>
  arrange(organisation, person) |>
  print(n = Inf)
```


```{r ISSEP}

```


```{r CLEAN HTML IN CANDIDATE NAMES}
# some names come with the markup of the page they were scraped from:
# "Aitor&nbsp;<span class="searchMatch">Caneda</span>&nbsp;Manso". The markup is
# read as HTML (tags dropped, entities decoded) and person_id is built again
vox_candidates <- vox_candidates |>
  mutate(candidate = map_chr(candidate, \(x)
           if (!is.na(x) && str_detect(x, "[<&]")) xml2::xml_text(xml2::read_html(str_c("<p>", x, "</p>"))) else x) |>
           str_squish(),
         person_id = candidate |>
           str_to_upper() |>
           stringi::stri_trans_general("Latin-ASCII") |>
           str_remove_all("[^A-Z ]") |>
           str_squish())

# must return 0 rows
vox_candidates |> filter(str_detect(candidate, "[<>&;]"))
```


## Data Enrichment

```{r HARMONIZE SEX}
# every source writes sex its own way ("M"/"F", "Hombre"/"Mujer", "Male"/"Female")
# and the municipal lists do not carry it at all. Harmonised to Female/Male, and
# filled, in this order: 1. what the list says, 2. what the other lists of the same
# person say, 3. the forename, with the INE census of names

# 1. the forenames of the INE census, each with the sex of most of its holders
ine_forenames_sex <- bind_rows(
  read_excel("nombres_por_edad_media.xlsx", sheet = "Hombres", skip = 5) |> mutate(sex_forename = "Male"),
  read_excel("nombres_por_edad_media.xlsx", sheet = "Mujeres", skip = 5) |> mutate(sex_forename = "Female")) |>
  mutate(forename = Nombre |>
           str_to_upper() |>
           stringi::stri_trans_general("Latin-ASCII") |>
           str_remove_all("[^A-Z ]") |>
           str_squish()) |>
  filter(!is.na(forename)) |>
  slice_max(Frecuencia, n = 1, by = forename, with_ties = FALSE) |>
  select(forename, sex_forename, frequency = Frecuencia) |>
  arrange(desc(frequency))

# 2. what each list says, harmonised ("MUJER" is checked before the "M" of Male)
vox_candidates <- vox_candidates |>
  mutate(sex = case_when(str_detect(str_to_upper(sex), "^(F|MUJER|FEMALE|FEMENINO)") ~ "Female",
                         str_detect(str_to_upper(sex), "^(M|H|MALE|HOMBRE|MASCULINO|VARON)") ~ "Male"))

# 3. the sex the person has in their other lists
sex_by_person <- vox_candidates |>
  filter(!is.na(sex)) |>
  count(person_id, sex) |>
  slice_max(n, n = 1, by = person_id, with_ties = FALSE) |>
  select(person_id, sex_person = sex)

# 4. the sex of the forename, trying in this order:
#    a. the first two words ("MARIA JOSE", "JOSE MARIA"); "M"/"MA" stand for Maria
#    b. the first word
#    c. forenames written together ("JUANCARLOS", "MARIAJESUS"): the longest census
#       forename the first word starts with
#    d. the second forename, skipping particles ("LEONIA ALEXANDRA", "J DEL ROCIO")
#    e. foreign or rare forenames the census does not carry, set by hand below
#    f. a misspelt first word ("PACUAL", "MAGADALENA"): the census forename one letter apart
#    g. what is left: ending in -A, Female; otherwise, Male
census_forenames <- ine_forenames_sex |>
  filter(frequency >= 100, !str_detect(forename, " "))

forename_corrections <- tribble(
  ~first_one,     ~sex_e,
  "ANABEIDY",     "Female", "ANATERESA",  "Female", "ANAVANESA",  "Female", "AHIATIMA",   "Female",
  "AUVELLY",      "Female", "CARUCHA",    "Female", "DOBROSLAVA", "Female", "EMERENCIA",  "Female",
  "EUDIVIGIS",    "Female", "EUKARIS",    "Female", "FAITIA",     "Female", "GLISELIA",   "Female",
  "IBOLYAKINGA",  "Female", "ISOLGITA",   "Female", "LUCIVANIA",  "Female", "MAGACEDA",   "Female",
  "MALLAURY",     "Female", "MARIUXI",    "Female", "MERCHINES",  "Female", "MISSANDRA",  "Female",
  "NAOULA",       "Female", "OSDANAY",    "Female", "OSLAIDA",    "Female", "SLADANA",    "Female",
  "VALBONA",      "Female", "YULEISIS",   "Female", "IDERMIS",    "Female", "ENERIET",    "Female",
  "ARGYRIS",      "Male",   "BEDROS",     "Male",   "CANCIANO",   "Male",   "CESIDIO",    "Male",
  "GARBEIL",      "Male",   "HARTIOUN",   "Male",   "KEFAH",      "Male",   "OLIVEIRO",   "Male",
  "SANTE",        "Male",   "SIQUEM",     "Male",   "TRAJANO",    "Male",   "TROYANO",    "Male",
  "VANGELIS",     "Male",   "AMROU",      "Male"
)

sex_by_forename <- vox_candidates |>
  distinct(person_id) |>
  mutate(name      = str_replace(person_id, "^MA? ", "MARIA "),
         first_two = word(name, 1, 2),
         first_one = word(name, 1),
         second    = word(str_squish(str_remove_all(name, "\\b(DE|DEL|LA|LAS|LOS)\\b")), 2)) |>
  left_join(ine_forenames_sex |> select(first_two = forename, sex_a = sex_forename), by = "first_two") |>
  left_join(ine_forenames_sex |> select(first_one = forename, sex_b = sex_forename), by = "first_one") |>
  left_join(ine_forenames_sex |> select(second = forename, sex_d = sex_forename), by = "second")

written_together <- sex_by_forename |>
  filter(is.na(sex_a), is.na(sex_b)) |>
  distinct(first_one) |>
  cross_join(census_forenames |> filter(str_length(forename) >= 4) |> select(forename, sex_forename)) |>
  filter(str_starts(first_one, forename), first_one != forename) |>
  slice_max(str_length(forename), n = 1, by = first_one, with_ties = FALSE) |>
  select(first_one, sex_c = sex_forename)

misspelt <- sex_by_forename |>
  filter(is.na(sex_a), is.na(sex_b), is.na(sex_d)) |>
  distinct(first_one)

distances <- adist(misspelt$first_one, census_forenames$forename)

misspelt <- misspelt |>
  mutate(closest  = census_forenames$forename[apply(distances, 1, which.min)],
         distance = apply(distances, 1, min),
         sex_f    = if_else(distance == 1, census_forenames$sex_forename[apply(distances, 1, which.min)], NA))

sex_by_forename <- sex_by_forename |>
  left_join(written_together, by = "first_one") |>
  left_join(misspelt |> select(first_one, closest, sex_f), by = "first_one") |>
  left_join(forename_corrections, by = "first_one") |>
  mutate(sex_g    = if_else(str_ends(first_one, "A"), "Female", "Male"),
         sex_name = coalesce(sex_a, sex_b, sex_c, sex_d, sex_e, sex_f, sex_g),
         sex_rule = case_when(!is.na(sex_a) ~ "a", !is.na(sex_b) ~ "b", !is.na(sex_c) ~ "c",
                              !is.na(sex_d) ~ "d", !is.na(sex_e) ~ "e", !is.na(sex_f) ~ "f", TRUE ~ "g"))

# names resolved by the weaker rules (f, g): check them
sex_by_forename |> filter(sex_rule %in% c("f", "g")) |> select(person_id, closest, sex_name, sex_rule) |> print(n = Inf)

sex_by_forename <- sex_by_forename |>
  select(person_id, sex_name)

vox_candidates <- vox_candidates |>
  select(-any_of(c("sex_person", "sex_name"))) |>
  left_join(sex_by_person, by = "person_id") |>
  left_join(sex_by_forename, by = "person_id") |>
  mutate(sex = coalesce(sex, sex_person, sex_name)) |>
  select(-sex_person, -sex_name)

count(vox_candidates, sex)

# names still without sex: must return 0 rows
vox_candidates |> filter(is.na(sex)) |> distinct(person_id)
```

```{r SIGNEES AND FOUNDERS}
# is_signee: signed the manifesto of reconversion.es; is_founder: in the founders table.
# As with the party organs, a name is found when its words are in the candidate's
# name; when it fits several candidates, the rule below picks one or leaves it out
reference_names <- bind_rows(
  signees_reconversion |> mutate(group = "signee"),
  founders |> select(name) |> mutate(group = "founder")) |>
  mutate(name_id = name |>
           str_to_upper() |>
           stringi::stri_trans_general("Latin-ASCII") |>
           str_remove_all("[^A-Z ]") |>
           str_squish()) |>
  distinct(group, name_id)

reference_words <- reference_names |>
  distinct(name_id) |>
  mutate(word = name_id) |>
  separate_longer_delim(word, " ") |>
  distinct() |>
  add_count(name_id, name = "n_words")

candidate_words <- vox_candidates |>
  distinct(person_id) |>
  mutate(word = person_id) |>
  separate_longer_delim(word, " ") |>
  distinct()

# Vox elections and national or regional lists of each candidate, to break ties
candidate_record <- vox_candidates |>
  group_by(person_id) |>
  summarise(n_elections       = n_distinct(election),
            national_regional = any(level != "Municipal"),
            .groups = "drop")

reference_matches <- reference_words |>
  filter(n_words >= 2) |>
  inner_join(candidate_words, by = "word", relationship = "many-to-many") |>
  count(name_id, person_id, n_words, name = "words_found") |>
  filter(words_found == n_words,
         # the words in the same order and together, except a middle name after the
         # first one: "Pedro Fernandez Barbadillo" is "Pedro Maria Fernandez Barbadillo",
         # but "Enrique Alvarez Lopez" is not "Enrique Lopez Alvarez" and
         # "Jose Maria Alvarez" is not "Jose Maria Figaredo Alvarez Sala"
         str_detect(person_id, str_c("\\b", word(name_id, 1), "\\b(.* )?", word(name_id, 2, -1), "\\b"))) |>
  left_join(candidate_record, by = "person_id") |>
  add_count(name_id, name = "n_candidates")

# names that fit several candidates ("Santiago Abascal": Abascal Conde and his father
# Abascal Escuza): the one who ran in national or regional lists in more elections
chosen <- reference_matches |>
  filter(n_candidates == 1 | national_regional) |>
  slice_max(n_elections, n = 1, by = name_id) |>
  add_count(name_id, name = "n_chosen") |>
  filter(n_chosen == 1)

# the same person under a shorter name ("Ana Velasco Vidalabarca" and
# "Ana Maria Velasco Vidalabarca") is kept as well
same_person <- reference_matches |>
  inner_join(chosen |> select(name_id, chosen_id = person_id), by = "name_id") |>
  filter(person_id != chosen_id,
         map2_lgl(str_split(person_id, " "), str_split(chosen_id, " "), \(p, c) all(p %in% c)))

# names left out: check them by hand
reference_matches |> filter(!name_id %in% chosen$name_id) |> arrange(name_id)

matched <- reference_names |>
  inner_join(bind_rows(chosen, same_person) |> select(name_id, person_id), by = "name_id",
             relationship = "many-to-many")   # a founder can also be a signee

vox_candidates <- vox_candidates |>
  mutate(is_signee  = person_id %in% matched$person_id[matched$group == "signee"],
         is_founder = person_id %in% matched$person_id[matched$group == "founder"])
```

```{r VOX's POSITIONS IN GOVERNMENT}
vox_executive_positions <- tribble(
  ~name,                               ~region,                ~position,                                                                                        ~start_date,  ~end_date,
  # === Castilla y León I (2022-2024) ===
  "JUAN GARCÍA-GALLARDO FRINGS",       "Castilla y León",      "Vicepresidente de la Junta de Castilla y León",                                                  "20/04/2022", "11/07/2024",
  "GONZALO SANTONJA GÓMEZ-AGERO",      "Castilla y León",      "Consejero de Cultura, Turismo y Deporte",                                                        "20/04/2022", "11/07/2024",  # left Vox, stayed in office
  "GERARDO DUEÑAS MERINO",             "Castilla y León",      "Consejero de Agricultura, Ganadería y Desarrollo Rural",                                         "20/04/2022", "11/07/2024",
  "MARIANO VEGANZONES DÍEZ",           "Castilla y León",      "Consejero de Industria, Comercio y Empleo",                                                      "20/04/2022", "11/07/2024",
  # === Comunidad Valenciana (2023-2024) ===
  "VICENTE BARRERA SIMÓ",              "Comunidad Valenciana", "Vicepresidente Primero y Conseller de Cultura y Deporte",                                        "17/07/2023", "11/07/2024",
  "ELISA NÚÑEZ SÁNCHEZ",               "Comunidad Valenciana", "Consellera de Justicia e Interior",                                                              "17/07/2023", "11/07/2024",
  "JOSÉ LUIS AGUIRRE LARRAURI",        "Comunidad Valenciana", "Conseller de Agricultura, Ganadería y Pesca",                                                    "17/07/2023", "11/07/2024",
  # === Extremadura I (2023-2024) ===
  "MARÍA DEL CAMINO LIMIA",            "Extremadura",          "Consejera de Gestión Forestal y Mundo Rural",                                                    "20/07/2023", "05/10/2023",
  "IGNACIO HIGUERO DE JUAN",           "Extremadura",          "Consejero de Gestión Forestal y Mundo Rural",                                                    "05/10/2023", "11/07/2024",  # left Vox, stayed in office
  # === Aragón I (2023-2024) ===
  "ALEJANDRO NOLASCO ASENSIO",         "Aragón",               "Vicepresidente Primero y Consejero de Desarrollo Territorial, Despoblación y Justicia",          "12/08/2023", "12/07/2024",
  "ÁNGEL SAMPER SECORÚN",              "Aragón",               "Consejero de Agricultura, Ganadería y Alimentación",                                             "12/08/2023", "12/07/2024",
  # === Región de Murcia (2023-2024) ===
  "JOSÉ ÁNGEL ANTELO PAREDES",         "Región de Murcia",     "Vicepresidente y Consejero de Interior, Emergencias y Ordenación del Territorio",                "15/09/2023", "12/07/2024",
  "JOSÉ MANUEL PANCORBO DE LA TORRE",  "Región de Murcia",     "Consejero de Fomento e Infraestructuras",                                                        "15/09/2023", "12/07/2024",
  # === Extremadura II (2026-) ===
  "ÓSCAR ARTURO FERNÁNDEZ CALLE",      "Extremadura",          "Vicepresidente y Consejero de Desregulación, Servicios Sociales y Familia",                      "30/04/2026", NA,
  "JUAN JOSÉ GARCÍA LOBATO",           "Extremadura",          "Consejero de Agricultura, Ganadería y Medio Natural",                                            "30/04/2026", NA,
  # === Aragón II (2026-) ===
  "ALEJANDRO NOLASCO ASENSIO",         "Aragón",               "Vicepresidente Primero y Consejero de Desregulación, Bienestar Social y Familia",                "04/05/2026", NA,
  "ARÁNZAZU SIMÓN PÉREZ",              "Aragón",               "Consejera de Agricultura, Ganadería y Alimentación",                                             "04/05/2026", NA,
  "LUIS FRANCISCO BIENDICHO GRACIA",   "Aragón",               "Consejero de Medio Ambiente y Turismo",                                                          "04/05/2026", NA,
  # === Castilla y León II (2026-) ===
  "CARLOS POLLÁN FERNÁNDEZ",           "Castilla y León",      "Vicepresidente Primero y Consejero de Desregulación, Familia y Ayudas Sociales",                 "15/06/2026", NA,
  "JOAQUÍN ANTONIO PINO",              "Castilla y León",      "Consejero de Agricultura, Ganadería, Medio Rural y Política Ambiental",                          "15/06/2026", NA,
  "JOSÉ ALBERTO DÍAZ PICO",            "Castilla y León",      "Consejero de Cultura, Turismo y Deporte",                                                        "15/06/2026", NA,
  # === Andalucía (2026-) ===
  "MANUEL GAVIRA FLORENTINO",          "Andalucía",            "Vicepresidente Segundo y Consejero de Turismo, Justicia, Desregulación y Administración Local",  "10/07/2026", NA
) |>
  # same format as person_id in vox_candidates: no accents, no punctuation
  mutate(name = name |>
           stringi::stri_trans_general("Latin-ASCII") |>
           str_remove_all("[^A-Z ]") |>
           str_squish())

# the regional election after which each government was formed
vox_executive_positions <- vox_executive_positions |>
  mutate(election = case_when(
    region == "Castilla y León"      & str_ends(start_date, "2022") ~ "autonomous_2022_02",
    region == "Castilla y León"      & str_ends(start_date, "2026") ~ "autonomous_2026_03",
    region == "Aragón"               & str_ends(start_date, "2023") ~ "autonomous_2023_05",
    region == "Aragón"               & str_ends(start_date, "2026") ~ "autonomous_2026_02",
    region == "Extremadura"          & str_ends(start_date, "2023") ~ "autonomous_2023_05",
    region == "Extremadura"          & str_ends(start_date, "2026") ~ "autonomous_2025_12",
    region == "Comunidad Valenciana"                                ~ "autonomous_2023_05",
    region == "Región de Murcia"                                    ~ "autonomous_2023_05",
    region == "Andalucía"                                           ~ "autonomous_2026_05"))

# one row per person and election (Limia and Higuero share a post, Nolasco repeats)
executive_by_candidate <- vox_executive_positions |>
  group_by(person_id = name, election) |>
  summarise(executive_position = str_c(position, collapse = " | "), .groups = "drop")

vox_candidates <- vox_candidates |>
  select(-any_of(c("executive_position", "has_executive_position"))) |>
  left_join(executive_by_candidate, by = c("person_id", "election")) |>
  mutate(has_executive_position = !is.na(executive_position))
```


```{r VOX CER MEMBERS y NO SE QUE OTRA COSA}
# one row per organic position, CEN flagged
vox_positions <- read_excel("VOX_CARGOS_ORGANOS.xlsx", sheet = "01 Personas y cargos") |>
  mutate(name       = Nombre,
         position   = Cargo,
         organ      = Organo,
         first_seen = ymd(`Primera constancia`),
         last_seen  = ymd(`Ultima constancia`),
         in_cen     = str_detect(Organo, "^CEN|^Comite Ejecutivo Nacional"),
         name_id    = name |>
           str_to_upper() |>
           stringi::stri_trans_general("Latin-ASCII") |>
           str_remove_all("[^A-Z ]") |>
           str_squish()) |>
  select(name, name_id, position, organ, first_seen, last_seen, in_cen)

# the organs give short names ("Alejandro Nolasco"), the lists full ones
# ("Alejandro Nolasco Asensio"): a name matches a candidate when all its words
# are in the candidate's name, and only if exactly one candidate fits
name_words <- vox_positions |>
  distinct(name_id) |>
  mutate(word = name_id) |>
  separate_longer_delim(word, " ") |>
  distinct() |>
  add_count(name_id, name = "n_words")

candidate_words <- vox_candidates |>
  distinct(person_id) |>
  mutate(word = person_id) |>
  separate_longer_delim(word, " ") |>
  distinct()

name_matches <- name_words |>
  inner_join(candidate_words, by = "word", relationship = "many-to-many") |>
  count(name_id, person_id, n_words, name = "words_found") |>
  filter(words_found == n_words) |>
  add_count(name_id, name = "n_candidates")

# names that fit several candidates: left out, check them by hand
name_matches |> filter(n_candidates > 1) |> arrange(name_id)

# last capture of the VOX website: whoever still appears then is still in the CEN
last_capture <- ymd("2026-09-06")

organ_info <- vox_positions |>
  inner_join(name_matches |> filter(n_candidates == 1) |> select(name_id, person_id), by = "name_id") |>
  group_by(person_id) |>
  summarise(organic_positions = str_c(unique(position), collapse = " | "),
            organ_first_seen  = first(sort(first_seen)),
            organ_last_seen   = last(sort(last_seen)),
            cen_from          = first(sort(first_seen[in_cen])),
            cen_to            = last(sort(last_seen[in_cen])),
            .groups = "drop") |>
  mutate(cen_to = if_else(cen_to == last_capture, NA, cen_to))   # NA = still in the CEN

vox_candidates <- vox_candidates |>
  select(-any_of(c("organic_positions", "organ_first_seen", "organ_last_seen", "in_cen", "cen_from", "cen_to",
                   "has_organic_position", "in_cen_at_election", "in_cen_before"))) |>
  left_join(organ_info, by = "person_id") |>
  mutate(has_organic_position = !is.na(organic_positions),
         # was in the CEN before this election (election month taken as its date)
         in_cen_before        = coalesce(cen_from < ym(str_extract(election, "[0-9]{4}_[0-9]{2}")), FALSE))
```
LIMPIAR y arreglar CEN

```{r SEATS ACTUALLY WON}
# seats Vox won in each list: one list = election + chamber + constituency
# (chamber keeps Congress and Senate apart; in municipal lists chamber is the municipality)
seats_won <- vox_candidates |>
  filter(is_titular %in% c(TRUE, NA)) |>
  group_by(election, chamber, province) |>
  summarise(seats_won = sum(is_elected %in% TRUE), .groups = "drop")

# did the number 1 of this same list (same chamber and constituency) win a seat
# in the previous election of that type? Whoever it was: not matched by name
previous_head_won <- vox_candidates |>
  filter(is_titular %in% c(TRUE, NA), list_position == 1) |>
  mutate(date = str_extract(election, "[0-9]{4}_[0-9]{2}")) |>
  group_by(chamber, province, date) |>
  summarise(head_won = any(is_elected %in% TRUE), .groups = "drop") |>
  arrange(chamber, province, date) |>
  group_by(chamber, province) |>
  mutate(previous_head_won = lag(head_won, default = FALSE)) |>
  ungroup() |>
  select(chamber, province, date, previous_head_won)

vox_candidates <- vox_candidates |>
  select(-any_of(c("seats_won", "top_won"))) |>
  mutate(date = str_extract(election, "[0-9]{4}_[0-9]{2}")) |>
  left_join(seats_won,         by = c("election", "chamber", "province")) |>
  left_join(previous_head_won, by = c("chamber", "province", "date")) |>
  # top_won: heads this list, and the head of this list won a seat last time
  mutate(top_won = is_titular %in% c(TRUE, NA) & list_position == 1 & previous_head_won %in% TRUE) |>
  select(-previous_head_won)
```

```{r VOX EXPECTED SEATS + TOP CANDIDATES}
# ---------------------------------------- 1. CIS estimate for each list (midpoint, rounded up)
# how the CIS spells the provinces that the lists spell differently
same_place <- c("ILLES BALEARS" = "BALEARES",
                "GIPUZKOA"      = "GUIPUZCOA",
                "BIZKAIA"       = "VIZCAYA",
                "ARABA/ALAVA"   = "ALAVA",
                "S.C. TENERIFE" = "SANTA CRUZ DE TENERIFE",
                "CIUDAD"        = "CIUDAD REAL")

cis_seats <- read_csv("cis_estimaciones_vox.csv", show_col_types = FALSE) |>
  mutate(date       = str_extract(election, "[0-9]{4}_[0-9]{2}"),
         # type of election of each estimate (the CIS does not estimate the Senate)
         cis_chamber = case_when(str_starts(election, "generales")   ~ "Congress",
                                 str_starts(election, "europeas")    ~ "European Parliament",
                                 str_starts(election, "autonomicas") ~ "Autonomical",
                                 str_starts(election, "municipales") ~ "Municipal"),
         place      = coalesce(province, circumscription) |>
                        str_to_upper() |>
                        stringi::stri_trans_general("Latin-ASCII"),
         cis_seats  = ceiling((min_seats + max_seats) / 2),
         # the file also carries rows that are not CIS estimates
         cis_source = case_when(str_starts(source, "CIS")     ~ "CIS constituency estimate",
                                str_starts(source, "electos") ~ "previous result, same constituency",
                                TRUE                          ~ "no CIS estimate before 2018 (0)")) |>
  add_count(date, cis_chamber, place) |>
  filter(n == 1) |>                  # two municipalities called Cieza: no way to tell them apart
  select(date, cis_chamber, place, cis_seats, cis_source)

# ------------------ 2. fallback: seats won in that same list in the previous election
previous_seats <- vox_candidates |>
  filter(is_titular %in% c(TRUE, NA)) |>
  mutate(date = str_extract(election, "[0-9]{4}_[0-9]{2}")) |>
  group_by(chamber, province, date) |>
  summarise(seats_won = sum(is_elected %in% TRUE), .groups = "drop") |>
  arrange(chamber, province, date) |>
  group_by(chamber, province) |>
  mutate(previous_seats = lag(seats_won)) |>          # NA = first time Vox ran this list
  ungroup() |>
  select(chamber, province, date, previous_seats)

# ----------------------------------------------- 3. expected seats into the lists
vox_candidates <- vox_candidates |>
  select(-any_of(c("expected_seats", "expected_seats_source", "top_expected"))) |>
  mutate(date        = str_extract(election, "[0-9]{4}_[0-9]{2}"),
         cis_chamber = case_when(chamber %in% c("Congress", "European Parliament") ~ chamber,
                                 level %in% c("Autonomical", "Municipal")          ~ level,
                                 TRUE                                              ~ "Senate"),
         place       = if_else(level == "Municipal", chamber, province) |>
                         str_to_upper() |>
                         stringi::stri_trans_general("Latin-ASCII"),
         place       = if_else(chamber == "European Parliament", "ESPANA", place),
         place       = coalesce(same_place[place], place)) |>
  left_join(cis_seats,      by = c("date", "cis_chamber", "place")) |>
  left_join(previous_seats, by = c("chamber", "province", "date")) |>
  mutate(expected_seats        = coalesce(cis_seats, previous_seats, 0),
         expected_seats_source = case_when(!is.na(cis_seats)      ~ cis_source,
                                           !is.na(previous_seats) ~ "previous result, same constituency",
                                           TRUE                   ~ "first contest (0)"),
         top_expected          = is_titular %in% c(TRUE, NA) & list_position <= expected_seats) |>
  select(-cis_chamber, -place, -cis_seats, -cis_source, -previous_seats)

# ---------------- 4. did the person win a seat in the election they ran in just before?
seat_by_date <- vox_candidates |>
  filter(!is.na(person_id), person_id != "") |>
  group_by(person_id, date) |>
  summarise(won_seat = any(is_elected %in% TRUE), .groups = "drop") |>
  arrange(person_id, date) |>
  group_by(person_id) |>
  mutate(had_seat_before = lag(won_seat, default = FALSE)) |>
  ungroup() |>
  select(person_id, date, had_seat_before)

vox_candidates <- vox_candidates |>
  select(-any_of("had_seat_before")) |>
  left_join(seat_by_date, by = c("person_id", "date")) |>
  mutate(had_seat_before = coalesce(had_seat_before, FALSE))
```

```{r ISCO VARIABLES JOIN}
# occupation before politics (ISCO-08) of the Congress candidates of July 2023 within
# the electoral threshold, and of the three titulars right below it in each constituency,
# who are neither heads of list nor above the threshold in that election
# (ISCO_CANDIDATOS_VOX_2023.xlsx, column GRUPO)
isco_candidates <- read_excel("ISCO_CANDIDATOS_VOX_2023.xlsx") |>
  select(person_id          = PERSON_ID,
         isco08_group       = GRUPO,
         isco08             = ISCO08,
         isco08_title       = ISCO08_TITLE,
         isco08_major       = ISCO08_MAJOR,
         isco08_major_title = ISCO08_MAJOR_TITLE)

vox_candidates <- vox_candidates |>
  select(-any_of(c("isco08_group", "isco08", "isco08_title", "isco08_major", "isco08_major_title"))) |>
  left_join(isco_candidates, by = "person_id")
```

```{r IN LISTS IN 2013-2017}
# people who already stood on a Vox list between 2013 and 2017, when Vox had
# almost no representation
early_candidates <- vox_candidates |>
  filter(between(year, 2013, 2017)) |>
  distinct(person_id)

vox_candidates <- vox_candidates |>
  mutate(in_lists_2013_2017 = person_id %in% early_candidates$person_id)
```


```{r PREV PARTY VARIABLE}
# ------------------------------------ 2. every candidacy for another party, with its year
other_party_candidacies <- bind_rows(
  pp_candidates |> mutate(prev_party = "PP") |> select(person_id, year, prev_party),
  cs_candidates |> mutate(prev_party = "Cs") |> select(person_id, year, prev_party),
  radical_right_candidates |>
    left_join(party_abbreviations, by = "party") |>
    select(person_id, year, prev_party)) |>
  filter(!is.na(person_id), person_id != "") |>
  mutate(year = as.integer(year),
         person_id = person_id |>
           # labels the bulletins print after the name ("IND" for independents, the
           # party, the section), which stop the same person from matching
           str_remove(" (CV BOCYLD |TITULARES )?NUM$") |>
           str_remove(" (IND|JN|AND|UPN|FN|MOVIMENT SOCIAL)$") |>
           # surnames split at the end of the line in the PDFs: "FERNAND EZ"
           str_replace("(?<=[A-Z]{3}) (EZ|Z)$", "\\1") |>
           str_squish()) |>
  distinct()

# ------------------------------------ 3. parties found in the press (PREV_PARTY_PRENSA_VOX.xlsx):
# candidates within the expected seats who had been in another party before joining Vox
press_prev_party <- read_excel("PREV_PARTY_PRENSA_VOX.xlsx") |>
  select(person_id = PERSON_ID, prev_party = PARTIDO_ANTERIOR) |>
  separate_longer_delim(prev_party, " | ")

# ------------------- 4. parties the person had been in BEFORE each Vox candidacy
vox_candidacy_years <- vox_candidates |>
  distinct(person_id, year)

prev_party_by_year <- bind_rows(
  # ran for another party in an earlier year
  vox_candidacy_years |>
    inner_join(other_party_candidacies, by = "person_id", suffix = c("", "_other"),
               relationship = "many-to-many") |>
    filter(year_other < year) |>
    select(person_id, year, prev_party),
  # member of another party before joining Vox, according to the press
  vox_candidacy_years |>
    inner_join(press_prev_party, by = "person_id", relationship = "many-to-many")) |>
  group_by(person_id, year) |>
  summarise(prev_party = str_c(sort(unique(prev_party)), collapse = " | "), .groups = "drop")

vox_candidates <- vox_candidates |>
  select(-any_of(c("prev_party", "has_prev_party"))) |>
  left_join(prev_party_by_year, by = c("person_id", "year")) |>
    mutate(has_prev_party = if_else(!is.na(prev_party), TRUE, NA))
```


```{r}
# electable_candidates <- 
  vox_candidates |>
  filter(top_expected | top_won) |> 
  select(chamber, year, province, list_position, candidate, is_elected, expected_seats, top_won, seats_won, prev_party) |> view()
```

## Electoral Results Vox

```{r TABLE VOX RESULTS}
# one row per election: general elections keep Congress and Senate apart,
# regional elections held on the same day are counted together
vox_results <- vox_candidates |>
  filter(is_titular %in% c(TRUE, NA)) |>
  mutate(contest = if_else(level == "National", str_c(election, " ", chamber), election)) |>
  group_by(level, date, contest) |>
  summarise(lists            = n_distinct(chamber, province),
            candidates       = n(),
            seats_won        = sum(is_elected %in% TRUE),
            lists_with_seats = n_distinct(chamber[is_elected %in% TRUE], province[is_elected %in% TRUE]),
            expected_seats   = sum(top_expected),
            .groups = "drop") |>
  arrange(factor(level, levels = c("National", "Autonomical", "Municipal")), date) |>
  select(Level = level, Election = contest, Lists = lists, Candidates = candidates,
         `Expected seats` = expected_seats, `Seats won` = seats_won, `Lists with seats` = lists_with_seats)

stargazer(as.data.frame(vox_results), summary = FALSE, rownames = FALSE, type = table_type,
          title = "Vox electoral results by election",
          notes = "Titular candidates only. Expected seats: candidates placed within the expected seats of their list.")
```

```{r FACETED CARTOGRAM MAP}
library(mapSpain)
library(sf)
library(readxl)
library(cartogram)

# Vox's vote and its access to regional government in the four phases of the party, on
# a contiguous cartogram (r-graph-gallery.com/331-basic-cartogram). Each region is
# resized by its area times Vox's vote share there: the total size of Spain does not
# change, and a region grows where Vox did better than its own average and shrinks where
# it did worse. The size compares Vox with itself, not with the other parties
phases <- tribble(
  ~phase,       ~phase_start,       ~phase_end,
  "2014-2017",  ymd("2014-01-01"),  ymd("2017-12-31"),
  "2018-2019",  ymd("2018-01-01"),  ymd("2019-12-31"),
  "2020-2022",  ymd("2020-01-01"),  ymd("2022-12-31"),
  "2023-2026",  ymd("2023-01-01"),  ymd("2026-12-31")
)

# ---- 1. Vox's vote share by region in the general elections of each phase -------------
# Congress results by constituency (Ministry of the Interior, infoelectoral, PROV_02 files).
# There was no general election in 2020-2022: that phase keeps the result of November
# 2019, the one in force during it
general_by_phase <- tribble(
  ~file,                   ~phase,       ~election,
  "PROV_02_201512_1.zip",  "2014-2017",  "December 2015",
  "PROV_02_201606_1.zip",  "2014-2017",  "June 2016",
  "PROV_02_201904_1.zip",  "2018-2019",  "April 2019",
  "PROV_02_201911_1.zip",  "2018-2019",  "November 2019",
  "PROV_02_201911_1.zip",  "2020-2022",  "November 2019",
  "PROV_02_202307_1.zip",  "2023-2026",  "July 2023"
)

# the party names are in row 2 or 3 of each file and the column labels in row 4; the
# provinces start in row 5. Where Vox did not run, its votes are empty (0 votes)
general_results <- tibble(file = unique(general_by_phase$file)) |>
  mutate(results = map(file, \(f) {
    raw   <- read_excel(unzip(f, exdir = tempdir()), col_names = FALSE, col_types = "text")
    vox   <- which(unlist(raw[2, ]) == "VOX" | unlist(raw[3, ]) == "VOX")[1]
    valid <- which(unlist(raw[4, ]) == "Votos válidos")
    rows  <- which(str_detect(raw[[2]], "^[0-9]+$"))
    tibble(cpro        = str_pad(raw[[2]][rows], 2, pad = "0"),
           vox_votes   = replace_na(as.numeric(raw[[vox]][rows]), 0),
           valid_votes = as.numeric(raw[[valid]][rows]))
  })) |>
  unnest(results)

# 52 constituencies in every file (must return 0 rows)
general_results |> count(file) |> filter(n != 52)

# provinces added up to regions, and the elections of each phase averaged
vote_by_phase <- general_by_phase |>
  left_join(general_results, by = "file", relationship = "many-to-many") |>
  left_join(esp_codelist |> distinct(cpro, codauto), by = "cpro") |>
  summarise(vote_share = 100 * sum(vox_votes) / sum(valid_votes), .by = c(phase, election, codauto)) |>
  summarise(vote_share = mean(vote_share), .by = c(phase, codauto))

# ---- 2. Vox in the regional government, or propping up a PP government --------------
# a Vox post counts in a phase if it started before the phase ended and ended after it
# started (posts still held have no end date)
vox_in_government <- vox_executive_positions |>
  mutate(start = dmy(start_date), end = coalesce(dmy(end_date), today())) |>
  left_join(tribble(~region,                 ~codauto,
                    "Andalucía",             "01",
                    "Aragón",                "02",
                    "Castilla y León",       "07",
                    "Comunidad Valenciana",  "10",
                    "Extremadura",           "11",
                    "Región de Murcia",      "14"), by = "region") |>
  cross_join(phases) |>
  filter(start <= phase_end, end >= phase_start) |>
  distinct(codauto, phase)

# PP governments that owed their investiture to the votes of Vox without Vox entering
# them, in office at some point of the phase
pp_needs_vox <- tribble(
  ~codauto, ~phase,       ~investiture,
  "01",     "2018-2019",  "Juan Manuel Moreno, January 2019 (PP-Cs)",
  "13",     "2018-2019",  "Isabel Díaz Ayuso, August 2019 (PP-Cs)",
  "14",     "2018-2019",  "Fernando López Miras, July 2019 (PP-Cs)",
  "01",     "2020-2022",  "Juan Manuel Moreno, in office until July 2022",
  "13",     "2020-2022",  "Isabel Díaz Ayuso, in office until 2021 and invested again with Vox votes in June 2021",
  "14",     "2020-2022",  "Fernando López Miras, in office until 2023",
  "04",     "2023-2026",  "Marga Prohens, July 2023"
)

# ---- 3. one contiguous cartogram per phase ----------------------------------------------
# equal-area projection (the cartogram works with areas); the Canary Islands are moved
# next to the peninsula before the calculation
regions <- esp_get_ccaa(epsg = "3035") |>
  select(codauto) |>
  mutate(area = as.numeric(st_area(geometry)))

# weight = area x vote share relative to Vox's national share in the phase. A region
# where Vox did not run keeps 5% of its size (a weight of zero cannot be drawn)
vox_cartogram <- phases$phase |>
  map(\(p) regions |>
        left_join(filter(vote_by_phase, phase == p), by = "codauto") |>
        mutate(phase    = p,
               relative = vote_share / weighted.mean(vote_share, area),
               weight   = area * pmax(relative, 0.05)) |>
        cartogram_cont("weight", itermax = 15)) |>
  bind_rows() |>
  mutate(status = case_when(
           str_c(codauto, phase) %in% str_c(vox_in_government$codauto, vox_in_government$phase) ~ "Vox in the government",
           str_c(codauto, phase) %in% str_c(pp_needs_vox$codauto, pp_needs_vox$phase)           ~ "PP government that needed Vox's votes",
           TRUE                                                                                  ~ "Neither") |>
           factor(levels = c("Vox in the government", "PP government that needed Vox's votes", "Neither")))

# ---- 4. the map ---------------------------------------------------------------------------
# national vote share in each phase, for the panel titles
phase_names <- general_by_phase |>
  left_join(general_results, by = "file", relationship = "many-to-many") |>
  summarise(vote_share = 100 * sum(vox_votes) / sum(valid_votes), .by = c(phase, election)) |>
  summarise(label = str_c(first(phase), "\n", str_c(election, collapse = " and "), ": ",
                          str_c(format(round(vote_share, 1), nsmall = 1), "%", collapse = " and ")),
            .by = phase) |>
  deframe()

ggplot(vox_cartogram) +
  geom_sf(aes(fill = status), colour = "white", linewidth = 0.2) +
  scale_fill_manual(values = c("Vox in the government"                 = "#1b7a3d",
                               "PP government that needed Vox's votes" = "#9fd3ad",
                               "Neither"                               = "grey70"),
                    drop = FALSE, name = NULL) +
  facet_wrap(~ phase, ncol = 2, labeller = as_labeller(phase_names)) +
  labs(title = "Vox's vote and its access to regional government",
       subtitle = "Regions resized by Vox's vote there: larger where Vox did better than its own average",
       caption = str_c("Size: area of the region times Vox's share of the valid vote in the general elections of the phase,\n",
                       "relative to its national share (a region where Vox did not run keeps 5% of its size).\n",
                       "Colour: Vox held a post in the regional government, or its votes were needed to invest a PP president,\n",
                       "at any point of the phase. Source: Ministry of the Interior (infoelectoral). Canary Islands moved next to the peninsula.")) +
  theme_void() +
  theme(text = element_text(colour = "#22211d"),
        plot.background = element_rect(fill = "#f5f5f4", colour = NA),
        panel.background = element_rect(fill = "#f5f5f4", colour = NA),
        legend.background = element_rect(fill = "#f5f5f4", colour = NA),
        plot.title = element_text(size = 16, hjust = 0.5, colour = "#4e4d47"),
        plot.subtitle = element_text(size = 9, hjust = 0.5, colour = "#4e4d47", margin = margin(b = 8)),
        plot.caption = element_text(size = 7, hjust = 0, colour = "grey40"),
        strip.text = element_text(size = 9, margin = margin(b = 4)),
        legend.position = "bottom",
        legend.text = element_text(size = 8),
        plot.margin = margin(10, 10, 10, 10))
```

```{r MAP COUNCILLERS PRE 2018}
library(mapSpain)
library(sf)

# Vox councillors elected in the 2015 municipal elections (Vox's only local seats before 2018)
councillors_2015 <- vox_candidates |>
  filter(level == "Municipal", year == 2015, is_elected %in% TRUE) |>
  count(province, chamber, name = "councillors") |>
  mutate(muni_key = chamber |>
           str_to_upper() |>
           stringi::stri_trans_general("Latin-ASCII") |>
           str_split("/") |>
           map_chr(\(x) str_c(sort(str_squish(x)), collapse = "/")))

municipalities_map <- esp_get_munic(epsg = "3857") |>
  left_join(provinces, by = c("cpro" = "province_code")) |>
  mutate(muni_key = name |>
           str_to_upper() |>
           stringi::stri_trans_general("Latin-ASCII") |>
           str_split("/") |>
           map_chr(\(x) str_c(sort(str_squish(x)), collapse = "/")))

councillors_map <- municipalities_map |>
  inner_join(councillors_2015, by = c("muni_key", "province"))

# councillors whose municipality was not found on the map (must return 0 rows)
councillors_2015 |> anti_join(st_drop_geometry(councillors_map), by = c("muni_key", "province"))

# one point per municipality, at its centroid
councillors_centroids <- st_centroid(councillors_map)

councillors_points <- councillors_centroids |>
  st_drop_geometry() |>
  bind_cols(as_tibble(st_coordinates(councillors_centroids)))

ggplot() +
  geom_sf(data = esp_get_prov(epsg = "3857"), fill = "grey95", colour = "grey70", linewidth = 0.2) +
  geom_sf(data = esp_get_can_box(epsg = "3857"), colour = "grey60") +
  geom_point(data = councillors_points, aes(X, Y, size = councillors), colour = "#009E73") +
  ggrepel::geom_text_repel(data = councillors_points, aes(X, Y, label = str_c(name, " (", councillors, ")")),
                           size = 3, min.segment.length = 0) +
  scale_size_continuous(range = c(2, 5), breaks = 1:3, name = "Councillors") +
  labs(title = "Vox councillors before 2018",
       subtitle = "Municipal elections of May 2015") +
  theme_void() +
  theme(legend.position = "bottom")
```


```{r TABLE VOX CANDIDATES BY LEVELS}
# profile of the candidates at each level (one row per candidacy)
candidates_by_level <- vox_candidates |>
  mutate(level = factor(level, levels = c("National", "Autonomical", "Municipal"))) |>
  group_by(level) |>
  summarise(`Candidacies`                        = n(),
            `Different people`                   = n_distinct(person_id),
            `Elections`                          = n_distinct(election),
            `Women (%)`                          = round(100 * mean(sex[!is.na(sex)] == "Female"), 1),
            `Elected (%)`                        = round(100 * mean(is_elected %in% TRUE), 1),
            `Within expected seats (%)`          = round(100 * mean(top_expected), 1),
            `Seat in previous election (%)`      = round(100 * mean(had_seat_before), 1),
            # has_prev_party is TRUE or NA: the NA count as no party on record
            `Ran for another party before (%)`   = round(100 * mean(has_prev_party %in% TRUE), 1),
            `In the CEN before the election (%)` = round(100 * mean(in_cen_before), 2),
            `Organic position in Vox (%)`        = round(100 * mean(has_organic_position), 2),
            .groups = "drop") |>
  mutate(level = as.character(level)) |>
  pivot_longer(-level, names_to = "Variable") |>
  pivot_wider(names_from = level, values_from = value)

stargazer(as.data.frame(candidates_by_level), summary = FALSE, rownames = FALSE, type = table_type,
          title = "Vox candidates by level of election",
          notes = "One row per candidacy. Sex of municipal candidates inferred from the forename (INE census).")

# the same variables for the candidates within the expected seats and the rest
candidates_by_expected <- vox_candidates |>
  mutate(group = if_else(top_expected, "Within expected seats", "Rest of the list")) |>
  group_by(group) |>
  summarise(`Candidacies`                        = n(),
            `Women (%)`                          = round(100 * mean(sex[!is.na(sex)] == "Female"), 1),
            `Elected (%)`                        = round(100 * mean(is_elected %in% TRUE), 1),
            `Seat in previous election (%)`      = round(100 * mean(had_seat_before), 1),
            `Ran for another party before (%)`   = round(100 * mean(has_prev_party %in% TRUE), 1),
            `In the CEN before the election (%)` = round(100 * mean(in_cen_before), 2),
            `Mean list position`                 = round(mean(list_position, na.rm = TRUE), 1),
            .groups = "drop") |>
  pivot_longer(-group, names_to = "Variable") |>
  pivot_wider(names_from = group, values_from = value)

stargazer(as.data.frame(candidates_by_expected), summary = FALSE, rownames = FALSE, type = table_type,
          title = "Candidates within and outside the expected seats")
```


```{r MAP VOX GOVERNMENT}
# regions where Vox has held regional government posts, by coalition period
government_by_region <- vox_executive_positions |>
  mutate(period = if_else(str_ends(start_date, "2026"), "2026 onwards", "2022-2024")) |>
  group_by(region, period) |>
  summarise(posts = n_distinct(position), .groups = "drop") |>   # Limia and Higuero held the same post
  mutate(codauto = case_match(region,
                              "Andalucía"            ~ "01",
                              "Aragón"               ~ "02",
                              "Castilla y León"      ~ "07",
                              "Comunidad Valenciana" ~ "10",
                              "Extremadura"          ~ "11",
                              "Región de Murcia"     ~ "14"))

government_map <- esp_get_ccaa(epsg = "3857") |>
  cross_join(tibble(period = c("2022-2024", "2026 onwards"))) |>
  left_join(government_by_region, by = c("codauto", "period"))

ggplot(government_map) +
  geom_sf(aes(fill = posts), colour = "white", linewidth = 0.2) +
  geom_sf(data = esp_get_can_box(epsg = "3857"), colour = "grey60") +
  scale_fill_gradient(low = "#b7e4c7", high = "#1b7a3d", na.value = "grey90",
                      breaks = 1:4, name = "Vox posts in\nthe regional government") +
  facet_wrap(~ period) +
  labs(title = "Vox in regional governments") +
  theme_void() +
  theme(legend.position = "bottom")
```


```{r TABLE VOX IN REGIONAL GOVERNMENTS}
# every government post, and whether the person had been a Vox candidate in that election
executive_table <- vox_executive_positions |>
  left_join(vox_candidates |> distinct(name = person_id, election) |> mutate(candidate = "Yes"),
            by = c("name", "election")) |>
  mutate(candidate = replace_na(candidate, "No"),
         end_date  = replace_na(end_date, "in office")) |>
  select(Region = region, Name = name, Position = position, From = start_date, To = end_date,
         `Vox candidate in that election` = candidate)

stargazer(as.data.frame(executive_table), summary = FALSE, rownames = FALSE, type = table_type,
          title = "Vox members of regional governments")
```


```{r TABLE PREVIOUS PARTY OF VOX CANDIDATES}
# candidacies by the parties the person had run for before (a candidacy can count for several)
previous_party_table <- vox_candidates |>
  filter(has_prev_party) |>
  separate_longer_delim(prev_party, " | ") |>
  mutate(level = factor(level, levels = c("National", "Autonomical", "Municipal"))) |>
  count(prev_party, level) |>
  pivot_wider(names_from = level, values_from = n, values_fill = 0) |>
  mutate(Total = National + Autonomical + Municipal) |>
  arrange(desc(Total)) |>
  rename(`Previous party` = prev_party)

stargazer(as.data.frame(previous_party_table), summary = FALSE, rownames = FALSE, type = table_type,
          title = "Vox candidacies by previous party",
          notes = "Parties the person had run for in an earlier year. A candidacy can count for several parties.")
```


```{r TABLE MULTI-LEVEL CAREERS}
# how many people ran at one, two or three levels, and how often they were elected
careers <- vox_candidates |>
  group_by(person_id) |>
  summarise(levels     = str_c(sort(unique(level)), collapse = " + "),
            n_levels   = n_distinct(level),
            elections  = n_distinct(election),
            ever_elected = any(is_elected %in% TRUE),
            .groups = "drop") |>
  group_by(`Levels` = levels) |>
  summarise(People                  = n(),
            `Mean elections`        = round(mean(elections), 2),
            `Ever elected (%)`      = round(100 * mean(ever_elected), 1),
            .groups = "drop") |>
  arrange(desc(People))

stargazer(as.data.frame(careers), summary = FALSE, rownames = FALSE, type = table_type,
          title = "Vox candidates by the levels of election they ran in")
```


```{r TERRITORIAL PENETRATION IN GENERAL ELECTIONS}
library(mapSpain)
library(sf)

# ------------------------------------ constituencies where Vox ran a Congress list
vox_general <- vox_candidates |>
  filter(chamber == "Congress") |>
  distinct(election, province) |>
  left_join(provinces, by = "province") |>
  mutate(presented = "Yes")

# constituencies that could not be placed in a province (must return 0 rows)
vox_general |> filter(is.na(province_code))

general_elections <- vox_general |>
  distinct(election) |>
  mutate(election_label = str_replace(str_extract(election, "[0-9]{4}_[0-9]{2}"), "_", "-"))

# ------------------------------------------ one map per election, all provinces
vox_presence_map <- esp_get_prov(epsg = "3857") |>
  cross_join(general_elections) |>
  left_join(vox_general |> select(election, cpro = province_code, presented),
            by = c("election", "cpro")) |>
  mutate(presented = replace_na(presented, "No"))

vox_presence_general <- ggplot(vox_presence_map) +
  geom_sf(aes(fill = presented), colour = "white", linewidth = 0.15) +
  geom_sf(data = esp_get_can_box(epsg = "3857"), fill = NA, colour = "grey60") +
  scale_fill_manual(values = c("Yes" = "#009E73", "No" = "grey90"),
                    name = "Vox list") +
  facet_wrap(~ election_label) +
  labs(title = "Provinces where Vox ran in each general election") +
  theme_void() +
  theme(legend.position = "bottom")

vox_presence_general

```


```{r TERRITORIAL PENETRATION IN AUTONOMICAL ELECTIONS}
library(mapSpain)
library(sf)

# regional constituencies that are not provinces: islands, districts, regional lists
constituency_to_province <- tribble(
  ~province_key,                ~province_code,
  "REGION DE MURCIA",           "30",
  "PRIMERA",                    "30",
  "SEGUNDA",                    "30",
  "TERCERA",                    "30",
  "CUARTA",                     "30",
  "QUINTA",                     "30",
  "CENTRO",                     "33",
  "OCCIDENTE",                  "33",
  "ORIENTE",                    "33",
  "MALLORCA",                   "07",
  "MENORCA",                    "07",
  "EIVISSA",                    "07",
  "FORMENTERA",                 "07",
  "GRAN CANARIA",               "35",
  "LANZAROTE",                  "35",
  "FUERTEVENTURA",              "35",
  "TENERIFE",                   "38",
  "LA PALMA",                   "38",
  "LA GOMERA",                  "38",
  "EL HIERRO",                  "38",
  "CIRCUNSCRIPCION AUTONOMICA", "35",
  "CIRCUNSCRIPCION AUTONOMICA", "38",
  "CIUDAD",                     "13",   # "Ciudad Real" cut in the source
  "ALAVA",                      "01"
) |>
  bind_rows(provinces |>
              mutate(province_key = province |> str_to_upper() |> stringi::stri_trans_general("Latin-ASCII")) |>
              select(province_key, province_code))

# ------------------------------- provinces that voted in each regional election
regional_lists <- autonomical_candidates |>
  filter(year >= 2014) |>
  mutate(province_key = province |> str_to_upper() |> stringi::stri_trans_general("Latin-ASCII") |> str_squish()) |>
  left_join(constituency_to_province, by = "province_key", relationship = "many-to-many")

# constituencies that could not be placed in a province (must return 0 rows)
regional_lists |> filter(is.na(province_code)) |> distinct(chamber, province)

regional_elections <- regional_lists |>
  distinct(election, cpro = province_code) |>
  left_join(esp_codelist |> distinct(cpro, ccaa = ccaa.shortname.es), by = "cpro") |>
  group_by(election) |>
  mutate(election_label = str_c(str_replace(str_extract(election, "[0-9]{4}_[0-9]{2}"), "_", "-"), "\n",
                                if_else(n_distinct(ccaa) == 1, first(ccaa), str_c(n_distinct(ccaa), " regions")))) |>
  ungroup()

vox_regional <- regional_lists |>
  filter(abbreviation == "VOX") |>
  distinct(election, cpro = province_code) |>
  mutate(presented = "Yes")

# ------------------------------------------ one map per election, all provinces
vox_presence_regional_map <- esp_get_prov(epsg = "3857") |>
  cross_join(regional_elections |> distinct(election, election_label)) |>
  left_join(regional_elections |> distinct(election, cpro) |> mutate(voted = TRUE), by = c("election", "cpro")) |>
  left_join(vox_regional, by = c("election", "cpro")) |>
  mutate(presented = case_when(!is.na(presented) ~ "Yes",
                               voted             ~ "No",
                               TRUE              ~ "No election"),
         presented = factor(presented, levels = c("Yes", "No", "No election"))) |>
  # keep only elections where Vox skipped at least one province that voted
  group_by(election) |>
  filter(any(presented == "No")) |>
  ungroup()

vox_presence_regional <- ggplot(vox_presence_regional_map) +
  geom_sf(aes(fill = presented), colour = "grey75", linewidth = 0.1) +
  geom_sf(data = esp_get_can_box(epsg = "3857"), fill = NA, colour = "grey60") +
  scale_fill_manual(values = c("Yes" = "#009E73", "No" = "grey65", "No election" = "white"),
                    name = "Vox list") +
  facet_wrap(~ election_label, ncol = 2) +
  labs(title = "Regional elections where Vox did not run everywhere") +
  theme_void() +
  theme(legend.position = "bottom",
        strip.text = element_text(size = 8))

vox_presence_regional
```



## Social Network

```{r EDGELIST CREATION}
library(tidyverse)

network_candidates <- vox_candidates |>
  mutate(
    contest = case_when(
      level == "National" ~ election,
      level == "Autonomical" ~ str_c(election, chamber, sep = " | "),
      level == "Municipal" ~ str_c(
        election,
        coalesce(province, ""),
        chamber,
        sep = " | "
      ),
      TRUE ~ NA_character_
    ),
    # in municipal elections there are no CIS estimates, so the ex ante rule (previous
    # result in the same municipality) misses the local breakthrough of 2019: the head
    # of a list that won at least one councillor is also electable
    electable = coalesce(top_expected, FALSE) | coalesce(top_won, FALSE) |
      (level == "Municipal" & list_position == 1 & is_titular %in% c(TRUE, NA) & is_elected %in% TRUE)
  ) |>
  filter(!is.na(contest), !is.na(person_id))

node_ids <- network_candidates |>
  filter(electable) |>
  distinct(person_id)

# ---- members of Vox's satellite organisations ---------------------------------------------
# every member is a node, whether or not they stood for Vox. The ones who did are matched
# with their candidate name, which is often longer ("IVAN VELEZ" is "ISMAEL IVAN VELEZ
# CIPRIANO"): the words of the member's name must appear in the candidate's name in the
# same order, with a middle name allowed after the first one
satellites <- c("DENAES", "Disenso", "ISSEP", "HazteOir", "CitizenGO", "Solidaridad", "Revuelta", "Homo Legens")

satellite_members <- people |>
  filter(organisation %in% satellites, str_count(person, " ") >= 1) |>
  distinct(organisation, person)

candidate_words <- network_candidates |>
  distinct(person_id) |>
  mutate(word = person_id) |>
  separate_longer_delim(word, " ") |>
  distinct()

# every candidate each member's name fits
satellite_fits <- satellite_members |>
  distinct(person) |>
  # particles are dropped from both names ("IGNACIO DE HOCES INIGUEZ" is "IGNACIO HOCES INIGUEZ")
  mutate(core = str_squish(str_remove_all(person, "\\b(DE|DEL|LA|LAS|LOS|Y)\\b")),
         word = word(core, -1)) |>
  inner_join(candidate_words, by = "word", relationship = "many-to-many") |>
  mutate(candidate_core = str_squish(str_remove_all(person_id, "\\b(DE|DEL|LA|LAS|LOS|Y)\\b")),
         pattern = str_c("^(.*?)\\b", word(core, 1), "\\b(.*?)\\b", word(core, 2, -1), "\\b"),
         before  = str_match(candidate_core, pattern)[, 2],
         between = str_match(candidate_core, pattern)[, 3]) |>
  # the words before the member's first name and between it and the surname must be
  # forenames of the INE census ("ISMAEL IVAN VELEZ", "ANA MARIA VELASCO"), never surnames:
  # "SUSANA DIAZ" is not "SUSANA GARCIA DIAZ". A member named with more than one surname
  # is specific enough to skip the test ("HERMANN TERTSCH VALLELERSUNDI" is "HERMANN
  # LEOPOLD TERTSCH VALLELERSUNDI", and Leopold is not in the census)
  filter(!is.na(before),
         str_count(core, " ") >= 2 |
           map_lgl(str_split(str_squish(str_c(before, " ", between)), " "),
                   \(w) all(w %in% c("", census_forenames$forename)))) |>
  distinct(person, person_id) |>
  mutate(electable = person_id %in% node_ids$person_id)

# a name that fits several candidates ("RICARDO CHAMORRO": Delmo and Arevalo) goes to the
# one who was electable, if only one was; otherwise it is left out
satellite_matches <- satellite_fits |>
  filter(n() == 1 | (electable & sum(electable) == 1), .by = person)

# members matched to a candidate under a longer name: check them
satellite_matches |> filter(person != person_id) |> print(n = Inf)

# names that fit several candidates and were left out: check them by hand
satellite_fits |> filter(!person %in% satellite_matches$person) |> print(n = Inf)

# members who never stood for Vox, or did so under a name that could not be matched:
# look for the second kind here
satellite_members |> filter(!person %in% satellite_fits$person) |> arrange(organisation, person) |> print(n = Inf)

# each member with the id of their node: the candidate's if matched, their own otherwise
satellite_nodes <- satellite_members |>
  left_join(satellite_matches |> select(person, person_id), by = "person") |>
  mutate(person_id = coalesce(person_id, person))

# the members who stood for Vox join the candidacy network even if never electable
node_ids <- node_ids |>
  bind_rows(satellite_nodes |> filter(person_id %in% network_candidates$person_id) |> distinct(person_id)) |>
  distinct()

# ---- families of the previous parties, for the colours of the nodes ------------------
# the far right split as in the text (Casals): nostalgic Francoist and Falangist right,
# ultra-Catholic and traditionalist parties, national-populist right, neo-Nazi fringe
# and personalist vehicles; the mainstream parties apart
party_families <- tribble(
  ~party,     ~family,
  "PP",       "PP",
  "AP",       "PP",
  "Cs",       "Cs",
  "FN",       "Francoist / Falangist",
  "JJEE",     "Francoist / Falangist",
  "FE-JONS",  "Francoist / Falangist",
  "FA",       "Francoist / Falangist",
  "FE-2000",  "Francoist / Falangist",
  "LF",       "Francoist / Falangist",
  "ADÑ",      "Francoist / Falangist",
  "AES",      "Ultra-Catholic / traditionalist",
  "CTC",      "Ultra-Catholic / traditionalist",
  "DN",       "National-populist",
  "E2000",    "National-populist",
  "PxC",      "National-populist",
  "VD",       "National-populist",
  "SOMI",     "National-populist",
  "PxL",      "National-populist",
  "NOSOTROS", "National-populist",
  "HN",       "National-populist",
  "ADN",      "National-populist",
  "MSR",      "Neo-Nazi",
  "AN",       "Neo-Nazi",
  "GIL",      "Personalist populist",
  "ARM",      "Personalist populist",
  "CDS",      "Other right and centre",
  "PADE",     "Other right and centre",
  "UPN",      "Other right and centre",
  "Foro",     "Other right and centre",
  "PDP",      "Other right and centre",
  "AG",       "Other right and centre",
  "SCD",      "Other right and centre",
  "ACTUA",    "Other right and centre",   # Actúa Baleares
  "SR",       "Other right and centre",   # Somos Región (Murcia)
  "CiU",      "Other right and centre",
  "PSOE",     "Left",
  "PSC",      "Left",
  "PCE-EPK",  "Left"
)

# a candidate with several families keeps the first one in this order: a far-right past
# is kept over a mainstream one, being the rarer and more telling of the two
family_order <- c("Neo-Nazi", "Francoist / Falangist", "Ultra-Catholic / traditionalist",
                  "National-populist", "Personalist populist", "PP", "Cs",
                  "Other right and centre", "Left", "None")

# the palette to set in Gephi (Appearance > Nodes > Partition > prev_party_family)
family_colours <- c("Neo-Nazi"                        = "#000000",
                    "Francoist / Falangist"           = "#D55E00",
                    "Ultra-Catholic / traditionalist" = "#6A3D9A",
                    "National-populist"               = "#009E73",
                    "Personalist populist"            = "#F0E442",
                    "PP"                              = "#0072B2",
                    "Cs"                              = "#E69F00",
                    "Other right and centre"          = "#56B4E9",
                    "Left"                            = "#CC79A7",
                    "None"                            = "#D9D9D9")

# a member of a satellite organisation is coloured by the organisation instead
satellite_colours <- c("DENAES"      = "#8C510A",
                       "Disenso"     = "#DFC27D",
                       "ISSEP"       = "#E7298A",
                       "HazteOir"    = "#66A61E",
                       "CitizenGO"   = "#1B7837",
                       "Solidaridad" = "#A50026",
                       "Revuelta"    = "#FB9A99",
                       "Homo Legens" = "#01665E")

# parties without a family (must return 0 rows)
vox_candidates |>
  filter(!is.na(prev_party)) |>
  separate_longer_delim(prev_party, " | ") |>
  distinct(prev_party) |>
  filter(!prev_party %in% party_families$party)

vox_nodes <- network_candidates |>
  semi_join(node_ids, by = "person_id") |>
  group_by(person_id) |>
  summarise(
    name = first(candidate),
    contests = n_distinct(contest),
    contests_electable = n_distinct(contest[electable]),
    has_prev_party = any(coalesce(has_prev_party, FALSE)),
    prev_party = {
      x <- prev_party[
        coalesce(has_prev_party, FALSE) &
          !is.na(prev_party)
      ]
      x <- x |>
        str_split(fixed(" | ")) |>
        unlist() |>
        str_squish() |>
        unique() |>
        sort()
      if (length(x) == 0) NA_character_ else str_c(x, collapse = " | ")
    },
    .groups = "drop"
  ) |>
  mutate(
    n_prev_parties = case_when(
      is.na(prev_party) ~ 0L,
      TRUE ~ str_count(prev_party, fixed(" | ")) + 1L
    ),
    # every family of the person's previous parties, and the one that colours the node
    prev_party_families = map_chr(str_split(replace_na(prev_party, ""), fixed(" | ")), \(p) {
      families <- party_families$family[party_families$party %in% p]
      if (length(families) == 0) "None" else str_c(intersect(family_order, families), collapse = " | ")
    }),
    prev_party_family = word(prev_party_families, 1, sep = fixed(" | "))
  ) |>
  # the members who never stood for Vox, as nodes of their own
  bind_rows(satellite_nodes |>
              filter(!person_id %in% network_candidates$person_id) |>
              distinct(person_id) |>
              mutate(name = str_to_title(person_id), contests = 0L, contests_electable = 0L,
                     has_prev_party = FALSE, n_prev_parties = 0L,
                     prev_party_families = "None", prev_party_family = "None")) |>
  # the satellite organisations of each node (in the order of `satellites`)
  left_join(satellite_nodes |>
              summarise(organisations = str_c(intersect(satellites, organisation), collapse = " | "),
                        .by = person_id),
            by = "person_id") |>
  mutate(satellite = !is.na(organisations),
         candidate = contests > 0,
         # the attribute that colours the node: the organisation for its members, the
         # family of the previous party for everybody else
         family    = if_else(satellite, word(organisations, 1, sep = fixed(" | ")), prev_party_family),
         colour    = unname(c(family_colours, satellite_colours)[family]))

participants <- network_candidates |>
  semi_join(node_ids, by = "person_id") |>
  group_by(contest, person_id) |>
  summarise(
    electable = any(electable),
    .groups = "drop"
  )

vox_edges <- participants |>
  inner_join(
    participants,
    by = "contest",
    suffix = c("_from", "_to"),
    relationship = "many-to-many"
  ) |>
  filter(
    person_id_from < person_id_to,
    electable_from | electable_to
  ) |>
  count(
    from = person_id_from,
    to = person_id_to,
    name = "weight"
  )

# members of the same satellite organisation are linked as well; a pair can be linked by
# both kinds of tie, recorded in `tie`
organisation_edges <- satellite_nodes |>
  distinct(organisation, person_id) |>
  inner_join(satellite_nodes |> distinct(organisation, person_id), by = "organisation",
             suffix = c("_from", "_to"), relationship = "many-to-many") |>
  filter(person_id_from < person_id_to) |>
  count(from = person_id_from, to = person_id_to, name = "weight")

vox_edges <- bind_rows(vox_edges |> mutate(tie = "candidacy"),
                       organisation_edges |> mutate(tie = "organisation")) |>
  summarise(weight = sum(weight), tie = str_c(sort(unique(tie)), collapse = " + "), .by = c(from, to))

vox_nodes |> count(family, colour, sort = TRUE)
vox_edges |> count(tie)

vox_nodes_gephi <- vox_nodes |>
  mutate(Id = as.character(person_id), Label = as.character(name)) |>
  select(Id, Label, family, prev_party, prev_party_family, prev_party_families, n_prev_parties,
         has_prev_party, organisations, satellite, candidate, contests, contests_electable, colour)

vox_edges_gephi <- vox_edges |>
  mutate(Source = as.character(from), Target = as.character(to),
         Type = "Undirected", Weight = as.numeric(weight)) |>
  select(Source, Target, Type, Weight, tie)

write_csv(vox_nodes_gephi, "vox_nodes_gephi.csv")
write_csv(vox_edges_gephi, "vox_edges_gephi.csv")
```


```{r GEPHI}

```

```{r SOCIAL NETWORK ANALYSIS}

```


## Regressions

```{r EXPECTED SEATS VARIABLE AS A PREDICTOR}
library(sandwich)
library(lmtest)
library(broom)

# ---- 1. one row per VOX list (election + chamber + constituency) ----------------------
expected_check <- vox_candidates |>
  filter(is_titular %in% c(TRUE, NA)) |>
  group_by(level, election, chamber, province, expected_seats, expected_seats_source) |>
  summarise(seats = sum(is_elected %in% TRUE), .groups = "drop") |>
  mutate(from_cis = expected_seats_source == "CIS constituency estimate")

expected_check |> count(level, expected_seats_source)

# ---- 2. do the expected seats predict the seats actually won? --------------------------
# standard errors clustered by election: constituencies of the same election share shocks
models <- list(
  "1 Pooled"      = lm(seats ~ expected_seats, data = expected_check),
  "2 + level FE"  = lm(seats ~ expected_seats + level, data = expected_check),
  "3 + source"    = lm(seats ~ expected_seats * from_cis + level, data = expected_check),
  "4 National"    = lm(seats ~ expected_seats, data = filter(expected_check, level == "National")),
  "5 Autonomical" = lm(seats ~ expected_seats, data = filter(expected_check, level == "Autonomical")),
  "6 Municipal"   = lm(seats ~ expected_seats, data = filter(expected_check, level == "Municipal"))
)

regression_table <- list()
for (m in names(models)) {
  regression_table[[m]] <- coeftest(models[[m]],
                                    vcov = vcovCL(models[[m]], cluster = ~ election)) |>
    tidy() |>
    mutate(model = m,
           n     = nobs(models[[m]]),
           r2    = summary(models[[m]])$r.squared)
}
regression_table <- bind_rows(regression_table) |>
  select(model, term, estimate, std.error, p.value, n, r2) |>
  mutate(across(c(estimate, std.error, r2), \(x) round(x, 3)),
         p.value = signif(p.value, 3))

regression_table |> print(n = Inf)

# ---- 3. the threshold: are the top_expected candidates the ones who got elected? --------
electable_check <- vox_candidates |>
  filter(is_titular %in% c(TRUE, NA)) |>
  mutate(is_elected = is_elected %in% TRUE)

electable_logit <- glm(is_elected ~ top_expected + level,
                       family = binomial, data = electable_check)

coeftest(electable_logit, vcov = vcovCL(electable_logit, cluster = ~ election)) |>
  tidy() |>
  mutate(odds_ratio = exp(estimate))

# share elected among top_expected and the rest
electable_check |>
  group_by(level, top_expected) |>
  summarise(candidates = n(), elected = sum(is_elected),
            share_elected = mean(is_elected), .groups = "drop")

# ---- 4. plot: expected seats against seats won -------------------------------------------
ggplot(expected_check, aes(expected_seats, seats)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey60") +
  geom_jitter(aes(colour = expected_seats_source), width = 0.1, height = 0.1, alpha = 0.5, size = 1.6) +
  geom_smooth(method = "lm", se = TRUE, colour = "black", linewidth = 0.7) +
  facet_wrap(~ level, scales = "free") +
  scale_colour_manual(values = c("CIS constituency estimate"          = "#0072B2",
                                 "previous result, same constituency" = "#E69F00",
                                 "no CIS estimate before 2018 (0)"    = "#CC79A7",
                                 "first contest (0)"                  = "#999999")) +
  labs(x = "Expected seats", y = "Seats actually won", colour = NULL,
       title = "Do the expected seats predict the seats Vox actually won?",
       subtitle = "One point per list. Dashed line: perfect prediction.") +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom", panel.grid.minor = element_blank())
```

```{r REGRESSION SEX}
library(sandwich)
library(lmtest)
library(broom)
library(survival)

# Who gets an electable place? The Congress candidates of July 2023 are split in two
# groups by ISCO_CANDIDATOS_VOX_2023.xlsx (isco08_group): those within the electoral
# threshold of their constituency ("Dentro del umbral", the heads of list included) and
# the three titulars right below it ("Justo debajo del umbral"), who are neither heads
# of list nor above the threshold in that election. One row per person
threshold_sample <- vox_candidates |>
  filter(election == "general_2023_07", chamber == "Congress",
         is_titular %in% c(TRUE, NA), !is.na(isco08_group)) |>
  slice_min(list_position, n = 1, by = person_id, with_ties = FALSE) |>
  mutate(electable      = isco08_group == "Dentro del umbral",
         sex            = factor(sex, levels = c("Male", "Female")),
         isco_class     = case_when(isco08_major == "1"  ~ "1 Managers",
                                    isco08_major == "2"  ~ "2 Professionals",
                                    !is.na(isco08_major) ~ "3 Rest",
                                    TRUE                 ~ "4 Unknown") |>
                            factor(levels = c("3 Rest", "1 Managers", "2 Professionals", "4 Unknown")),
         prev_party_any = has_prev_party %in% TRUE) |>
  select(person_id, candidate, province, list_position, electable, isco08_group, sex,
         isco08, isco08_title, isco08_major, isco08_major_title, isco_class,
         prev_party, prev_party_any, in_lists_2013_2017)

threshold_sample |> count(isco08_group, electable)

# the constituencies that identify the comparison: those with people in both groups
threshold_sample |>
  summarise(within = sum(electable), below = sum(!electable), .by = province) |>
  arrange(desc(within)) |>
  print(n = Inf)

# ---- 1. sex: descriptives and Fisher exact test ----------------------------------------
sex_table <- table(sex = threshold_sample$sex, electable = threshold_sample$electable)
sex_table
prop.table(sex_table, margin = 1) |> round(3)
fisher.test(sex_table)

# ---- 2. sex as a predictor of an electable place ----------------------------------------
# the odds of an electable place are compared in two ways:
#   a. a logit over the whole sample, standard errors clustered by constituency
#   b. a conditional logit, which compares the candidates of the same constituency only
#      (constituencies with no electable candidate drop out of this model)
# list_position is not a control here: the two groups are defined by it
threshold_models <- list(
  "1 Logit"             = glm(electable ~ sex, family = binomial, data = threshold_sample),
  "2 Conditional logit" = clogit(electable ~ sex + strata(province), data = threshold_sample)
)

# the table of every model: clustered standard errors for the logits, the model's own for
# the conditional logit, which is already within constituency
threshold_table <- list()
for (m in names(threshold_models)) {
  threshold_table[[m]] <- (if (inherits(threshold_models[[m]], "coxph"))
                             tidy(threshold_models[[m]])
                           else
                             coeftest(threshold_models[[m]],
                                      vcov = vcovCL(threshold_models[[m]], cluster = ~ province)) |>
                               tidy()) |>
    mutate(model = m, n = nobs(threshold_models[[m]]))
}
threshold_table <- bind_rows(threshold_table) |>
  filter(term != "(Intercept)") |>
  mutate(odds_ratio = exp(estimate)) |>
  select(model, term, estimate, std.error, odds_ratio, p.value, n) |>
  mutate(across(c(estimate, std.error, odds_ratio), \(x) round(x, 3)),
         p.value = signif(p.value, 3))

threshold_table |> print(n = Inf)
```

```{r REGRESSION ISCO + SEX}
# The occupation before politics (ISCO-08 major group) added to sex. The four digit codes
# are collapsed into managers (1), professionals (2), the rest (3-9) and "Unknown", the
# candidates whose occupation was not found: they are 2 of the 44 within the threshold and
# 61 of the 137 below it, so the category also measures how much press each group gets
threshold_sample |> count(electable, isco_class)
threshold_sample |> count(electable, isco08_major, isco08_major_title) |> print(n = Inf)

isco_table <- table(isco = threshold_sample$isco_class, electable = threshold_sample$electable)
isco_table
prop.table(isco_table, margin = 1) |> round(3)
fisher.test(isco_table)

# the same, dropping the candidates without a known occupation
fisher.test(isco_table[rownames(isco_table) != "4 Unknown", ])

threshold_models <- list(
  "1 Logit, ISCO"            = glm(electable ~ isco_class, family = binomial, data = threshold_sample),
  "2 Logit, ISCO + sex"      = glm(electable ~ isco_class + sex, family = binomial, data = threshold_sample),
  "3 Conditional logit"      = clogit(electable ~ isco_class + sex + strata(province), data = threshold_sample),
  "4 Logit, ISCO known only" = glm(electable ~ isco_class + sex, family = binomial,
                                   data = filter(threshold_sample, isco_class != "4 Unknown"))
)

threshold_table <- list()
for (m in names(threshold_models)) {
  threshold_table[[m]] <- (if (inherits(threshold_models[[m]], "coxph"))
                             tidy(threshold_models[[m]])
                           else
                             coeftest(threshold_models[[m]],
                                      vcov = vcovCL(threshold_models[[m]], cluster = ~ province)) |>
                               tidy()) |>
    mutate(model = m, n = nobs(threshold_models[[m]]))
}
threshold_table <- bind_rows(threshold_table) |>
  filter(term != "(Intercept)") |>
  mutate(odds_ratio = exp(estimate)) |>
  select(model, term, estimate, std.error, odds_ratio, p.value, n) |>
  mutate(across(c(estimate, std.error, odds_ratio), \(x) round(x, 3)),
         p.value = signif(p.value, 3))

threshold_table |> print(n = Inf)
```


```{r REGRESSION ISCO + SEX + PREV_PARTY}
# A previous party added to the occupation and sex. has_prev_party is TRUE or NA, because
# the absence of a previous party cannot be told apart from the absence of evidence of one;
# to compare the two groups every NA is read here as FALSE (prev_party_any, built with the
# sample), so the coefficient is the effect of a previous party *on record*
threshold_sample |> count(electable, prev_party_any)
threshold_sample |> filter(prev_party_any) |> count(electable, prev_party) |> print(n = Inf)

party_table <- table(prev_party = threshold_sample$prev_party_any, electable = threshold_sample$electable)
party_table
prop.table(party_table, margin = 1) |> round(3)
fisher.test(party_table)

threshold_models <- list(
  "1 Logit, party"              = glm(electable ~ prev_party_any, family = binomial, data = threshold_sample),
  "2 Logit, ISCO + sex + party" = glm(electable ~ isco_class + sex + prev_party_any,
                                      family = binomial, data = threshold_sample),
  "3 Conditional logit"         = clogit(electable ~ isco_class + sex + prev_party_any + strata(province),
                                         data = threshold_sample),
  "4 + lists of 2013-2017"      = glm(electable ~ isco_class + sex + prev_party_any + in_lists_2013_2017,
                                      family = binomial, data = threshold_sample)
)

threshold_table <- list()
for (m in names(threshold_models)) {
  threshold_table[[m]] <- (if (inherits(threshold_models[[m]], "coxph"))
                             tidy(threshold_models[[m]])
                           else
                             coeftest(threshold_models[[m]],
                                      vcov = vcovCL(threshold_models[[m]], cluster = ~ province)) |>
                               tidy()) |>
    mutate(model = m, n = nobs(threshold_models[[m]]))
}
threshold_table <- bind_rows(threshold_table) |>
  filter(term != "(Intercept)") |>
  mutate(odds_ratio = exp(estimate)) |>
  select(model, term, estimate, std.error, odds_ratio, p.value, n) |>
  mutate(across(c(estimate, std.error, odds_ratio), \(x) round(x, 3)),
         p.value = signif(p.value, 3))

threshold_table |> print(n = Inf)

# an odds ratio in the thousands with a huge standard error means a category has no
# variation left (complete separation): read the Fisher tables above instead
```

```{r CIS 2019 POLITICAL DETERMINANTS OF VOX SUPPORT}
library(haven)

# Figure 4.12 of the literature (average marginal effect of the political determinants of
# the support for Vox) rebuilt with the microdata of the postelectoral survey of the general
# election of November 2019: CIS, study 3269 (CIS 2019b)
cis_file <- "MD3269/3269.sav"

if (!file.exists(cis_file)) {
  zip_file <- "MD3269.zip"
  download.file("https://www.cis.es/documents/d/cis/MD3269", zip_file, mode = "wb",
                method = "libcurl", headers = c("User-Agent" = "Mozilla/5.0"))
  unzip(zip_file, exdir = "MD3269", junkpaths = TRUE)
}

cis_2019 <- read_sav(cis_file)

# ---- 1. the variables of the figure ------------------------------------------------------
# B22R      vote recalled in November 2019 (18 = VOX); blank, void, abstention and no answer out
# C3        left-right self-placement, 1 (left) to 10 (right)
# C12A      attendance at religious services: "frequent" is every Sunday or more often; the
#           non-believers, who are not asked about it, do not attend
# B1        interest in politics: "a lot" or "quite a lot"
# B20       did what happened in Catalonia have an effect on your vote? (self-reported)
# A9_1..3   the three most important problems of Spain: 18 immigration, 45 independence of
#           Catalonia. The figure is reproduced with a problem named in any of the three
#           answers; naming it first is the robustness check below
# B30       national identity: "I feel only Spanish" is the exclusionary one
# controls: sex, age, subjective class, education, household income, size of the
#           municipality (the smaller, the more rural) and fixed effects of the autonomous
#           community. Class, education and income keep a "N.C." level, so that the item
#           non-response (40% of the income question) does not empty the sample, and the
#           income brackets are grouped: with the raw ones nobody in the lowest voted VOX
vox_support <- cis_2019 |>
  mutate(vote                = as.numeric(B22R),
         vox                 = as.integer(vote == 18),
         left_right          = if_else(C3 <= 10, as.numeric(C3), NA),
         church_attendance   = case_when(C12A %in% c(5, 6) ~ 1L, C12A == 9 ~ NA_integer_, TRUE ~ 0L),
         political_interest  = case_when(B1 %in% c(1, 2) ~ 1L, B1 %in% c(3, 4) ~ 0L, TRUE ~ NA_integer_),
         catalan_effect      = case_when(B20 == 1 ~ 1L, B20 == 2 ~ 0L, TRUE ~ NA_integer_),
         problem_immigration = as.integer(A9_1 == 18 | A9_2 == 18 | A9_3 == 18) |> replace_na(0L),
         problem_catalonia   = as.integer(A9_1 == 45 | A9_2 == 45 | A9_3 == 45) |> replace_na(0L),
         first_immigration   = as.integer(A9_1 == 18),
         first_catalonia     = as.integer(A9_1 == 45),
         spanish_only        = case_when(B30 == 1 ~ 1L, B30 %in% c(2, 3, 4, 5, 7) ~ 0L, TRUE ~ NA_integer_),
         female              = as.integer(C9 == 2),
         age                 = as.numeric(C10),
         class               = factor(if_else(CLASESUB %in% 1:5, as.character(CLASESUB), "N.C.")),
         education           = factor(if_else(ESTUDIOS %in% 1:6, as.character(ESTUDIOS), "N.C.")),
         income              = case_when(C19 %in% 1:4  ~ "1 up to 900",
                                         C19 %in% 5:6  ~ "2 901-1800",
                                         C19 %in% 7:8  ~ "3 1801-3000",
                                         C19 %in% 9:11 ~ "4 over 3000",
                                         TRUE          ~ "5 N.C.") |> factor(),
         rurality            = 8 - as.numeric(TAMUNI),
         ccaa                = factor(as.numeric(CCAA)),
         weight              = as.numeric(PESO)) |>
  filter(!is.na(vote), !vote %in% c(77, 96, 97, 98, 99)) |>
  select(vox, left_right, church_attendance, political_interest, catalan_effect,
         problem_immigration, problem_catalonia, first_immigration, first_catalonia,
         spanish_only, female, age, class, education, income, rurality, ccaa, weight) |>
  drop_na(vox, left_right, church_attendance, political_interest, catalan_effect,
          problem_immigration, problem_catalonia, first_immigration, first_catalonia,
          spanish_only, female, age, rurality)

cat("voters:", nrow(vox_support), " VOX:", sum(vox_support$vox),
    " share:", round(100 * mean(vox_support$vox), 1), "%\n")

vox_support |>
  summarise(n = n(),
            across(c(left_right, church_attendance, political_interest, catalan_effect,
                     problem_immigration, problem_catalonia, spanish_only), \(x) round(mean(x), 3)),
            .by = vox)

# ---- 2. logit of the vote for Vox ---------------------------------------------------------
# the unweighted logit is the model of the figure; the other two are checks: the same model
# with the weights of the survey, and with the problem named in the first place only
vox_models <- list(
  "1 Logit"          = glm(vox ~ left_right + church_attendance + political_interest +
                             catalan_effect + problem_immigration + problem_catalonia +
                             spanish_only + female + age + class + education + income +
                             rurality + ccaa,
                           family = binomial, data = vox_support),
  "2 Weighted"       = glm(vox ~ left_right + church_attendance + political_interest +
                             catalan_effect + problem_immigration + problem_catalonia +
                             spanish_only + female + age + class + education + income +
                             rurality + ccaa,
                           family = quasibinomial, data = vox_support, weights = weight),
  "3 First problem"  = glm(vox ~ left_right + church_attendance + political_interest +
                             catalan_effect + first_immigration + first_catalonia +
                             spanish_only + female + age + class + education + income +
                             rurality + ccaa,
                           family = binomial, data = vox_support)
)

tidy(vox_models[["1 Logit"]]) |>
  filter(!str_detect(term, "^(ccaa|class|education|income)")) |>
  mutate(odds_ratio = exp(estimate), across(where(is.numeric), \(x) signif(x, 3)))

# the Basque Country has no VOX voter in the sample, so its fixed effect is separated and
# its 110 respondents get a predicted probability of zero
vox_support |> summarise(vox = sum(vox), n = n(), .by = ccaa) |> arrange(vox) |> head(3)

# ---- 3. average marginal effects, in percentage points ------------------------------------
# for the dummies, the average change of the predicted probability between 1 and 0; for the
# left-right scale, the average slope of one point of the scale. The confidence intervals
# are drawn from the coefficients (1000 draws of a normal with the estimated variance matrix)
political_terms <- c("Left-right position"                          = "left_right",
                     "(Frequent) church attendance"                 = "church_attendance",
                     "Political interested"                         = "political_interest",
                     "Catalan crisis had an effect (self-reported)" = "catalan_effect",
                     "Most important problem = Immigration"         = "problem_immigration",
                     "Most important problem = Catalan crisis"      = "problem_catalonia",
                     "Exclusionary Spanish identity"                = "spanish_only")

set.seed(20191110)
margins <- list()
for (m in names(vox_models)) {
  design <- model.matrix(vox_models[[m]])
  beta   <- coef(vox_models[[m]])
  draws  <- t(MASS::mvrnorm(1000, beta, vcov(vox_models[[m]])))

  for (label in names(political_terms)) {
    v <- political_terms[[label]]
    if (m == "3 First problem") v <- str_replace(v, "^problem", "first")
    if (v == "left_right") {                      # continuous: the average derivative
      effects <- colMeans(dlogis(design %*% draws) * rep(draws[v, ], each = nrow(design)))
      point   <- mean(dlogis(design %*% beta) * beta[[v]])
    } else {                                      # dummy: the average discrete change
      design1 <- design; design1[, v] <- 1
      design0 <- design; design0[, v] <- 0
      effects <- colMeans(plogis(design1 %*% draws) - plogis(design0 %*% draws))
      point   <- mean(plogis(design1 %*% beta) - plogis(design0 %*% beta))
    }
    margins[[str_c(m, label)]] <- tibble(model = m, term = label,
                                         ame       = 100 * point,
                                         conf_low  = 100 * quantile(effects, 0.025),
                                         conf_high = 100 * quantile(effects, 0.975))
  }
}
margins <- bind_rows(margins) |>
  mutate(term = factor(term, levels = rev(names(political_terms))))

margins |>
  mutate(across(c(ame, conf_low, conf_high), \(x) round(x, 2))) |>
  arrange(model, desc(term)) |>
  print(n = Inf)

# ---- 4. the figure ------------------------------------------------------------------------
ggplot(filter(margins, model == "1 Logit"), aes(ame, term)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey75", linewidth = 1.2) +
  geom_linerange(aes(xmin = conf_low, xmax = conf_high), colour = "grey35") +
  geom_point(size = 2.2, colour = "grey20") +
  geom_text(aes(label = str_replace(as.character(signif(ame, 2)), "^(-?)0", "\\1")),
            vjust = -1.1, size = 3, family = "serif") +
  scale_x_continuous(limits = c(-5, 15), breaks = seq(-5, 15, 5)) +
  labs(x = NULL, y = NULL,
       title = "Political determinants of support for VOX",
       caption = str_c("Confidence intervals (95%) | Controls not illustrated: gender, age, ",
                       "class (self-reported), education, income, rurality, CCAA fixed effects\n",
                       "Average marginal effects on the probability of voting VOX, in percentage ",
                       "points. Own elaboration with CIS study 3269 (2019b), n = ", nrow(vox_support))) +
  theme_minimal(base_size = 11, base_family = "serif") +
  theme(plot.title = element_text(hjust = 0.5, size = 12),
        plot.caption = element_text(hjust = 0, size = 7, colour = "grey30"),
        panel.grid = element_blank(),
        panel.grid.major.y = element_line(linetype = "dotted", colour = "grey80"),
        axis.line.x = element_line(colour = "grey40"),
        axis.ticks.x = element_line(colour = "grey40"),
        plot.margin = margin(10, 20, 10, 10))
```

