---
title: "VOX_CANDIDATES"
author: "Noah Emanuel Fürup"
format: pdf
editor: visual
---

```{r LIBRARIES}
library(tidyverse)
library(readxl)
library(sf)
library(rnaturalearth)
library(countrycode)
```

#MAP ELECTORAL SYSTEMS:

```{r READ IDEA ELECTORAL SYSTEM DATASET}
idea <- read_excel("idea_export_electoral_system_design_database.xlsx")
names(idea)   # mira aquí cómo se llaman de verdad las columnas

```

```{r}
systems <- idea |>
  filter(!is.na(ISO3), !is.na(`Electoral system family`)) |>
  group_by(ISO3) |>
  slice_max(Year, n = 1, with_ties = FALSE) |>
  ungroup() |>
  select(ISO3, Country, Year,
         family = `Electoral system family`,
         system = `Electoral system for national legislature`)

europe <- ne_countries(scale = "medium", continent = "Europe", returnclass = "sf") |>
  filter(admin != "Russia") |>
  left_join(systems, by = c("iso_a3" = "ISO3"))

ggplot(europe) +
  geom_sf(aes(fill = family), colour = "white", linewidth = 0.2) +
  coord_sf(xlim = c(-25, 45), ylim = c(34, 72)) +
  scale_fill_brewer(palette = "Set2", na.value = "grey90", name = "Electoral system family") +
  labs(title = "Electoral system families in Europe",
       caption = "Source: International IDEA, Electoral System Design Database") +
  theme_void()

# ggsave("europe_electoral_systems.png", width = 10, height = 7, dpi = 300)

# what didn't match
europe |> st_drop_geometry() |> filter(is.na(family)) |> select(admin)
```

```{r}
# the exact labels first
europe |> st_drop_geometry() |> count(family)

# share of each family
europe |>
  st_drop_geometry() |>
  filter(!is.na(family)) |>
  count(family) |>
  mutate(pct = round(100 * n / sum(n), 1)) |>
  arrange(desc(n))

# PR only, as a single figure
europe |>
  st_drop_geometry() |>
  filter(!is.na(family)) |>
  summarise(countries = n(),
            pr = sum(str_detect(family, "Proportional")),
            pct_pr = round(100 * pr / countries, 1))
```

Extraction of non-autonomical elections candidates: european and general elections

Source: - 2014-2019: github.com/JaimeObregon/infoelectoral - 2023-2024: infoelectoral.interior.gob.es

```{r SPANISH LETERS}
latin1 <- locale(encoding = "ISO-8859-1")
```

```{r GITHUB MIRROR}
MIRROR <- "https://raw.githubusercontent.com/JaimeObregon/infoelectoral/main/files"
```

The github has the files in .DAT, a file type that doesn't separate in ";" or "\|". In order to separate each column, we need to map where the separation between each variable occurs. I asked Claude to give me the exact character positions where the separation is. I used fwf_positions of ReadR to write down the column names, given the positions where the variables lie in the .DAT file.

#GENERAL ELECTIONS

```{r FIXED POSITIONS}
# File 03 — candidacies (parties)
cols_03 <- fwf_cols(
  party_code   = c(  9,  14),
  abbreviation = c( 15,  64),
  full_name    = c( 65, 214)
)

# File 04 — candidates (people)
cols_04 <- fwf_cols(
  province_code  = c( 10,  11),
  party_code     = c( 16,  21),
  list_position  = c( 22,  24),
  candidate_type = c( 25,  25),   # T = titular, S = substitute
  first_name     = c( 26,  50),
  surname_1      = c( 51,  75),
  surname_2      = c( 76, 100),
  sex            = c(101, 101),   # M / F
  elected        = c(120, 120)    # S = elected
)

# up to 2000 the name is not split: one field, and no sex column
cols_04_single <- fwf_cols(
  province_code  = c( 10,  11),
  party_code     = c( 16,  21),
  list_position  = c( 22,  24),
  candidate_type = c( 25,  25),
  first_name     = c( 26, 100),   # name and both surnames together
  elected        = c(120, 120)
)
```

```{r ELECTIONS 2014-2019}
LOCAL_MIRROR <- "infoelectoral/files"
MIRROR <- "https://raw.githubusercontent.com/JaimeObregon/infoelectoral/main/files"
PORTAL <- "https://infoelectoral.interior.gob.es/estaticos/docxl/apliextr"

elections <- tribble(
  ~process, ~year,  ~month, ~chamber,              ~mirror_folder,
  "02",     "1977", "06",   "Congress",            "congreso",
  "02",     "1979", "03",   "Congress",            "congreso",
  "02",     "1982", "10",   "Congress",            "congreso",
  "02",     "1986", "06",   "Congress",            "congreso",
  "02",     "1989", "10",   "Congress",            "congreso",
  "02",     "1993", "06",   "Congress",            "congreso",
  "02",     "1996", "03",   "Congress",            "congreso",
  "02",     "2000", "03",   "Congress",            "congreso",
  "02",     "2004", "03",   "Congress",            "congreso",
  "02",     "2008", "03",   "Congress",            "congreso",
  "02",     "2011", "11",   "Congress",            "congreso",
  "02",     "2015", "12",   "Congress",            "congreso",
  "02",     "2016", "06",   "Congress",            "congreso",
  "02",     "2019", "04",   "Congress",            "congreso",
  "02",     "2019", "11",   "Congress",            "congreso",
  "02",     "2023", "07",   "Congress",            "congreso",
  "03",     "1977", "06",   "Senate",              "senado",
  "03",     "1979", "03",   "Senate",              "senado",
  "03",     "1982", "10",   "Senate",              "senado",
  "03",     "1986", "06",   "Senate",              "senado",
  "03",     "1989", "10",   "Senate",              "senado",
  "03",     "1993", "06",   "Senate",              "senado",
  "03",     "1996", "03",   "Senate",              "senado",
  "03",     "2000", "03",   "Senate",              "senado",
  "03",     "2004", "03",   "Senate",              "senado",
  "03",     "2008", "03",   "Senate",              "senado",
  "03",     "2011", "11",   "Senate",              "senado",
  "03",     "2015", "12",   "Senate",              "senado",
  "03",     "2016", "06",   "Senate",              "senado",
  "03",     "2019", "04",   "Senate",              "senado",
  "03",     "2019", "11",   "Senate",              "senado",
  "03",     "2023", "07",   "Senate",              "senado",
  "07",     "2014", "05",   "European Parliament", "europeas",
  "07",     "2019", "05",   "European Parliament", "europeas",
  "07",     "2024", "06",   "European Parliament", "europeas"
)

```

```{r PROVINCES}
provinces <- tribble(
  ~province_code, ~province,
  "01", "Araba/Alava",    "02", "Albacete",      "03", "Alicante",
  "04", "Almeria",        "05", "Avila",         "06", "Badajoz",
  "07", "Illes Balears",  "08", "Barcelona",     "09", "Burgos",
  "10", "Caceres",        "11", "Cadiz",         "12", "Castellon",
  "13", "Ciudad Real",    "14", "Cordoba",       "15", "A Coruna",
  "16", "Cuenca",         "17", "Girona",        "18", "Granada",
  "19", "Guadalajara",    "20", "Gipuzkoa",      "21", "Huelva",
  "22", "Huesca",         "23", "Jaen",          "24", "Leon",
  "25", "Lleida",         "26", "La Rioja",      "27", "Lugo",
  "28", "Madrid",         "29", "Malaga",        "30", "Murcia",
  "31", "Navarra",        "32", "Ourense",       "33", "Asturias",
  "34", "Palencia",       "35", "Las Palmas",    "36", "Pontevedra",
  "37", "Salamanca",      "38", "S.C. Tenerife", "39", "Cantabria",
  "40", "Segovia",        "41", "Sevilla",       "42", "Soria",
  "43", "Tarragona",      "44", "Teruel",        "45", "Toledo",
  "46", "Valencia",       "47", "Valladolid",    "48", "Bizkaia",
  "49", "Zamora",         "50", "Zaragoza",      "51", "Ceuta",
  "52", "Melilla",        "99", "Single national constituency"
)
```

```{r CODES MUNICIPALITIES}
# 26codmun and 26codislas
```

```{r DOWNLOAD FILES GITHUB}
candidate_files <- list()

for (i in seq_len(nrow(elections))) {

  process       <- elections$process[i]
  year          <- elections$year[i]
  month         <- elections$month[i]
  chamber       <- elections$chamber[i]
  mirror_folder <- elections$mirror_folder[i]

  yy     <- str_sub(year, 3, 4)
  folder <- str_c(process, year, month, "_TOTA")

  local_dir <- str_c(LOCAL_MIRROR, "/", mirror_folder, "/", folder)

  # the GitHub mirror lacks a few processes (Congress 1989 and 1993, and
  # everything after 2019), so take those from the ministry's download area
  if (!dir.exists(local_dir)) {
    message("Fetching ", folder, " from the ministry portal ...")
    zip_file <- tempfile(fileext = ".zip")
    download.file(str_c(PORTAL, "/", folder, ".zip"), zip_file,
                  mode = "wb", quiet = TRUE)
    dir.create(local_dir, recursive = TRUE, showWarnings = FALSE)
    unzip(zip_file, exdir = local_dir)
  }

  path_03 <- list.files(local_dir, str_c("^03", process, yy, month, "[.]DAT$"),
                        full.names = TRUE, recursive = TRUE, ignore.case = TRUE)[1]
  path_04 <- list.files(local_dir, str_c("^04", process, yy, month, "[.]DAT$"),
                        full.names = TRUE, recursive = TRUE, ignore.case = TRUE)[1]

  message("Reading ", chamber, " ", year, "-", month, " ...")

  parties <- read_fwf(path_03, cols_03, locale = latin1,
                      col_types = cols(.default = "c")) |>
    mutate(across(everything(), str_trim))

  # up to 2000 the name comes in a single field; from 2004 it is split in three.
  # Decide from the data: the second-surname block is only filled in the split one.
  probe <- read_lines(path_04, n_max = 500, locale = latin1)
  split <- mean(str_squish(str_sub(probe, 76, 100)) != "") > 0.5

  raw <- read_fwf(path_04, if (split) cols_04 else cols_04_single,
                  locale = latin1, col_types = cols(.default = "c")) |>
    mutate(across(everything(), str_trim))

  # give every process the same columns, so bind_rows and transmute never break
  if (!"surname_1" %in% names(raw)) raw <- mutate(raw, surname_1 = NA_character_)
  if (!"surname_2" %in% names(raw)) raw <- mutate(raw, surname_2 = NA_character_)
  if (!"sex"       %in% names(raw)) raw <- mutate(raw, sex       = NA_character_)

  candidate_files[[i]] <- raw |>
    left_join(parties, by = "party_code") |>
    mutate(chamber = chamber, year = year, month = month)
}

all_candidates <- bind_rows(candidate_files) |>
  left_join(provinces, by = "province_code") |>
  transmute(
    election      = str_c(if_else(chamber == "European Parliament",
                                  "european", "general"),
                          "_", year, "_", month),
    year          = as.integer(year),
    chamber,
    province,
    list_position = as.integer(list_position),
    # str_c returns NA if ANY part is NA, and the surnames are empty before 2004
    candidate     = str_squish(str_c(coalesce(first_name, ""), " ",
                                     coalesce(surname_1,  ""), " ",
                                     coalesce(surname_2,  ""))),
    sex           = recode(sex, "M" = "Male", "F" = "Female",
                           .default = NA_character_),
    is_titular    = candidate_type == "T",
    is_elected    = elected == "S",
    abbreviation,
    level         = "National"
  ) |>
  arrange(year, chamber, province, list_position)

all_candidates |> count(year, chamber) |> print(n = Inf)
all_candidates |>
  filter(str_detect(coalesce(abbreviation, ""), regex("vox", ignore_case = TRUE))) |>
  count(year, chamber) |> print(n = Inf)
```

```{r JOIN PROVINCES AND TRANSMUTE}
all_candidates <- bind_rows(candidate_files) |>
  left_join(provinces, by = "province_code") |>
  transmute(
    election      = str_c(if_else(chamber == "European Parliament",
                                  "european", "general"),
                          "_", year, "_", month),
    year          = as.integer(year),
    chamber,
    province,
    list_position = as.integer(list_position),
    # str_c returns NA if ANY part is NA, and the surnames are empty before 2004
    candidate     = str_squish(str_c(coalesce(first_name, ""), " ",
                                     coalesce(surname_1,  ""), " ",
                                     coalesce(surname_2,  ""))),
    sex           = recode(sex, "M" = "Male", "F" = "Female",
                           .default = NA_character_),
    is_titular    = candidate_type == "T",
    is_elected    = elected == "S",
    abbreviation,
    level         = "National"
  ) |>
  arrange(year, chamber, province, list_position)

# sanity check: every process must be present
all_candidates |> count(year, chamber) |> print(n = Inf)
all_candidates |>
  filter(str_detect(coalesce(abbreviation, ""), regex("vox", ignore_case = TRUE))) |>
  count(year, chamber) |> print(n = Inf)
```

```{r ADD 2023-2024 GENERAL AND EUROPEAN ELECTIONS}
PORTAL <- "https://infoelectoral.interior.gob.es/estaticos/docxl/apliextr"

elections_new <- tribble(
  ~process, ~year,  ~month, ~chamber,
  "02",     "2023", "07",   "Congress",
  "03",     "2023", "07",   "Senate",
  "07",     "2024", "06",   "European Parliament"
)
```

```{r DOWNLOAD 2023-2024}
portal_candidates <- list()

for (i in seq_len(nrow(elections_new))) {

  process <- elections_new$process[i]
  year    <- elections_new$year[i]
  month   <- elections_new$month[i]
  chamber <- elections_new$chamber[i]

  folder   <- str_c(process, year, month, "_TOTA")
  zip_file <- tempfile(fileext = ".zip")
  unzip_to <- tempfile()

  message("Downloading ", chamber, " ", year, "-", month, " ...")

  download.file(str_c(PORTAL, "/", folder, ".zip"), zip_file,
                mode = "wb", quiet = TRUE)
  unzip(zip_file, exdir = unzip_to)

  path_03 <- list.files(unzip_to, "^03.*[.]DAT$", full.names = TRUE,
                        recursive = TRUE, ignore.case = TRUE)[1]
  path_04 <- list.files(unzip_to, "^04.*[.]DAT$", full.names = TRUE,
                        recursive = TRUE, ignore.case = TRUE)[1]

  parties <- read_fwf(path_03, cols_03, locale = latin1,
                      col_types = cols(.default = "c")) |>
    mutate(across(everything(), str_trim))

  portal_candidates[[i]] <- read_fwf(path_04, cols_04, locale = latin1,
                                  col_types = cols(.default = "c")) |>
    mutate(across(everything(), str_trim)) |>
    left_join(parties, by = "party_code") |>
    mutate(chamber = chamber, year = year, month = month)
}
```

```{r JOIN PROVINCES AND TRANSMUTE}
portal_candidates <- bind_rows(portal_candidates) |>
  left_join(provinces, by = "province_code") |>
  transmute(
    election      = str_c(if_else(chamber == "European Parliament",
                                  "european", "general"),
                          "_", year, "_", month),
    year          = as.integer(year),
    chamber,
    province,
    list_position = as.integer(list_position),
        candidate = str_squish(str_c(coalesce(first_name, ""), " ",
                                 coalesce(surname_1,  ""), " ",
                                 coalesce(surname_2,  ""))),
    sex           = recode(sex, "M" = "Male", "F" = "Female",
                           .default = NA_character_),
    is_titular    = candidate_type == "T",
    is_elected    = elected == "S",
    abbreviation,
    full_name
  )

all_candidates <- bind_rows(all_candidates, portal_candidates) |>
  arrange(year, chamber, province, list_position) |> 
  select(-full_name)
```

```{r TAKE DON/ DOÑA OUT}
# Handled upstream in READ ALL AUTONOMICAL CANDIDATES: the honorific is stripped
# from `candidate `before the union (this chunk referred to `candidato_norm'
# a column that does not exist in autonomical_candidates).
all_candidates <- all_candidates |>
  mutate(candidate = str_squish(str_remove(candidate,
                                           regex("^(don|do[ñn]a|d[.ª])\\s+",
                                                 ignore_case = TRUE))))
```

```{r FILTER PP, CS AND VOX}
pp_candidates <- all_candidates |> 
  filter(str_detect(abbreviation, "PP"))

vox_candidates <- all_candidates |> 
  filter(str_detect(abbreviation, "VOX"))

cs_candidates <- all_candidates |> 
  filter(str_detect(abbreviation, "C's"))
  
```

#AUTONOMICAL ELECTIONS

Claude Code For Autonomical elections: two attempts to confirm it is reproducible.

Prompt 1:

"Quiero que busques todas las listas de todas las elecciones autonómicas que han habido en España desde 2015 hasta 2026 y que pongas en el archivo excel en el excel de CANDIDATOS_AUTONOMICAS. Tienes que ir a la web para recopilar las siguientes variables de cada eleccion: CANDIDATO, CCAA, ELECCION, CIRCUNSCRIPCION, ORDEN, ELEGIDO (TRUE/FALSE)"

Prompt 2:

"You're a Computational Social Scientist. I want you to search for every candidate list of all autonomical elections in Spain from 2015 to 2026 of the following parties: PP and VOX. You have to get into the web and recopilate the following variables for each election:

election: election (write it in like this: "autonomous_2015_05", where 2015 would be the year and 05 the month where the election took place) year: year of the election chamber: the chamber (such as Parlament of Andalucia, Parlament of Catalonia...) province: the province that was being represented by the candidate list_position: the position of the candidate within that list candidate: the candidate's full name is_titular: whether he was the titular of that seat or he was a substitution candidate (TRUE/FALSE), is_elected: whether he was elected (TRUE/FALSE) abbreviation: the abbreviation of the party's name

the final file should be a csv file."

```{r TOTAL ELECTIONS AUTONOMICAL/MUNICIPAL}
vox_candidates |>
  filter(level %in% c("Autonomical", "Municipal")) |>
  group_by(level) |>
  summarise(elections = n_distinct(election), years = n_distinct(year))

# autonomicas: cada region vota por su cuenta, asi que la eleccion es region-anyo
vox_candidates |>
  filter(level == "Autonomical") |>
  distinct(chamber, year) |>
  count(chamber, name = "elections") |>
  arrange(desc(elections))

vox_candidates |>
  filter(level == "Municipal") |>
  count(election, year, name = "candidates")
```

```{r READ ALL AUTONOMICAL CANDIDATES}
# =============================================================================
# PP and VOX candidates in the regional (autonomic) elections
#
# These do NOT come from infoelectoral: the Ministry of the Interior does not
# publish regional elections ("son competencia de cada Comunidad Autonoma"), so
# both files were built from the official proclamation bulletins collected by the
# Junta Electoral Central. That is why they are read from disk instead of a URL.
#
#   2000-2012 : PP only          (176 constituency-lists, from the JEC bulletins)
#   2015-2026 : PP and VOX
# =============================================================================

autonomical_pre2015 <- read_csv(
  "autonomicas_scrape/out_pre2015/pp_autonomicas_2000_2012.csv",
  show_col_types = FALSE
) |>
  transmute(
    election,
    year          = as.integer(year),
    chamber,
    province,
    list_position = as.integer(list_position),
    candidate     = str_squish(candidate),
    sex,
    is_titular,
    is_elected,                      # not resolved for this period: all NA
    abbreviation,
    level         = "Autonomical",
    source        = source_doc,
    list_completeness
  )

autonomical_2015_on <- read_csv(
  "candidatos_pp_vox_autonomicas_2015_2026.csv",
  show_col_types = FALSE
) |>
  transmute(
    election,
    year          = as.integer(year),
    chamber,
    province,
    list_position = as.integer(list_position),
    candidate     = str_squish(candidate),
    sex,
    is_titular,
    is_elected,
    abbreviation,
    level         = "Autonomical",
    source        = source_doc)

autonomical_candidates <- bind_rows(autonomical_pre2015, autonomical_2015_on) |>
  arrange(year, chamber, province, desc(is_titular), list_position)
```

```{r READ CIUDADANOS CANDIDATES AUTONOMICAL}
cs_candidates_autonomical <- read_csv("candidatos_cs_autonomicas_2015_2026.csv")
cs_candidates_autonomical
```

```{r FILTER PP AND VOX CANDIDATES}
pp_candidates <- pp_candidates |> 
  bind_rows(pp_candidates, autonomical_candidates |> filter(str_detect(abbreviation, "PP")))

vox_candidates <- vox_candidates |> 
  bind_rows(vox_candidates, autonomical_candidates |> filter(str_detect(abbreviation, "VOX")))

cs_candidates <- cs_candidates |> 
  bind_rows(cs_candidates, cs_candidates_autonomical)

```

```{r}

```

```{r ADD SEX}
# library(readxl)
# hombres <- read_excel("nombres_por_edad_media.xlsx", sheet = "Hombres", skip = 6) |>
#   rename(name = 2, freq = 3) |>
#   transmute(name = as.character(name), freq = as.numeric(freq), sex_ine = "Male")
# 
# mujeres <- read_excel("nombres_por_edad_media.xlsx", sheet = "Mujeres", skip = 6) |>
#   rename(name = 2, freq = 3) |>
#   transmute(name = as.character(name), freq = as.numeric(freq), sex_ine = "Female")
# 
# ine_names <- bind_rows(hombres, mujeres) |>
#   mutate(name = name |>
#            str_to_upper() |>
#            stringi::stri_trans_general("Latin-ASCII") |>
#            str_remove_all("[^A-Z ]") |>
#            str_squish()) |>
#   filter(!is.na(name), name != "") |>
#   group_by(name) |>
#   slice_max(freq, n = 1, with_ties = FALSE) |>
#   ungroup() |>
#   select(name, sex_ine)
# 
# autonomical_candidates <- autonomical_candidates |>
#   mutate(
#     clean = candidate |>
#       str_to_upper() |>
#       stringi::stri_trans_general("Latin-ASCII") |>
#       str_remove_all("[^A-Z ]") |>
#       str_squish(),
#     tok1 = word(clean, 1),
#     tok2 = word(clean, 1, 2),
#     tok3 = word(clean, 1, 3)
#   ) |>
#   left_join(rename(ine_names, s1 = sex_ine), by = c("tok1" = "name")) |>
#   left_join(rename(ine_names, s2 = sex_ine), by = c("tok2" = "name")) |>
#   left_join(rename(ine_names, s3 = sex_ine), by = c("tok3" = "name")) |>
#   mutate(sex = coalesce(s3, s2, s1)) |>
#   select(-clean, -tok1, -tok2, -tok3, -s1, -s2, -s3)

```

Added 103 changes manually in sex in excel.

# MUNICIPAL ELECTIONS

```{r READ ALL VOX MUNICIPAL CANDIDATES}
# municipality -> province, from the INE list (one sheet per province). The key
# ignores accents, case and the order of the bilingual halves, because the file
# writes "Alicante/Alacant" and the INE "Alacant/Alicante".
municipalities <- list()
for (s in excel_sheets("26codmun.xlsx")) {
  municipalities[[s]] <- read_excel("26codmun.xlsx", sheet = s, skip = 2,
                                    col_types = "text") |>
    filter(!is.na(CMUN)) |>
    transmute(province_code = CPRO, municipality = NOMBRE)
}
municipalities <- bind_rows(municipalities) |>
  left_join(provinces, by = "province_code") |>
  mutate(muni_key = municipality |>
           str_to_upper() |>
           stringi::stri_trans_general("Latin-ASCII") |>
           str_split("/") |>
           map_chr(~ str_c(sort(str_squish(.x)), collapse = "/"))) |>
  distinct(muni_key, province)

municipal_vox <- readxl::read_xlsx("CANDIDATOS_MUNICIPALES.xlsx") |>
  mutate(muni_key = CIRCUNSCRIPCION |>
           str_to_upper() |>
           stringi::stri_trans_general("Latin-ASCII") |>
           str_split("/") |>
           map_chr(~ str_c(sort(str_squish(.x)), collapse = "/"))) |>
  left_join(municipalities, by = "muni_key") |>
  transmute(
    election      = ELECCION,
    year          = as.integer(str_extract(ELECCION, "\\d{4}")),
    chamber       = CIRCUNSCRIPCION,   # the municipality
    province,                          # its province, from the INE list
    list_position = as.integer(ORDEN),
    candidate     = CANDIDATO,
    is_elected    = ELEGIDO,
    source        = FUENTE,
    level         = "Municipal"
  )
# municipalities the INE list no longer carries under that name (it is dated
# 1 January 2026 and these data reach 2015), assigned by hand
manual_provinces <- tribble(
  ~chamber,                       ~province_manual,
  "Castell-Platja d'Aro",         "Girona",
  "Sant Carles de la Ràpita",     "Tarragona",
  "Moncada",                      "Valencia",
  "Alfarp",                       "Valencia",
  "Castelló de la Plana",         "Castellon",
  "Guadiana del Caudillo",        "Badajoz",
  "Maó-Mahón",                    "Illes Balears"
)

municipal_vox <- municipal_vox |>
  left_join(manual_provinces, by = "chamber") |>
  mutate(province = coalesce(province, province_manual)) |>
  select(-province_manual)
```

```{r PP MUNICIPAL CANDIDATES}
# PP candidates in the municipal elections. The ministry's file 04 for municipal
# elections carries the municipality, and up to 2003 the name is in one field, so
# read 26-100 as a single block: in the split layout the three fields are adjacent
# and str_squish collapses the padding to the same result.
PORTAL <- "https://infoelectoral.interior.gob.es/estaticos/docxl/apliextr"
dir.create("datos_pp_locales", showWarnings = FALSE)

cols_04_local <- fwf_cols(
  province_code  = c( 10,  11),
  muni_code      = c( 13,  15),   # 12 is a control digit
  party_code     = c( 16,  21),
  list_position  = c( 22,  24),
  candidate_type = c( 25,  25),
  full_name_cand = c( 26, 100),
  sex            = c(101, 101),
  elected        = c(120, 120)
)

# municipality names, one sheet per province
municipalities <- list()
for (s in excel_sheets("26codmun.xlsx")) {
  municipalities[[s]] <- read_excel("26codmun.xlsx", sheet = s, skip = 2,
                                    col_types = "text")
}
municipalities <- bind_rows(municipalities) |>
  filter(!is.na(CMUN)) |>
  transmute(province_code = CPRO, muni_code = CMUN, municipality = NOMBRE)

local_candidates <- list()
years <- c("2003", "2007", "2011", "2015", "2019", "2023")

for (i in seq_along(years)) {

  folder   <- str_c("04", years[i], "05_TOTA")
  zip_file <- str_c("datos_pp_locales/", folder, ".zip")
  unzip_to <- tempfile()

  if (!file.exists(zip_file))
    download.file(str_c(PORTAL, "/", folder, ".zip"), zip_file, mode = "wb", quiet = TRUE)
  unzip(zip_file, exdir = unzip_to)

  path_03 <- list.files(unzip_to, "^03.*[.]DAT$", full.names = TRUE,
                        recursive = TRUE, ignore.case = TRUE)[1]
  path_04 <- list.files(unzip_to, "^04.*[.]DAT$", full.names = TRUE,
                        recursive = TRUE, ignore.case = TRUE)[1]

  # the full name is the reliable filter: many local parties are abbreviated PPx
  # ("PPN" is Partido Progresista Nerjeño, "P.P.SO." Plataforma del Pueblo Soriano)
  parties <- read_fwf(path_03, cols_03, locale = latin1,
                      col_types = cols(.default = "c")) |>
    mutate(across(everything(), str_trim)) |>
    filter(str_detect(stringi::stri_trans_general(str_to_upper(full_name),
                                                  "Latin-ASCII"),
                      "PARTIDO POPULAR|PARTIT POPULAR"))

  message("Reading municipal ", years[i], " ...")

  local_candidates[[i]] <- read_fwf(path_04, cols_04_local, locale = latin1,
                                    col_types = cols(.default = "c")) |>
    mutate(across(everything(), str_trim)) |>
    inner_join(parties, by = "party_code") |>
    mutate(year = years[i])
}

municipal_pp <- bind_rows(local_candidates) |>
  left_join(municipalities, by = c("province_code", "muni_code")) |>
  left_join(provinces, by = "province_code") |>
  transmute(
    election      = str_c("municipal_", year, "_05"),
    year          = as.integer(year),
    chamber       = municipality,
    province,
    list_position = as.integer(list_position),
    candidate     = str_squish(full_name_cand),
    sex           = recode(sex, "M" = "Male", "F" = "Female", .default = NA_character_),
    is_titular    = candidate_type == "T",
    is_elected    = elected == "S",
    abbreviation  = "PP",
    level         = "Municipal"
  ) |>
  arrange(year, province, chamber, list_position)
```

```{r FILTER PP AND VOX CANDIDATES}
pp_candidates <-
  bind_rows(pp_candidates, municipal_pp)

vox_candidates <-
  bind_rows(vox_candidates, municipal_vox)

```

```{r ADD SEATS DISTRICT AND PARTY}
seats_district <- all_candidates |>
  filter(is_elected, is_titular = TRUE) |>
  count(election, chamber, province, name = "seats_district")

party_seats <- all_candidates |>
  filter(is_elected, is_titular) |>
  count(election, chamber, province, abbreviation, name = "seats_party")

all_candidates <- all_candidates |>
  left_join(seats_district,   by = c("election", "chamber", "province")) |>
  left_join(party_seats, by = c("election", "chamber", "province", "abbreviation")) |>
  mutate(seats_party = replace_na(seats_party, 0))
```

```{r NORMALIZE NAMES AND ELECTION DATE}
vox_candidates <- vox_candidates |>
  mutate(
    # crude person id: uppercase, no accents, no punctuation
    person_id = candidate |>
      str_to_upper() |>
      stringi::stri_trans_general("Latin-ASCII") |>
      str_remove_all("[^A-Z ]") |>
      str_squish()
    # Date?
  ) 

pp_candidates <- pp_candidates |>
  mutate(
    # crude person id: uppercase, no accents, no punctuation
    person_id = candidate |>
      str_to_upper() |>
      stringi::stri_trans_general("Latin-ASCII") |>
      str_remove_all("[^A-Z ]") |>
      str_squish()
    # Date?
  ) 

cs_candidates <- cs_candidates |>
  mutate(
    # crude person id: uppercase, no accents, no punctuation
    person_id = candidate |>
      str_to_upper() |>
      stringi::stri_trans_general("Latin-ASCII") |>
      str_remove_all("[^A-Z ]") |>
      str_squish()
    # Date?
  ) 

```

```{r TABLE ELECTED NON ELECTED}
vox_candidates |>
  summarise(
    total_candidates = n_distinct(person_id, na.rm = TRUE),
    total_elected    = n_distinct(person_id[is_elected %in% TRUE], na.rm = TRUE)
  ) 
```

```{r VOX CANDIDATES THAT COME FROM ALTERNATIVA DE GOVERN}
read_csv("alternativa_de_govern_candidates.csv") |> 
  mutate(
    # crude person id: uppercase, no accents, no punctuation
    person_id = candidate |>
      str_to_upper() |>
      stringi::stri_trans_general("Latin-ASCII") |>
      str_remove_all("[^A-Z ]") |>
      str_squish()
    # Date?
  ) |> 
  mutate(from_alternativa = person_id %in% vox_candidates$person_id[!is.na(vox_candidates$person_id)]) |>
  filter(from_alternativa == TRUE) |> 
  select(candidate, chamber)
  
```

```{r FROM PP VARIABLE}
vox_candidates <- vox_candidates |>
  mutate(from_pp = person_id %in% pp_candidates$person_id[!is.na(pp_candidates$person_id)])
```

Manual from PP

```{r RADICAL RIGHT CANDIDATES}
radical_right_candidates <- bind_rows(
  read_excel("CANDIDATOS_PARTIDOS_HISTORICOS.xlsx", sheet = "01 Candidatos") |>
    select(Partido, `Tipo de candidatura`, Candidato, Anyo, Nivel),
  read_excel("CANDIDATOS_PARTIDOS_HISTORICOS.xlsx", sheet = "06 Autonomicas") |>
    select(Partido, `Tipo de candidatura`, Candidato, Anyo, Calidad) |>
    mutate(Nivel = "Autonomicas")) |>
  mutate(person_id = Candidato |>
           str_to_upper() |>
           stringi::stri_trans_general("Latin-ASCII") |>
           str_remove_all("[^A-Z ]") |>
           str_squish())
```

```{r CS CANDIDATES}

cs_candidates <- read_csv("candidatos_cs_autonomicas_2015_2026.csv", show_col_types = FALSE) |>
  bind_rows(all_candidates |> filter(abbreviation == "C's")) |>
  mutate(person_id = candidate |>
           str_to_upper() |>
           stringi::stri_trans_general("Latin-ASCII") |>
           str_remove_all("[^A-Z ]") |>
           str_squish())
```

```{r}
# =============================================================================
# People listed on the websites of the satellite organisations of VOX,
# read from the Internet Archive. Pure R, ~2 minutes, writes one CSV.
#
# Two things keep it fast: only the pages that list people are asked for (one
# CDX query per path prefix, collapsed by digest so that only real content
# changes come back), and at most one capture per page and year is downloaded,
# all of them in parallel.
#
# The date is the date of the CAPTURE, not of the appointment: the person was
# listed on that page on that day, not necessarily appointed then.
# =============================================================================

library(tidyverse)
library(httr2)
library(rvest)
library(readxl)

# path prefixes that hold the team, the board, the faculty
pages <- tribble(
  ~organisation, ~prefix,
  "DENAES",      "denaes.es/quienes-somos",
  "Disenso",     "fundaciondisenso.org/nosotros",
  "ISSEP",       "issep.es/nosotros",
  "ISSEP",       "issep.es/profesores",
  "HazteOir",    "hazteoir.org/conocenos",
  "HazteOir",    "www.hazteoir.org/conocenos",
  "CitizenGO",   "citizengo.org/es/conocenos",
  "CitizenGO",   "www.citizengo.org/es/conocenos",
  "CitizenGO",   "citizengo.org/en/about-us",
  "Solidaridad", "sindicatosolidaridad.es/contacto"
)

# --- 1. what the Archive has --------------------------------------------------
captures <- pages |>
  mutate(cdx = map(prefix, \(p)
    request("http://web.archive.org/cdx/search/cdx") |>
      req_url_query(url = p, matchType = "prefix", output = "json",
                    fl = "timestamp,original", collapse = "digest",
                    filter = c("statuscode:200", "mimetype:text/html"),
                    limit = 2000, .multi = "explode") |>
      req_retry(max_tries = 3) |>
      req_perform() |> resp_body_json() |>
      (\(x) tibble(timestamp = map_chr(x[-1], 1), original = map_chr(x[-1], 2)))())) |>
  unnest(cdx) |>
  mutate(date = ymd(str_sub(timestamp, 1, 8)), year = year(date),
         snapshot = str_c("https://web.archive.org/web/", timestamp, "id_/", original)) |>
  # one capture per page and year is enough to follow who is on a board
  slice_max(date, n = 1, by = c(original, year), with_ties = FALSE)

# --- 2. download them in parallel ---------------------------------------------
responses <- captures$snapshot |>
  map(\(u) request(u) |> req_timeout(30) |> req_error(is_error = \(r) FALSE)) |>
  req_perform_parallel(max_active = 5, on_error = "continue", progress = TRUE)

lines <- captures |>
  mutate(html = map_chr(responses, possibly(\(r) resp_body_string(r), NA_character_))) |>
  filter(!is.na(html)) |>
  # only the code is dropped. These sites wrap their content in <header> and
  # <aside>, so removing those takes the board away with the menu; the census
  # test below is what actually separates people from navigation.
  mutate(text = map(html, \(h) {
    page <- read_html(h)
    page |> html_elements("script, style, noscript") |> xml2::xml_remove()
    page |> html_elements("body") |> html_text2() |> str_split_1("\n") |> str_squish()
  }), .keep = "unused") |>
  unnest(text) |>
  filter(text != "")

# --- 3. which of those lines is a person --------------------------------------
PARTICLES <- c("de", "del", "la", "las", "los", "y", "e", "da", "do", "van", "von", "i")

# labels, menus and headlines the capitalisation rule would otherwise accept
NOT_A_PERSON <- str_c(
  "EQUIPO|CONSEJO|PATRONATO|DIRECCION|COMITE|JUNTA|FUNDACION|INSTITUTO|SINDICATO|",
  "ASOCIACION|UNIVERSIT|SCIENCES PO|COLLEGE|SCHOOL|REVISTA|DIARIO|DIGEST|OBJECTIVE|",
  "TELEVISION|NEWSLETTER|GACETA|BLOG|ESCUELA|CENTRO|OBSERVATORIO|PREMIO|PROGRAMA|",
  "CURSO|JORNADA|CONGRESO|ESPANA|NACIONAL|INTERNACIONAL|MOVIMIENTO|INICIO|CONTACTO|",
  "PRENSA|AGENDA|BOLETIN|TIENDA|AVISO|LEGAL|PRIVACIDAD|COOKIES|DISCURSO|ENTREVISTA|",
  "ARTICULO|CONFERENCIA|PONENCIA|VIDEO|FOUNDATION|CAMPAIGN|MANAGER|IP ADDRESS|",
  "SOBRE NOSOTROS|^LEER|^VER |^FISCAL|ENGLISH|ESPANOL|FRANCAIS|DEUTSCH|",
  "ISSEP|DISENSO|CITIZENGO|HAZTEOIR|DENAES")

# what may follow a name and be kept. The column is "role or description" on
# purpose: these sites print a post ("Presidente") and a profession
# ("Filosofo", "Pediatra") in exactly the same place
ROLES <- str_c(
  "PRESIDENT|VICEPRESIDENT|DIRECTOR|SECRETARI|TESORER|VOCAL|PATRON|COORDINADOR|",
  "RESPONSABLE|GERENTE|PORTAVOZ|FUNDADOR|MIEMBRO|ASESOR|PROFESOR|CATEDRATIC|",
  "DOCENTE|INVESTIGADOR|ANALISTA|CONSEJER|DELEGAD|ESCRITOR|PERIODISTA|FILOSOF|",
  "HISTORIADOR|ABOGAD|ECONOMISTA|MILITAR|EMPRESARI|MAGISTRAD|SOCIOLOG|POLITOLOG|",
  "JURISTA|DECANO|RECTOR|EDITOR|COLUMNISTA|DIPUTAD|SENADOR|EURODIPUTAD")

flat <- function(x) str_to_upper(stringi::stri_trans_general(x, "Latin-ASCII"))

# The real test of whether a line is a person: the INE census lists, which the
# project already uses to assign sex to candidates. A word list of things that
# are NOT names never ends; a list of the names that exist does.
ine_forenames <- c("Hombres", "Mujeres") |>
  map(\(s) read_excel("nombres_por_edad_media.xlsx", sheet = s, skip = 5)$Nombre) |>
  list_c() |> flat() |> str_squish()

# nicknames and rare forenames the census does not carry
ALSO_PEOPLE <- c("CAKE", "EDMALY")

# 2 to 6 words, all of them capitalised except the particles
is_person <- function(x) {
  capitalised <- map_lgl(str_split(x, " "), \(w) {
    w <- w[!str_to_lower(flat(w)) %in% PARTICLES]
    length(w) >= 2 && all(str_detect(w, "^\\p{Lu}\\p{L}"))
  })
  str_length(x) >= 5 & str_length(x) <= 60 &
    str_count(x, "\\S+") >= 2 & str_count(x, "\\S+") <= 6 &
    !str_detect(x, "[0-9@#/|]") & capitalised &
    flat(word(x, 1)) %in% c(ine_forenames, ALSO_PEOPLE) &
    !str_detect(flat(x), NOT_A_PERSON) & !str_detect(flat(x), ROLES)
}

people <- lines |>
  # "Ignacio Arsuaga Rato, Presidente" is a person and their role on one line;
  # without splitting it, the line is thrown away for containing "Presidente"
  separate_wider_delim(text, ", ", names = c("person", "inline_role"),
                       too_many = "merge", too_few = "align_start") |>
  mutate(inline_role = replace_na(inline_role, ""),
         # the other layout puts the role on the line below the name
         next_line = lead(person), .by = c(organisation, original, timestamp)) |>
  filter(is_person(person)) |>
  mutate(role = case_when(
           str_detect(flat(inline_role), ROLES) ~ inline_role,
           str_detect(flat(next_line), ROLES) & str_length(next_line) <= 90 ~ next_line,
           .default = ""),
         # Disenso prints the same board in caps on one page and in title case
         # on another: "AMANDO DE MIGUEL" and "Amando de Miguel" are one person
         person = if_else(person == str_to_upper(person), str_to_title(person), person),
         person = str_replace_all(person, str_c("(?<= )(", str_c(str_to_title(PARTICLES),
                                                                 collapse = "|"), ")(?= )"),
                                  str_to_lower))

# the same person is listed with and without the second surname ("Ana Velasco",
# "Ana Velasco Vidal-Abarca"): the longest form wins
variants <- people |>
  distinct(person) |>
  # compound surnames are opened at the hyphen, so that "Alejo Vidal Quadras"
  # falls inside "Alejo Vidal-Quadras Roca"
  mutate(tokens = str_split(flat(person), "[ -]"), first = map_chr(tokens, 1))

people <- variants |>
  rename(long = person, long_tokens = tokens) |>
  inner_join(rename(variants, short = person, short_tokens = tokens), by = "first",
             relationship = "many-to-many") |>
  filter(map2_lgl(short_tokens, long_tokens, \(s, l) all(s %in% l)),
         lengths(short_tokens) <= lengths(long_tokens)) |>
  slice_max(lengths(long_tokens), n = 1, by = short, with_ties = FALSE) |>
  select(person = short, canonical = long) |>
  right_join(people, by = "person") |>
  mutate(person = coalesce(canonical, person)) |>
  summarise(.by = c(organisation, person),
            roles      = str_c(sort(unique(role[role != ""])), collapse = " | "),
            first_seen = min(date),
            last_seen  = max(date),
            n_captures = n_distinct(timestamp),
            source     = first(str_c("https://web.archive.org/web/", timestamp, "/", original))) |>
  arrange(organisation, person)

# --- 4. Solidaridad, from the BOE ---------------------------------------------
# The union's website archives no team page, so its officers are taken from the
# four announcements of the deposit of its statutes (Direccion General de
# Trabajo, union SPDSTE, deposit number 99105954). Searched at boe.es for
# SPDSTE in the title. Rafael Martinez de la Gandara, who signs all four, is
# the civil servant issuing the resolution and is NOT part of the union.
solidaridad_boe <- tribble(
  ~person,                              ~roles,             ~date,        ~boe,
  "Santiago Alarcón Gordon",            "Solicitante del depósito", "2020-08-12", "BOE-B-2020-26040",
  "Paula Bernardina Hidalgo de la Cruz","Promotora",        "2020-08-12", "BOE-B-2020-26040",
  "Raquel Moreno Barba",                "Promotora",        "2020-08-12", "BOE-B-2020-26040",
  "Ángel García Ochoa",                 "Promotor",         "2020-08-12", "BOE-B-2020-26040",
  "Santiago Alarcón Gordon",            "Solicitante del depósito", "2021-02-15", "BOE-B-2021-7089",
  "Rodrigo Alonso Fernández",           "Secretario General", "2021-02-15", "BOE-B-2021-7089",
  "Javier Tomás de la Cruz Bazo",       "Solicitante del depósito", "2025-05-23", "BOE-B-2025-18656",
  "Jordi Albert de la Fuente Miró",     "Secretario General", "2025-05-23", "BOE-B-2025-18656",
  "Javier Tomás de la Cruz Bazo",       "Solicitante del depósito", "2026-07-03", "BOE-B-2026-22639",
  "Jordi Albert de la Fuente Miró",     "Secretario General", "2026-07-03", "BOE-B-2026-22639"
) |>
  summarise(.by = person,
            organisation = "Solidaridad",
            roles      = str_c(sort(unique(roles)), collapse = " | "),
            first_seen = min(ymd(date)),
            last_seen  = max(ymd(date)),
            n_captures = n(),
            source     = str_c("https://www.boe.es/diario_boe/txt.php?id=", first(boe)))

people <- bind_rows(people, solidaridad_boe) |> arrange(organisation, person)

people |> count(organisation, name = "people")

# who sits in more than one of them
people |>
  summarise(.by = person, organisations = str_c(sort(unique(organisation)), collapse = " | "),
            n = n_distinct(organisation)) |>
  filter(n > 1) |>
  arrange(desc(n), person)

```

```{r}
people <- people |> 
   mutate(person = person |>
           str_to_upper() |>
           stringi::stri_trans_general("Latin-ASCII") |>
           str_remove_all("[^A-Z ]") |>
           str_squish()) |> 
  filter(!person %in% c("ADOLFO SUAREZ", "YA SOMOS", "SEGUNDO APELLIDO"))
```

```{r}
# =============================================================================
# Retention of VOX top candidates across regional, Congress and European elections
#
# "Top candidate" = head of a constituency list (position 1) in Congress and
# regional elections. The European list is a single national list, so position 1
# is one person per election: its top is widened to the first TOP_N_EU names.
# =============================================================================

# "Top candidate" = the first three names of each constituency list. Where an
# election has a single constituency (European Parliament, or regions such as
# Madrid, Murcia or La Rioja), that is simply the first three of that one list.
TOP_N <- 3

vox_top <- vox_candidates |>
  filter(is_titular, !is.na(person_id)) |>
  mutate(
    arena = case_when(
      chamber == "Congress"            ~ "Congress",
      chamber == "European Parliament" ~ "European",
      level   == "Autonomical"         ~ "Autonomical"
    ),
    cycle_date = ymd(str_c(str_extract(election, "\\d{4}_\\d{2}"), "_01"))
  ) |>
  filter(!is.na(arena), list_position <= TOP_N) |>
  distinct(person_id, arena, election, cycle_date)

# every VOX candidacy of any kind (municipal and Senate included), for the
# broader question "had this person already stood for VOX before?"
vox_any <- vox_candidates |>
  filter(!is.na(person_id)) |>
  mutate(any_date = ymd(str_c(str_extract(election, "\\d{4}_\\d{2}"), "_01"))) |>
  distinct(person_id, any_date)

# ----------------------------------------------------------------- 1. RECYCLING
# of the top candidates of each election, how many had already been top
# candidates (or any candidate) in an EARLIER election
top_before <- vox_top |>
  inner_join(vox_top |> select(person_id, prev_date = cycle_date),
             by = join_by(person_id, cycle_date > prev_date)) |>
  distinct(person_id, arena, election) |>
  mutate(was_top_before = TRUE)

any_before <- vox_top |>
  inner_join(vox_any, by = join_by(person_id, cycle_date > any_date)) |>
  distinct(person_id, arena, election) |>
  mutate(was_candidate_before = TRUE)

recycling <- vox_top |>
  left_join(top_before, by = c("person_id", "arena", "election")) |>
  left_join(any_before, by = c("person_id", "arena", "election")) |>
  mutate(across(c(was_top_before, was_candidate_before), ~ coalesce(.x, FALSE))) |>
  group_by(arena, election, cycle_date) |>
  summarise(
    top_candidates        = n(),
    share_top_before      = mean(was_top_before),
    share_candidate_before = mean(was_candidate_before),
    .groups = "drop"
  ) |>
  arrange(cycle_date, arena)

# ----------------------------------------------------------------- 2. RETENTION
# of the top candidates of each election, how many are top candidates again in a
# LATER election, in any of the three arenas
top_later <- vox_top |>
  inner_join(vox_top |> select(person_id, next_arena = arena, next_date = cycle_date),
             by = join_by(person_id, cycle_date < next_date))

retention <- vox_top |>
  left_join(top_later |> distinct(person_id, arena, election) |> mutate(top_again = TRUE),
            by = c("person_id", "arena", "election")) |>
  mutate(top_again = coalesce(top_again, FALSE)) |>
  group_by(arena, election, cycle_date) |>
  summarise(top_candidates = n(), share_top_again = mean(top_again), .groups = "drop") |>
  arrange(cycle_date, arena)

# --------------------------------------------------------------- 3. TRANSITIONS
# where do top candidates go next: arena of the earlier top candidacy (rows) by
# arena of a later one (columns), counted in people
transitions <- top_later |>
  distinct(person_id, arena, next_arena) |>
  count(arena, next_arena) |>
  pivot_wider(names_from = next_arena, values_from = n, values_fill = 0)

# --------------------------------------------------------------------- 4. PLOT
recycling |>
  ggplot(aes(cycle_date, share_top_before, colour = arena)) +
  geom_line() +
  geom_point(aes(size = top_candidates)) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = NULL, y = "Top candidates already top candidates before",
       colour = NULL, size = "Top candidates") +
  theme_minimal()

recycling
retention
transitions
```

```{r IMPLANTACION TERRITORIAL VOX}
vox_candidates |>
  filter(chamber == "Congress") |>
  distinct(election, province) |>
  count(election, name = "constituencies_contested")
```
```{r}
library(mapSpain)
library(sf)

END_YEAR <- 2019   # include 2019 to see the jump after Andalusia; 2018 to stop before

# ------------------------------------------------------ province -> region lookup
prov_lookup <- provinces |>
  filter(province_code != "99") |>
  left_join(esp_codelist |> distinct(cpro, ccaa = ccaa.shortname.es),
            by = c("province_code" = "cpro")) |>
  mutate(prov_key = province |> str_to_upper() |>
           stringi::stri_trans_general("Latin-ASCII"))

# ------------------------------------ every list Vox presented, by constituency
presence <- bind_rows(
  vox_candidates |>
    filter(chamber == "Congress" | level == "Autonomical") |>
    transmute(election,
              type = if_else(chamber == "Congress", "Congress", "Regional"),
              province, unit = province),
  municipal_vox |>
    transmute(election, type = "Municipal", province, unit = chamber)
) |>
  mutate(
    date_lab = str_replace(str_extract(election, "\\d{4}_\\d{2}"), "_", "-"),
    year     = as.integer(str_sub(date_lab, 1, 4)),
    key      = province |> str_to_upper() |>
                 stringi::stri_trans_general("Latin-ASCII") |> str_squish(),
    # regional constituencies that are not provinces: islands, districts
    prov_key = case_when(
      str_detect(key, "ALAVA|ARABA")                                   ~ "ARABA/ALAVA",
      str_detect(key, "ASTURIAS") | key %in% c("CENTRO", "OCCIDENTE", "ORIENTE") ~ "ASTURIAS",
      key %in% c("MALLORCA", "MENORCA", "EIVISSA", "IBIZA", "FORMENTERA",
                 "BALEARES")                                           ~ "ILLES BALEARS",
      key %in% c("GRAN CANARIA", "LANZAROTE", "FUERTEVENTURA")         ~ "LAS PALMAS",
      key %in% c("TENERIFE", "LA PALMA", "LA GOMERA", "EL HIERRO",
                 "SANTA CRUZ DE TENERIFE")                             ~ "S.C. TENERIFE",
      key %in% c("PRIMERA", "SEGUNDA", "TERCERA", "CUARTA", "QUINTA",
                 "REGION DE MURCIA")                                   ~ "MURCIA",
      key == "COMUNIDAD DE MADRID"                                     ~ "MADRID",
      key == "VIZCAYA"   ~ "BIZKAIA",  key == "GUIPUZCOA" ~ "GIPUZKOA",
      key == "GERONA"    ~ "GIRONA",   key == "LERIDA"    ~ "LLEIDA",
      key == "ORENSE"    ~ "OURENSE",  key == "LA CORUNA" ~ "A CORUNA",
      TRUE ~ key
    )
  ) |>
  left_join(prov_lookup |> select(prov_key, province_code, province_label = province, ccaa),
            by = "prov_key")

# constituencies that could not be placed in a province (should be very few)
presence |> filter(is.na(province_code)) |> distinct(type, province)

# ----------------------------------------------------------- 1. presence matrix
lists_count <- presence |>
  filter(!is.na(province_code), year <= END_YEAR) |>
  group_by(type, date_lab, province_label) |>
  summarise(lists = n_distinct(unit), .groups = "drop")

tiles <- lists_count |>
  distinct(type, date_lab) |>
  cross_join(prov_lookup |> select(ccaa, province_label = province)) |>
  left_join(lists_count, by = c("type", "date_lab", "province_label")) |>
  mutate(type = factor(type, levels = c("Municipal", "Regional", "Congress")))



# ------------------------------------------ 2. year Vox first stood in each province
entry <- presence |>
  filter(!is.na(province_code)) |>
  group_by(type, province_code) |>
  summarise(first_year = min(year), .groups = "drop")

library(cowplot)
library(mapSpain)
library(sf)

types       <- c("Municipal", "Regional", "Congress")
year_levels <- as.character(sort(unique(entry$first_year)))
# qualitative palette: one distinct colour per year (Okabe-Ito, colour-blind safe)
fill_values <- setNames(c(palette.colors(length(year_levels) + 1, "Okabe-Ito")[-1],
                          "grey90"),
                        c(year_levels, "Never"))

maps_facet <- map(.x = types,
                  .f = function(x) esp_get_prov(epsg = "3857") %>%
                    left_join(entry %>% filter(type == x),
                              by = c("cpro" = "province_code")) %>%
                    mutate(type       = x,
                           first_year = replace_na(as.character(first_year), "Never")) %>%
                    ggplot(aes(fill = first_year)) +
                    geom_sf(colour = "white", linewidth = 0.15, show.legend = TRUE) +
                    geom_sf(data = esp_get_can_box(epsg = "3857"),
                            inherit.aes = FALSE, fill = NA, colour = "grey60") +
                    scale_fill_manual(values = fill_values,
                                      limits = names(fill_values),
                                      name = 'First Vox\ncandidacy') +
                    facet_wrap(~type) +
                    theme_minimal() +
                    theme(axis.text = element_blank(),
                          axis.ticks = element_blank(),
                          panel.grid = element_blank()))

vox_entry_maps <- plot_grid(plotlist = c(map(.x = maps_facet,
                                             .f = function(x) x + theme(legend.position = 'none')),
                                         list(get_legend(maps_facet[[1]]))),
                            nrow = 2)

vox_entry_maps
save_plot("vox_entry_maps.png", vox_entry_maps, base_width = 9, base_height = 7)

# ------------------------------------------------------- the Basque case, in figures
tiles |> filter(ccaa == "País Vasco", !is.na(lists)) |> arrange(type, date_lab)
```



```{r MAPA CONCEJALES PRE-2018}
library(mapSpain)
library(sf)

# ------------------------------------------------ elected officials before 2018
# municipal_vox has no abbreviation column, so it does not reach vox_candidates:
# take it directly and stack it with the national and regional candidacies
elected_pre2018 <- bind_rows(
  vox_candidates |>
    filter(level != "Municipal") |>
    transmute(election, year, level, chamber, province, candidate,
              is_elected = as.logical(is_elected)),
  municipal_vox |>
    transmute(election, year, level, chamber, province, candidate,
              is_elected = as.logical(is_elected))
) |>
  filter(is_elected, year < 2018)

# check: before 2018 everything should be municipal
elected_pre2018 |> count(level, election)

# ------------------------------------------------- municipality -> INE code
# matched on name AND province, so that names repeated across provinces
# (Mieres, Torrent, Sada...) go to the right municipality
muni_codes <- list()
for (s in excel_sheets("26codmun.xlsx")) {
  muni_codes[[s]] <- read_excel("26codmun.xlsx", sheet = s, skip = 2,
                                col_types = "text") |>
    filter(!is.na(CMUN)) |>
    transmute(province_code = CPRO, LAU_CODE = str_c(CPRO, CMUN),
              municipality = NOMBRE)
}
muni_codes <- bind_rows(muni_codes) |>
  left_join(provinces, by = "province_code") |>
  mutate(muni_key = municipality |>
           str_to_upper() |>
           stringi::stri_trans_general("Latin-ASCII") |>
           str_split("/") |>
           map_chr(~ str_c(sort(str_squish(.x)), collapse = "/"))) |>
  select(LAU_CODE, muni_key, province)

councillors <- elected_pre2018 |>
  filter(level == "Municipal") |>
  mutate(muni_key = chamber |>
           str_to_upper() |>
           stringi::stri_trans_general("Latin-ASCII") |>
           str_split("/") |>
           map_chr(~ str_c(sort(str_squish(.x)), collapse = "/"))) |>
  left_join(muni_codes, by = c("muni_key", "province")) |>
  count(LAU_CODE, municipality = chamber, province, name = "councillors")

# any municipality that could not be placed on the map
councillors |> filter(is.na(LAU_CODE))

map_points <- esp_get_munic(epsg = "3857") |>
  mutate(LAU_CODE = str_c(cpro, cmun)) |>
  inner_join(councillors |> select(LAU_CODE, municipality, councillors),
             by = "LAU_CODE") |>
  st_set_agr("constant") |>          # silences the "attributes constant" warning
  st_point_on_surface()

# plain X/Y columns for the labels, whatever the geometry column is called
map_labels <- map_points |>
  bind_cols(as_tibble(st_coordinates(map_points))) |>
  st_drop_geometry()

ggplot() +
  geom_sf(data = esp_get_prov(epsg = "3857"),
          fill = "grey96", colour = "grey75", linewidth = 0.2) +
  geom_sf(data = esp_get_can_box(epsg = "3857"), colour = "grey60") +
  geom_sf(data = map_points, aes(size = councillors),
          colour = "#5BC035", alpha = 0.85) +
  ggrepel::geom_text_repel(
    data = map_labels,
    aes(x = X, y = Y, label = municipality),
    size = 2.6, min.segment.length = 0
  ) +
  scale_size_area(max_size = 6, breaks = c(1, 2, 3)) +
  labs(title    = "Vox elected officials before 2018",
       subtitle = "Municipal councillors elected in the May 2015 local elections",
       size     = "Councillors",
       caption  = "Source: Ministerio del Interior; boundaries: IGN / mapSpain") +
  theme_void() +
  theme(legend.position = "bottom")

```

```{r}
# ----------------------------------------------- sex, imputed where it is missing
# municipal_vox carries no sex column, and several general elections (2016 among
# them) have the sex field empty in the ministry's files: fill those from the INE
# forename tables, assigning a sex only when one is at least 15 times more common
ine_sex <- bind_rows(
  read_excel("nombres_por_edad_media.xlsx", sheet = "Hombres", skip = 5,
             col_types = "text") |> select(token = 2, n = 3) |> mutate(sex_ine = "Male"),
  read_excel("nombres_por_edad_media.xlsx", sheet = "Mujeres", skip = 5,
             col_types = "text") |> select(token = 2, n = 3) |> mutate(sex_ine = "Female")
) |>
  mutate(token = str_to_upper(stringi::stri_trans_general(token, "Latin-ASCII")),
         n = as.numeric(n)) |>
  filter(!is.na(token), !is.na(n)) |>
  pivot_wider(names_from = sex_ine, values_from = n, values_fn = sum, values_fill = 0) |>
  mutate(sex_ine = case_when(Male   >= 15 * Female ~ "Male",
                             Female >= 15 * Male   ~ "Female")) |>
  select(token, sex_ine)

women_data <- bind_rows(
  vox_candidates |>
    filter(level != "Municipal") |>
    transmute(election, year, candidate, sex, is_titular,
              is_elected = as.logical(is_elected),
              type = case_when(chamber == "European Parliament" ~ "European",
                               chamber == "Congress"            ~ "Congress",
                               chamber == "Senate"              ~ "Senate",
                               level   == "Autonomical"         ~ "Regional")),
  municipal_vox |>
    transmute(election, year, candidate, sex = NA_character_, is_titular = TRUE,
              is_elected = as.logical(is_elected), type = "Municipal")
) |>
  mutate(clean = candidate |> str_to_upper() |>
           stringi::stri_trans_general("Latin-ASCII") |>
           str_remove_all("[^A-Z ]") |> str_squish(),
         tok2 = word(clean, 1, 2),     # compound forenames: MARIA CARMEN
         tok1 = word(clean, 1)) |>
  left_join(rename(ine_sex, sex_2 = sex_ine), by = c("tok2" = "token")) |>
  left_join(rename(ine_sex, sex_1 = sex_ine), by = c("tok1" = "token")) |>
  mutate(sex_source = if_else(is.na(sex), "INE", "official"),
         sex        = coalesce(sex, sex_2, sex_1)) |>
  filter(is_titular) |>
  select(election, year, type, candidate, sex, sex_source, is_elected)

# how much had to be imputed, and how much is still unknown
women_data |> count(type, sex_source, known = !is.na(sex))

# -------------------------------------------------------- W1, W2: overall
women_data |>
  filter(!is.na(sex)) |>
  summarise(candidates       = n(),
            women_candidates = mean(sex == "Female"),                    # W1
            elected          = sum(is_elected, na.rm = TRUE),
            women_elected    = mean(sex[is_elected %in% TRUE] == "Female")) # W2

# ------------------------------------------------ W3, W4: by type and election
women_gap <- women_data |>
  filter(!is.na(sex)) |>
  group_by(type, election, year) |>
  summarise(candidates       = n(),
            elected          = sum(is_elected, na.rm = TRUE),
            women_candidates = mean(sex == "Female"),
            women_elected    = if_else(elected > 0,
                                       mean(sex[is_elected %in% TRUE] == "Female"),
                                       NA_real_),
            gap              = women_candidates - women_elected,
            .groups = "drop") |>
  arrange(year)

women_gap |> filter(elected >= 10) |> arrange(desc(gap))                # W3, W4

# ------------------------------------------------------------------- figure
women_gap |>
  filter(elected >= 10) |>
  ggplot(aes(y = reorder(election, year))) +
  geom_segment(aes(x = women_elected, xend = women_candidates, yend = election),
               colour = "grey70") +
  geom_point(aes(x = women_candidates, colour = "All candidates"), size = 2.5) +
  geom_point(aes(x = women_elected,    colour = "Elected"),        size = 2.5) +
  geom_vline(xintercept = 0.5, linetype = "dashed", colour = "grey60") +
  scale_x_continuous(labels = scales::percent, limits = c(0, 0.6)) +
  scale_colour_manual(values = c("All candidates" = "#E69F00", "Elected" = "#0072B2")) +
  facet_grid(type ~ ., scales = "free_y", space = "free_y") +
  labs(x = "Share of women", y = NULL, colour = NULL,
       title = "Women among Vox candidates and elected officials") +
  theme_minimal() +
  theme(legend.position = "bottom")
```



```{r}
library(tidyverse)
library(ggbump)

# ---- data: every VOX candidacy ----------------------------------------------

founders <- vox_candidates |>
  filter(chamber == "European Parliament", year == 2014) |>
  arrange(list_position) |>
  distinct(person_id, .keep_all = TRUE) |>
  transmute(person_id, candidate = str_to_title(candidate), pos_2014 = list_position)

best_later <- bind_rows(
  vox_candidates |> filter(chamber == "Congress", is_titular) |> mutate(stop = "Congress"),
  vox_candidates |> filter(level == "Autonomical")            |> mutate(stop = "Regional"),
  vox_candidates |> filter(level == "Municipal")              |> mutate(stop = "Municipal")
) |>
  filter(person_id %in% founders$person_id) |>
  group_by(person_id, stop) |>
  summarise(best = min(list_position), .groups = "drop")

# ---- layout -----------------------------------------------------------------
stops   <- c("2014 European list", "Congress", "Regional", "Municipal")
stayers <- unique(best_later$person_id)
n_total <- nrow(founders)
n_stay  <- length(stayers)

stop1 <- founders |>
  transmute(person_id, candidate, stop = stops[1], best = pos_2014, y = pos_2014)

# later stops: only who stood there, ranked by best position (ties: 2014 position)
later <- best_later |>
  left_join(founders, by = "person_id") |>
  group_by(stop) |>
  arrange(best, pos_2014, .by_group = TRUE) |>
  mutate(y = row_number() * 2 - 1) |>
  ungroup() |>
  select(-pos_2014)

heads <- best_later |> filter(best == 1) |> distinct(person_id) |> pull(person_id)

pal <- setNames(c("#D55E00", "#0072B2", "#CC79A7", "#E69F00", "#009E73", "#56B4E9",
                  "#882255", "#44AA99", "#999933", "#AA4499")[seq_along(heads)], heads)

bump <- bind_rows(stop1, later) |>
  mutate(x      = match(stop, stops),
         stayed = person_id %in% stayers,
         col    = if_else(person_id %in% heads, person_id, "other"))

lines     <- filter(bump, stayed)
last_stop <- lines |> group_by(person_id) |> filter(x == max(x)) |> ungroup()

# ---- plot -------------------------------------------------------------------
ggplot(lines, aes(x, y, group = person_id)) +
  geom_bump(data = filter(lines, col == "other"), colour = "grey80", linewidth = 0.7, smooth = 8) +
  geom_bump(data = filter(lines, col != "other"), aes(colour = col), linewidth = 1.3, smooth = 8) +
  geom_point(data = filter(bump, !stayed), colour = "grey88", size = 1.8) +
  geom_point(data = filter(lines, x == 1, col == "other"), colour = "grey65", size = 2) +
  geom_point(data = filter(lines, x == 1, col != "other"), aes(colour = col), size = 2.8) +
  geom_point(data = filter(lines, x > 1, col == "other"), colour = "grey65", size = 5.5) +
  geom_point(data = filter(lines, x > 1, col != "other"), aes(colour = col), size = 5.5) +
  geom_text(data = filter(lines, x > 1), aes(label = best),
            colour = "white", size = 2.3, fontface = "bold") +
  geom_text(data = filter(bump, x == 1),
            aes(x = 0.93, label = str_c(best, ". ", candidate),
                colour = if_else(stayed, col, "gone")),
            hjust = 1, size = 2.5) +
  geom_label(data = last_stop,
             aes(x = x + 0.08, label = candidate, colour = if_else(col == "other", "lab", col)),
             hjust = 0, size = 2.5, fill = "white", label.size = 0,
             label.padding = unit(0.1, "lines")) +
  annotate("text", x = 0.93, y = n_total + 2.5, hjust = 1, vjust = 1, size = 3, colour = "grey45",
           label = str_c(n_total - n_stay, " of the ", n_total,
                         " founders (light grey) never stood for Vox again")) +
  scale_colour_manual(values = c(pal, other = "grey45", gone = "grey75", lab = "grey35")) +
  scale_y_reverse(expand = expansion(add = c(3, 1))) +
  scale_x_continuous(breaks = 1:4, position = "top",
                     labels = c("2014 European\nlist", "Congress\n(best list position)",
                                "Regional\n(best list position)", "Municipal\n(best list position)"),
                     expand = expansion(add = c(1.6, 1.3))) +
  labs(title = "Founders: who is left from the first 2014 election",
       subtitle = "Each line only passes through the types of election the founder later stood in. Number = best list position. Coloured: founders who later headed a list.",
       x = NULL, y = NULL) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "none",
        panel.grid = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_text(face = "bold", colour = "grey25"),
        plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(colour = "grey40", size = 9))

ggsave("founders_bump.png", width = 13, height = 11, dpi = 300, bg = "white")
```
```{r}
vox_candidates |> 
  filter(election == "general_2019_04") |> 
  filter(is_elected == TRUE) |> 
  distinct(person_id) |> 
  count()

vox_candidates |> 
  filter(election == "general_2019_11") |> 
  filter(is_elected == TRUE) |> 
  distinct(person_id) |> 
  count()

vox_candidates |> 
  filter(election == "municipales_2019_05") |> 
  filter(is_elected == TRUE) |> 
  distinct(person_id) |> 
  count()

vox_candidates |> 
  filter(election == "general_2023_07") |> 
  filter(is_elected == TRUE) |> 
  distinct(person_id) |> 
  count()

vox_candidates |> 
  filter(election == "municipales_2023_05") |> 
  filter(is_elected == TRUE) |> 
  distinct(person_id) |> 
  count()

vox_candidates |> 
  filter(election == "autonomous_2019_04" | election == "autonomous_2019_05") |> 
  filter(is_elected == TRUE) |> 
  distinct(person_id) |> 
  count()

vox_candidates |> 
  filter(is_elected == TRUE) |> 
  filter(str_detect(election, "autonomous")) |> 
  filter(election != "autonomous_2019_04" | election != "autonomous_2019_05" | election != "autonomous_2018_12" | election != "autonomous_2019_05") |> 
  
  distinct(person_id) |> 
  count()
```


```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

# 1. Una fila por persona-elección, con el tramo de lista
gen <- vox_candidates |>
  filter(election == "Congress", !is.na(person_id), !is.na(list_position)) |>
  distinct(person_id, year, province, list_position) |>
  mutate(
    tier = case_when(
      list_position == 1 ~ "Head of list",
      list_position <= 3 ~ "Positions 2-3",
      TRUE               ~ "Rest of list"
    ),
    tier = factor(tier, levels = c("Rest of list", "Positions 2-3", "Head of list"))
  )

# 2. Pares de convocatorias consecutivas
years <- sort(unique(gen$year))
pairs <- tibble(year_prev = head(years, -1), year_next = tail(years, -1))

# 3. Para cada cohorte en t-1, ¿quién vuelve en t y en qué puesto?
retention_tier <- pairs |>
  inner_join(gen, by = c("year_prev" = "year"), relationship = "many-to-many") |>
  left_join(
    gen |> select(person_id, year, pos_next = list_position, prov_next = province),
    by = c("person_id", "year_next" = "year")
  ) |>
  summarise(
    n_cohort        = n(),
    retention       = mean(!is.na(pos_next)),
    stayed_head     = mean(pos_next == 1, na.rm = TRUE),
    changed_prov    = mean(prov_next != province, na.rm = TRUE),
    .by = c(year_prev, year_next, tier)
  ) |>
  arrange(year_prev, desc(tier))

retention_tier
```

```{r}
vox_candidates |> head(2) |> 
  print()
```

```{r NETWORK ANALYSIS}
vox_candidates <- vox_candidates |>
  mutate(
    from_cs = person_id %in% cs_candidates$person_id[!is.na(cs_candidates$person_id)],
    from_radical_right = person_id %in% radical_right_candidates$person_id[!is.na(radical_right_candidates$person_id)])

vox_candidates |>
  group_by(from_pp, from_cs, from_radical_right) |>
  summarise(n = n(),
            mean_list_position = mean(list_position, na.rm = TRUE),
            median_list_position = median(list_position, na.rm = TRUE),
            .groups = "drop") |>
  arrange(desc(n))

# one row per person and party they ran for: the edge list of the network
affiliations <- bind_rows(
  vox_candidates |> distinct(person_id) |> mutate(party = "VOX"),
  pp_candidates |> distinct(person_id) |> mutate(party = "PP"),
  cs_candidates |> distinct(person_id) |> mutate(party = "Cs"),
  radical_right_candidates |> distinct(person_id, Partido) |> rename(party = Partido)) |>
  filter(!is.na(person_id), person_id != "")

# people who ran for more than one party
movers <- affiliations |>
  group_by(person_id) |>
  filter(n_distinct(party) > 1) |>
  ungroup()

# party-to-party edges, weighted by how many people they share
party_edges <- movers |>
  inner_join(movers, by = "person_id", relationship = "many-to-many") |>
  filter(party.x < party.y) |>
  count(party.x, party.y, name = "weight") |>
  arrange(desc(weight))

party_edges |> filter(party.y == "VOX")
```

```{r PP AND VOX NAME VERIFICATOR ONLY SAME_PROVINCE}

# The overlap between PP and VOX is matched on the normalised name, so common
# names produce false positives. The INE publishes forename and surname
# frequencies: assuming independence, the expected number of people in Spain
# carrying that exact combination gives a triage order for manual review.
# The file is https://www.ine.es/daco/daco42/nombyapel/apellidos_frecuencia.xls
SPAIN <- 47.4e6

surnames <- read_excel("apellidos_frecuencia.xls",
                       sheet = "Apellidos >=100", skip = 4,
                       col_types = "text") |>          # the file uses ".." for
  select(token = 2, n = 3) |>                          # suppressed values
  transmute(token = str_to_upper(stringi::stri_trans_general(token, "Latin-ASCII")),
            n_surname = as.numeric(n)) |>
  filter(!is.na(token), !is.na(n_surname)) |>
  group_by(token) |> summarise(n_surname = max(n_surname), .groups = "drop")

forenames <- bind_rows(
  read_excel("nombres_por_edad_media.xlsx", sheet = "Hombres", skip = 5,
             col_types = "text") |> select(token = 2, n = 3),
  read_excel("nombres_por_edad_media.xlsx", sheet = "Mujeres", skip = 5,
             col_types = "text") |> select(token = 2, n = 3)) |>
  transmute(token = str_to_upper(stringi::stri_trans_general(token, "Latin-ASCII")),
            n_forename = as.numeric(n)) |>
  filter(!is.na(token), !is.na(n_forename)) |>
  group_by(token) |> summarise(n_forename = sum(n_forename), .groups = "drop")

# one row per party with where and when the person stood, so the context itself
# tells you whether it is plausibly the same individual
pp_context <- pp_candidates |>
  filter(!is.na(person_id)) |>
  group_by(person_id) |>
  summarise(pp_years  = str_c(sort(unique(year)), collapse = ", "),
            pp_places = str_c(sort(unique(province)), collapse = ", "),
            .groups = "drop")

vox_context <- vox_candidates |>
  filter(!is.na(person_id)) |>
  group_by(person_id) |>
  summarise(vox_years  = str_c(sort(unique(year)), collapse = ", "),
            vox_places = str_c(sort(unique(province)), collapse = ", "),
            .groups = "drop")

homonym_review <- inner_join(pp_context, vox_context, by = "person_id") |>
  mutate(tok = str_split(person_id, " "),
         surname_2 = map_chr(tok, ~ .x[length(.x)]),
         surname_1 = map_chr(tok, ~ .x[length(.x) - 1]),
         forename  = map_chr(tok, ~ str_c(.x[1:(length(.x) - 2)], collapse = " "))) |>
  select(-tok) |>
  left_join(forenames, by = c("forename"  = "token")) |>
  left_join(rename(surnames, n_surname_1 = n_surname), by = c("surname_1" = "token")) |>
  left_join(rename(surnames, n_surname_2 = n_surname), by = c("surname_2" = "token")) |>
  mutate(
    expected_bearers = round(n_forename * n_surname_1 * n_surname_2 / SPAIN^2, 2),
    same_province    = pp_places == vox_places,
    verdict          = NA_character_          # fill in by hand
  ) |>
  arrange(desc(expected_bearers)) |>
  select(person_id, expected_bearers, same_province,
         pp_years, pp_places, vox_years, vox_places, verdict) |> 
  filter(same_province == TRUE)

# write_csv(homonym_review, "homonym_review.csv")
# homonym_review
```

```{r NEWCOMER VARIABLE}
vox_candidates <- vox_candidates |> 
   group_by(person_id) |>
  mutate(
    appearance    = dense_rank(election_date),   # not row_number: someone can
    is_newcomer   = appearance == 1,             # run for Congress + Senate
    prev_position = lag(list_position, order_by = election_date),
    prev_elected  = lag(is_elected,    order_by = election_date)
  ) |>
  ungroup()
```

```{r}

```

```{r ADD WHO WAS IN PP OR OTHER PARTIES BEFORE (NOT NEWCOMER)}

```

```{r}
library(readxl)

# --- 1. leer los dos ficheros de resultados provinciales ------------------
results <- list()

for (f in c("PROV_02_201606_1.xlsx", "PROV_02_201904_1.xlsx")) {

  raw <- read_excel(f, sheet = "Circunscripciones", col_names = FALSE)

  type_i  <- which(rowSums(raw == "Votos", na.rm = TRUE) > 0)[1]
  party_i <- type_i - 1

  party_row <- str_squish(as.character(unlist(raw[party_i, ])))
  type_row  <- str_squish(as.character(unlist(raw[type_i, ])))
  body      <- raw[(type_i + 1):nrow(raw), ]

  votes_cols <- which(type_row == "Votos"     & !is.na(party_row))
  seats_cols <- which(type_row == "Diputados" & !is.na(party_row))

  votes_long <- bind_cols(
      tibble(province = str_squish(as.character(body[[3]]))),
      set_names(body[votes_cols], party_row[votes_cols])
    ) |>
    pivot_longer(-province, names_to = "abbreviation", values_to = "votes")

  seats_long <- bind_cols(
      tibble(province = str_squish(as.character(body[[3]]))),
      set_names(body[seats_cols], party_row[seats_cols])
    ) |>
    pivot_longer(-province, names_to = "abbreviation", values_to = "seats")

  results[[f]] <- left_join(votes_long, seats_long,
                            by = c("province", "abbreviation")) |>
    mutate(
      votes    = as.numeric(votes),
      seats    = as.numeric(seats),
      province = province |>
        stringi::stri_trans_general("Latin-ASCII") |>
        str_squish(),
      base_election = if_else(str_detect(f, "201606"),
                              "general_2016_06", "general_2019_04")
    ) |>
    filter(votes > 0)
}

results <- bind_rows(results)

# --- 2. magnitud de cada circunscripcion ---------------------------------
magnitude <- results |>
  filter(base_election == "general_2019_04") |>
  group_by(province) |>
  summarise(magnitude = sum(seats), .groups = "drop")

# --- 3. escenarios de voto nacional de VOX -------------------------------
# marzo 2019: la referencia es 2016 (VOX partia de casi cero)
# octubre 2019: la referencia es abril 2019 (VOX ya tenia 10,3% real)
scenarios <- bind_rows(
  tibble(moment = "march_2019",   base_election = "general_2016_06",
         vox_nat = c(0.08, 0.10, 0.12, 0.14)),
  tibble(moment = "october_2019", base_election = "general_2019_04",
         vox_nat = c(0.08, 0.10, 0.12, 0.14))
)

# --- 4. proyeccion: swing uniforme, renormalizando el resto --------------
others <- results |>
  filter(abbreviation != "VOX") |>
  group_by(base_election, province) |>
  mutate(share_others = votes / sum(votes)) |>
  ungroup() |>
  select(base_election, province, abbreviation, share_others)

expected <- bind_rows(
  left_join(others, scenarios, by = "base_election",
            relationship = "many-to-many") |>
    mutate(share_exp = share_others * (1 - vox_nat)),
  left_join(distinct(others, base_election, province), scenarios,
            by = "base_election", relationship = "many-to-many") |>
    mutate(abbreviation = "VOX", share_exp = vox_nat)
)

# --- 5. D'Hondt -----------------------------------------------------------
seats_exp <- expected |>
  left_join(magnitude, by = "province") |>
  filter(share_exp >= 0.03) |>
  mutate(divisor = map(magnitude, seq_len)) |>
  unnest(divisor) |>
  mutate(quotient = share_exp / divisor) |>
  group_by(moment, vox_nat, province) |>
  arrange(desc(quotient), .by_group = TRUE) |>
  filter(row_number() <= magnitude) |>
  count(moment, vox_nat, province, abbreviation, name = "seats") |>
  ungroup()

vox_expected <- seats_exp |>
  filter(abbreviation == "VOX") |>
  select(-abbreviation) |>
  pivot_wider(names_from = vox_nat, values_from = seats, values_fill = 0) |>
  arrange(moment, province)

vox_expected

# --- 6. comprobaciones ----------------------------------------------------
count(results, base_election)
n_distinct(results$province)          # deben ser 52
sum(magnitude$magnitude)              # debe dar 350
anti_join(magnitude, distinct(all_candidates, province), by = "province")
```

```{r}
library(readxl)

# --- 1. leer los dos ficheros de resultados provinciales ------------------
results <- list()

for (f in c("PROV_02_201606_1.xlsx", "PROV_02_201904_1.xlsx")) {

  m <- as.matrix(read_excel(f, sheet = "Circunscripciones",
                            col_names = FALSE, .name_repair = "minimal"))
  mode(m) <- "character"

  hits    <- which(m == "Votos", arr.ind = TRUE)
  type_i  <- min(hits[, "row"])
  party_i <- type_i - 1

  party_row <- str_squish(m[party_i, ])
  type_row  <- str_squish(m[type_i, ])

  votes_cols <- which(type_row == "Votos" & !is.na(party_row))
  seats_cols <- votes_cols + 1

  stopifnot(length(votes_cols) > 0,
            all(type_row[seats_cols] == "Diputados"))

  body_rows <- (type_i + 1):nrow(m)

  province <- m[body_rows, 3] |>
    stringi::stri_trans_general("Latin-ASCII") |>
    str_squish() |>
    str_replace("^Araba.*", "Araba")

  votes_m <- m[body_rows, votes_cols, drop = FALSE]
  seats_m <- m[body_rows, seats_cols, drop = FALSE]
  colnames(votes_m) <- party_row[votes_cols]
  colnames(seats_m) <- party_row[votes_cols]

  votes_long <- as_tibble(votes_m, .name_repair = "unique") |>
    mutate(province = province) |>
    pivot_longer(-province, names_to = "abbreviation", values_to = "votes")

  seats_long <- as_tibble(seats_m, .name_repair = "unique") |>
    mutate(province = province) |>
    pivot_longer(-province, names_to = "abbreviation", values_to = "seats")

  results[[f]] <- left_join(votes_long, seats_long,
                            by = c("province", "abbreviation")) |>
    mutate(
      votes = as.numeric(votes),
      seats = as.numeric(seats),
      base_election = if_else(str_detect(f, "201606"),
                              "general_2016_06", "general_2019_04")
    ) |>
    filter(!is.na(province), province != "", !is.na(votes), votes > 0) |>
      filter(!is.na(province), province != "", province != "Total",
           !is.na(votes), votes > 0)
}

results <- bind_rows(results)

stopifnot(n_distinct(results$province) == 52)

# --- 2. magnitud de cada circunscripcion ---------------------------------
magnitude <- results |>
  filter(base_election == "general_2019_04") |>
  group_by(province) |>
  summarise(magnitude = sum(seats, na.rm = TRUE), .groups = "drop")

stopifnot(sum(magnitude$magnitude) == 350)

# --- 3. escenarios de voto nacional de VOX -------------------------------
scenarios <- bind_rows(
  tibble(moment = "march_2019",   base_election = "general_2016_06",
         vox_nat = c(0.08, 0.10, 0.12, 0.14)),
  tibble(moment = "october_2019", base_election = "general_2019_04",
         vox_nat = c(0.08, 0.10, 0.12, 0.14))
)

# --- 4. proyeccion: swing uniforme, renormalizando el resto --------------
others <- results |>
  filter(abbreviation != "VOX") |>
  group_by(base_election, province) |>
  mutate(share_others = votes / sum(votes)) |>
  ungroup() |>
  select(base_election, province, abbreviation, share_others)

expected <- bind_rows(
  left_join(others, scenarios, by = "base_election",
            relationship = "many-to-many") |>
    mutate(share_exp = share_others * (1 - vox_nat)),
  left_join(distinct(others, base_election, province), scenarios,
            by = "base_election", relationship = "many-to-many") |>
    mutate(abbreviation = "VOX", share_exp = vox_nat)
) |>
  left_join(magnitude, by = "province")

stopifnot(!any(is.na(expected$magnitude)))

# --- 5. D'Hondt -----------------------------------------------------------
seats_exp <- expected |>
  filter(share_exp >= 0.03) |>
  mutate(divisor = map(magnitude, seq_len)) |>
  unnest(divisor) |>
  mutate(quotient = share_exp / divisor) |>
  group_by(moment, vox_nat, province) |>
  arrange(desc(quotient), .by_group = TRUE) |>
  filter(row_number() <= magnitude) |>
  count(moment, vox_nat, province, abbreviation, name = "seats") |>
  ungroup()

vox_expected <- seats_exp |>
  filter(abbreviation == "VOX") |>
  select(-abbreviation) |>
  pivot_wider(names_from = vox_nat, values_from = seats, values_fill = 0) |>
  arrange(moment, province)

vox_expected |> 
  arrange(province)
```

```{r}
results |>
  filter(base_election == "general_2019_04") |>
  distinct(province) |>
  anti_join(filter(results, base_election == "general_2016_06") |> distinct(province),
            by = "province")
```

```{r}
# --- expectativa media de escaños por provincia y momento -----------------
expectation <- seats_exp |>
  filter(abbreviation == "VOX") |>
  group_by(moment, province) |>
  summarise(exp_seats = mean(seats), .groups = "drop") |>
  bind_rows(                                   # provincias con 0 en todos los escenarios
    crossing(moment = c("march_2019", "october_2019"),
             province = magnitude$province) |>
      anti_join(filter(seats_exp, abbreviation == "VOX"),
                by = c("moment", "province")) |>
      mutate(exp_seats = 0)
  ) |>
  pivot_wider(names_from = moment, values_from = exp_seats,
              names_prefix = "exp_") |>
  mutate(exp_gain = exp_october_2019 - exp_march_2019)

# --- listas de abril y noviembre, solo Congreso y titulares ---------------
lists_2019 <- vox_candidates |>
  filter(chamber == "Congress", is_titular,
         election %in% c("general_2019_04", "general_2019_11")) |>
  mutate(province = str_replace(province, "^Araba.*", "Araba"))

# --- rotación por provincia ----------------------------------------------
churn <- lists_2019 |>
  group_by(province, person_id) |>
  summarise(in_april = any(election == "general_2019_04"),
            in_nov   = any(election == "general_2019_11"),
            .groups  = "drop_last") |>
  summarise(
    n_april   = sum(in_april),
    n_nov     = sum(in_nov),
    stayed    = sum(in_april & in_nov),
    dropped   = sum(in_april & !in_nov),
    added     = sum(!in_april & in_nov),
    churn_rate = dropped / n_april,
    .groups   = "drop"
  ) |>
  left_join(expectation, by = "province")

churn |> arrange(desc(exp_october_2019))

# --- total de nombres que cambiaron --------------------------------------
churn |> summarise(across(c(n_april, stayed, dropped, added), sum))

# --- rotación solo en los puestos
```

```{r}
# lo que sabían en octubre: el resultado REAL de abril, no una proyección
seats_april <- vox_candidates |>
  filter(election == "general_2019_04", chamber == "Congress",
         is_titular, is_elected) |>
  mutate(province = str_replace(province, "^Araba.*", "Araba")) |>
  count(province, name = "seats_april")

bands <- expectation |>
  left_join(seats_april, by = "province") |>
  mutate(seats_april = replace_na(seats_april, 0))

# ¿quién es nuevo en la lista de noviembre?
november <- vox_candidates |>
  filter(election == "general_2019_11", chamber == "Congress", is_titular) |>
  mutate(province = str_replace(province, "^Araba.*", "Araba"))

april_ids <- vox_candidates |>
  filter(election == "general_2019_04", chamber == "Congress", is_titular) |>
  mutate(province = str_replace(province, "^Araba.*", "Araba")) |>
  distinct(province, person_id)

test <- november |>
  left_join(bands, by = "province") |>
  left_join(mutate(april_ids, was_in_april = TRUE),
            by = c("province", "person_id")) |>
  mutate(
    is_new = is.na(was_in_april),
    band = case_when(
      list_position <= seats_april       ~ "1_ya ganable en abril",
      list_position <= exp_october_2019  ~ "2_nuevo ganable",
      TRUE                               ~ "3_no ganable"
    )
  )

test |>
  group_by(band) |>
  summarise(n = n(), share_new = mean(is_new))

# solo en provincias donde la expectativa creció
test |>
  filter(exp_october_2019 > seats_april) |>
  group_by(band) |>
  summarise(n = n(), share_new = mean(is_new))

# posición exacta donde entran los nuevos, respecto al corte de abril
test |>
  mutate(dist_cutoff = list_position - seats_april) |>
  filter(dist_cutoff >= -3, dist_cutoff <= 6) |>
  group_by(dist_cutoff) |>
  summarise(n = n(), share_new = mean(is_new))
```

```{r}
test |>
  filter(!is_new, exp_october_2019 > seats_april) |>
  left_join(
    vox_candidates |>
      filter(election == "general_2019_04", chamber == "Congress", is_titular) |>
      mutate(province = str_replace(province, "^Araba.*", "Araba")) |>
      select(province, person_id, pos_april = list_position),
    by = c("province", "person_id")
  ) |>
  mutate(move = pos_april - list_position) |>
  group_by(band) |>
  summarise(n = n(), mean_move = mean(move, na.rm = TRUE),
            promoted = mean(move > 0, na.rm = TRUE))
```

```{r}
test |>
  group_by(province) |>
  summarise(identica = all(!is_new), n = n()) |>
  count(identica)
```

```{r}
test |>
  group_by(province) |>
  summarise(first_new = min(list_position[is_new]),
            seats_april = first(seats_april),
            .groups = "drop") |>
  count(first_new > seats_april)
```

```{r}
test |>
  group_by(province) |>
  summarise(first_new = min(list_position[is_new]),
            seats_april = first(seats_april), .groups = "drop") |>
  filter(is.na(first_new) | is.na(seats_april) | first_new <= seats_april) |>
  left_join(select(test, province, list_position, candidate, is_new),
            by = "province") |>
  filter(is_new) |>
  arrange(province, list_position) |> 
  view()
  
```

```{r}

```
