---
title: "VOX_CANDIDATES"
author: "Noah Emanuel Fürup"
format: pdf
editor: visual
---

```{r LIBRARIES}
library(tidyverse)
```

Extraction of non-autonomical elections candidates: european and general elections

Source: - 2014-2019: github.com/JaimeObregon/infoelectoral - 2023-2024: infoelectoral.interior.gob.es

```{r SPANISH LETERS}
latin1 <- locale(encoding = "ISO-8859-1")
```

```{r GITHUB MIRROR}
MIRROR <- "https://raw.githubusercontent.com/JaimeObregon/infoelectoral/main/files"
```

The github has the files in .DAT, a file type that doesn't separate in ";" or "\|". In order to separate each column, we need to map where the separation between each variable occurs. I asked Claude to give me the exact character positions where the separation is. I used fwf_positions of ReadR to write down the column names, given the positions where the variables lie in the .DAT file.

#GENERAL ELECTIONS

```{r FIXED POSITIONS}
# File 03 — candidacies (parties)
cols_03 <- fwf_cols(
  party_code   = c(  9,  14),
  abbreviation = c( 15,  64),
  full_name    = c( 65, 214)
)

# File 04 — candidates (people)
cols_04 <- fwf_cols(
  province_code  = c( 10,  11),
  party_code     = c( 16,  21),
  list_position  = c( 22,  24),
  candidate_type = c( 25,  25),   # T = titular, S = substitute
  first_name     = c( 26,  50),
  surname_1      = c( 51,  75),
  surname_2      = c( 76, 100),
  sex            = c(101, 101),   # M / F
  elected        = c(120, 120)    # S = elected
)

# up to 2000 the name is not split: one field, and no sex column
cols_04_single <- fwf_cols(
  province_code  = c( 10,  11),
  party_code     = c( 16,  21),
  list_position  = c( 22,  24),
  candidate_type = c( 25,  25),
  first_name     = c( 26, 100),   # name and both surnames together
  elected        = c(120, 120)
)
```

```{r ELECTIONS}
# the ministry's server is slow: R's 60-second default aborts mid-download
options(timeout = 600)

LOCAL_MIRROR <- "infoelectoral/files"
MIRROR <- "https://raw.githubusercontent.com/JaimeObregon/infoelectoral/main/files"
PORTAL <- "https://infoelectoral.interior.gob.es/estaticos/docxl/apliextr"

# From 1996 on. Congress 2000 is left out on purpose: the ministry's file for that
# one process contains only the 350 elected deputies, not the candidacies.
elections <- tribble(
  ~process, ~year,  ~month, ~chamber,              ~mirror_folder,
  "02",     "1996", "03",   "Congress",            "congreso",
  "02",     "2004", "03",   "Congress",            "congreso",
  "02",     "2008", "03",   "Congress",            "congreso",
  "02",     "2011", "11",   "Congress",            "congreso",
  "02",     "2015", "12",   "Congress",            "congreso",
  "02",     "2016", "06",   "Congress",            "congreso",
  "02",     "2019", "04",   "Congress",            "congreso",
  "02",     "2019", "11",   "Congress",            "congreso",
  "02",     "2023", "07",   "Congress",            "congreso",
  "03",     "1996", "03",   "Senate",              "senado",
  "03",     "2000", "03",   "Senate",              "senado",
  "03",     "2004", "03",   "Senate",              "senado",
  "03",     "2008", "03",   "Senate",              "senado",
  "03",     "2011", "11",   "Senate",              "senado",
  "03",     "2015", "12",   "Senate",              "senado",
  "03",     "2016", "06",   "Senate",              "senado",
  "03",     "2019", "04",   "Senate",              "senado",
  "03",     "2019", "11",   "Senate",              "senado",
  "03",     "2023", "07",   "Senate",              "senado",
  "07",     "1999", "06",   "European Parliament", "europeas",
  "07",     "2004", "06",   "European Parliament", "europeas",
  "07",     "2009", "06",   "European Parliament", "europeas",
  "07",     "2014", "05",   "European Parliament", "europeas",
  "07",     "2019", "05",   "European Parliament", "europeas",
  "07",     "2024", "06",   "European Parliament", "europeas"
)

```


```{r PROVINCES}
provinces <- tribble(
  ~province_code, ~province,
  "01", "Araba/Alava",    "02", "Albacete",      "03", "Alicante",
  "04", "Almeria",        "05", "Avila",         "06", "Badajoz",
  "07", "Illes Balears",  "08", "Barcelona",     "09", "Burgos",
  "10", "Caceres",        "11", "Cadiz",         "12", "Castellon",
  "13", "Ciudad Real",    "14", "Cordoba",       "15", "A Coruna",
  "16", "Cuenca",         "17", "Girona",        "18", "Granada",
  "19", "Guadalajara",    "20", "Gipuzkoa",      "21", "Huelva",
  "22", "Huesca",         "23", "Jaen",          "24", "Leon",
  "25", "Lleida",         "26", "La Rioja",      "27", "Lugo",
  "28", "Madrid",         "29", "Malaga",        "30", "Murcia",
  "31", "Navarra",        "32", "Ourense",       "33", "Asturias",
  "34", "Palencia",       "35", "Las Palmas",    "36", "Pontevedra",
  "37", "Salamanca",      "38", "S.C. Tenerife", "39", "Cantabria",
  "40", "Segovia",        "41", "Sevilla",       "42", "Soria",
  "43", "Tarragona",      "44", "Teruel",        "45", "Toledo",
  "46", "Valencia",       "47", "Valladolid",    "48", "Bizkaia",
  "49", "Zamora",         "50", "Zaragoza",      "51", "Ceuta",
  "52", "Melilla",        "99", "Single national constituency"
)
```

```{r CODES MUNICIPALITIES}
# 26codmun and 26codislas
```


```{r DOWNLOAD FILES GITHUB}
candidate_files <- list()

for (i in seq_len(nrow(elections))) {

  process       <- elections$process[i]
  year          <- elections$year[i]
  month         <- elections$month[i]
  chamber       <- elections$chamber[i]
  mirror_folder <- elections$mirror_folder[i]

  yy     <- str_sub(year, 3, 4)
  folder <- str_c(process, year, month, "_TOTA")

  local_dir <- str_c(LOCAL_MIRROR, "/", mirror_folder, "/", folder)

  # the GitHub mirror lacks a few processes (Congress 1989 and 1993, and
  # everything after 2019), so take those from the ministry's download area
  if (!dir.exists(local_dir)) {
    message("Fetching ", folder, " from the ministry portal ...")
    zip_file <- tempfile(fileext = ".zip")
    download.file(str_c(PORTAL, "/", folder, ".zip"), zip_file,
                  mode = "wb", quiet = TRUE)
    dir.create(local_dir, recursive = TRUE, showWarnings = FALSE)
    unzip(zip_file, exdir = local_dir)
  }

  path_03 <- list.files(local_dir, str_c("^03", process, yy, month, "[.]DAT$"),
                        full.names = TRUE, recursive = TRUE, ignore.case = TRUE)[1]
  path_04 <- list.files(local_dir, str_c("^04", process, yy, month, "[.]DAT$"),
                        full.names = TRUE, recursive = TRUE, ignore.case = TRUE)[1]

  message("Reading ", chamber, " ", year, "-", month, " ...")

  parties <- read_fwf(path_03, cols_03, locale = latin1,
                      col_types = cols(.default = "c")) |>
    mutate(across(everything(), str_trim))

  # up to 2000 the name comes in a single field; from 2004 it is split in three.
  # Decide from the data: the second-surname block is only filled in the split one.
  probe <- read_lines(path_04, n_max = 500, locale = latin1)
  split <- mean(str_squish(str_sub(probe, 76, 100)) != "") > 0.5

  raw <- read_fwf(path_04, if (split) cols_04 else cols_04_single,
                  locale = latin1, col_types = cols(.default = "c")) |>
    mutate(across(everything(), str_trim))

  # give every process the same columns, so bind_rows and transmute never break
  if (!"surname_1" %in% names(raw)) raw <- mutate(raw, surname_1 = NA_character_)
  if (!"surname_2" %in% names(raw)) raw <- mutate(raw, surname_2 = NA_character_)
  if (!"sex"       %in% names(raw)) raw <- mutate(raw, sex       = NA_character_)

  candidate_files[[i]] <- raw |>
    left_join(parties, by = "party_code") |>
    mutate(chamber = chamber, year = year, month = month)
}
```

`all_candidates` is built in ONE place only, in the chunk below. Building it in
several chunks is what silently dropped most processes before: re-running any of
them out of order overwrote the object.

```{r JOIN PROVINCES AND TRANSMUTE}
all_candidates <- bind_rows(candidate_files) |>
  left_join(provinces, by = "province_code") |>
  transmute(
    election      = str_c(if_else(chamber == "European Parliament",
                                  "european", "general"),
                          "_", year, "_", month),
    year          = as.integer(year),
    chamber,
    province,
    list_position = as.integer(list_position),
    # str_c returns NA if ANY part is NA, and the surnames are empty before 2004
    candidate     = str_squish(str_c(coalesce(first_name, ""), " ",
                                     coalesce(surname_1,  ""), " ",
                                     coalesce(surname_2,  ""))),
    sex           = recode(sex, "M" = "Male", "F" = "Female",
                           .default = NA_character_),
    is_titular    = candidate_type == "T",
    is_elected    = elected == "S",
    abbreviation,
    level         = "National"
  ) |>
  arrange(year, chamber, province, list_position)

# sanity check: every process must be present
all_candidates |> count(year, chamber) |> print(n = Inf)
all_candidates |>
  filter(str_detect(coalesce(abbreviation, ""), regex("vox", ignore_case = TRUE))) |>
  count(year, chamber) |> print(n = Inf)
```


The 2023 and 2024 processes used to be downloaded in a separate block that
rebuilt `all_candidates`. They are now rows of the `elections` tribble above and
are read by the same loop, so that block has been removed: it duplicated those
rows and, run out of order, wiped every earlier process.



```{r TAKE DON/ DOÑA OUT}
# Handled upstream in READ ALL AUTONOMICAL CANDIDATES: the honorific is stripped
# from `candidate `before the union (this chunk referred to `candidato_norm'
# a column that does not exist in autonomical_candidates).
all_candidates <- all_candidates |>
  mutate(candidate = str_squish(str_remove(candidate,
                                           regex("^(don|do[ñn]a|d[.ª])\\s+",
                                                 ignore_case = TRUE))))
```

```{r FILTER PP AND VOX}
# Anchor the abbreviations: a bare "PP|AP" also matches APROBAR, PPSO, MAP and
# anything else merely containing those letters. Coalesce first, because
# str_detect(NA) is NA and filter() drops those rows without saying so.
pp_candidates <- all_candidates |>
  filter(str_detect(coalesce(abbreviation, ""),
                    regex("^(PP|P\\.P\\.|PPC|PPdeG|PPCV|PPN|PPG)\\b|\\bPP\\b",
                          ignore_case = TRUE)))

# before 1989 there is no PP: it is Alianza Popular, and in 1982 and 1986 it ran
# in coalition (AP-PDP, Coalicion Popular, AP-PDP-PL). Kept as its own object so
# the AP years are not silently folded into "PP".
ap_candidates <- all_candidates |>
  filter(str_detect(coalesce(abbreviation, ""),
                    regex("^(AP|A\\.P\\.|AP-|CP)\\b|COALICION POPULAR",
                          ignore_case = TRUE)))

vox_candidates <- all_candidates |>
  filter(str_detect(coalesce(abbreviation, ""), regex("\\bVOX\\b", ignore_case = TRUE)))
```

#AUTONOMICAL ELECTIONS 


Claude Code For Autonomical elections: two attempts to confirm it is reproducible.

Prompt 1:

"Quiero que busques todas las listas de todas las elecciones autonómicas que han habido en España desde 2015 hasta 2026 y que pongas en el archivo excel en el excel de CANDIDATOS_AUTONOMICAS. Tienes que ir a la web para recopilar las siguientes variables de cada eleccion: CANDIDATO, CCAA, ELECCION, CIRCUNSCRIPCION, ORDEN, ELEGIDO (TRUE/FALSE)"

Prompt 2:

"You're a Computational Social Scientist. I want you to search for every candidate list of all autonomical elections in Spain from 2015 to 2026 of the following parties: PP and VOX. You have to get into the web and recopilate the following variables for each election:

election: election (write it in like this: "autonomous_2015_05", where 2015 would be the year and 05 the month where the election took place) year: year of the election chamber: the chamber (such as Parlament of Andalucia, Parlament of Catalonia...) province: the province that was being represented by the candidate list_position: the position of the candidate within that list candidate: the candidate's full name is_titular: whether he was the titular of that seat or he was a substitution candidate (TRUE/FALSE), is_elected: whether he was elected (TRUE/FALSE) abbreviation: the abbreviation of the party's name

the final file should be a csv file."

```{r READ ALL AUTONOMICAL CANDIDATES}
# =============================================================================
# PP and VOX candidates in the regional (autonomic) elections
#
# These do NOT come from infoelectoral: the Ministry of the Interior does not
# publish regional elections ("son competencia de cada Comunidad Autonoma"), so
# both files were built from the official proclamation bulletins collected by the
# Junta Electoral Central. That is why they are read from disk instead of a URL.
#
#   2000-2012 : PP only          (176 constituency-lists, from the JEC bulletins)
#   2015-2026 : PP and VOX
# =============================================================================

autonomical_pre2015 <- read_csv(
  "autonomicas_scrape/out_pre2015/pp_autonomicas_2000_2012.csv",
  show_col_types = FALSE
) |>
  transmute(
    election,
    year          = as.integer(year),
    chamber,
    province,
    list_position = as.integer(list_position),
    candidate     = str_squish(candidate),
    sex,
    is_titular,
    is_elected,                      # not resolved for this period: all NA
    abbreviation,
    level         = "Autonomical",
    source        = source_doc,
    list_completeness
  )

autonomical_2015_on <- read_csv(
  "candidatos_pp_vox_autonomicas_2015_2026.csv",
  show_col_types = FALSE
) |>
  transmute(
    election,
    year          = as.integer(year),
    chamber,
    province,
    list_position = as.integer(list_position),
    candidate     = str_squish(candidate),
    sex,
    is_titular,
    is_elected,
    abbreviation,
    level         = "Autonomical",
    source        = source_doc,
    list_completeness = "complete"   # every list here matches its legal length
  )

autonomical_candidates <- bind_rows(autonomical_pre2015, autonomical_2015_on) |>
  arrange(year, chamber, province, desc(is_titular), list_position)
```

```{r ADD SEX}
# library(readxl)
# hombres <- read_excel("nombres_por_edad_media.xlsx", sheet = "Hombres", skip = 6) |>
#   rename(name = 2, freq = 3) |>
#   transmute(name = as.character(name), freq = as.numeric(freq), sex_ine = "Male")
# 
# mujeres <- read_excel("nombres_por_edad_media.xlsx", sheet = "Mujeres", skip = 6) |>
#   rename(name = 2, freq = 3) |>
#   transmute(name = as.character(name), freq = as.numeric(freq), sex_ine = "Female")
# 
# ine_names <- bind_rows(hombres, mujeres) |>
#   mutate(name = name |>
#            str_to_upper() |>
#            stringi::stri_trans_general("Latin-ASCII") |>
#            str_remove_all("[^A-Z ]") |>
#            str_squish()) |>
#   filter(!is.na(name), name != "") |>
#   group_by(name) |>
#   slice_max(freq, n = 1, with_ties = FALSE) |>
#   ungroup() |>
#   select(name, sex_ine)
# 
# autonomical_candidates <- autonomical_candidates |>
#   mutate(
#     clean = candidate |>
#       str_to_upper() |>
#       stringi::stri_trans_general("Latin-ASCII") |>
#       str_remove_all("[^A-Z ]") |>
#       str_squish(),
#     tok1 = word(clean, 1),
#     tok2 = word(clean, 1, 2),
#     tok3 = word(clean, 1, 3)
#   ) |>
#   left_join(rename(ine_names, s1 = sex_ine), by = c("tok1" = "name")) |>
#   left_join(rename(ine_names, s2 = sex_ine), by = c("tok2" = "name")) |>
#   left_join(rename(ine_names, s3 = sex_ine), by = c("tok3" = "name")) |>
#   mutate(sex = coalesce(s3, s2, s1)) |>
#   select(-clean, -tok1, -tok2, -tok3, -s1, -s2, -s3)

```

Added 103 changes manually in sex in excel.



```{r READ ALL VOX MUNICIPAL CANDIDATES}
municipal_candidates <- readxl::read_xlsx("CANDIDATOS_MUNICIPALES.xlsx") |>
  transmute(
    election      = ELECCION,
    year          = as.integer(str_extract(ELECCION, "\\d{4}")),
    chamber       = NA_character_,
    ccaa          = CCAA,
    province      = CIRCUNSCRIPCION,
    list_position = as.integer(ORDEN),
    candidate     = CANDIDATO,
    is_elected    = ELEGIDO,
    source        = FUENTE,
    level         = "Municipal"
  )

all_candidates <- all_candidates |>   # `level` is already set upstream
  bind_rows(municipal_candidates)
```

```{r PP COUNCILERS}
# =============================================================================
# PP councillors and mayors from 2000 on
#
# Councillors come from the Ministry of the Interior microdata (same .DAT format
# already used above for Congress/Senate/European), mayors from the Ministry of
# Territorial Policy (SIL), because the mayor is chosen by the council after the
# election and therefore is not in the electoral files.
# =============================================================================

PORTAL <- "https://infoelectoral.interior.gob.es/estaticos/docxl/apliextr"

dir.create("datos_pp_locales/infoelectoral", recursive = TRUE, showWarnings = FALSE)
dir.create("datos_pp_locales/sil",           recursive = TRUE, showWarnings = FALSE)

elections_local <- tribble(
  ~process, ~year,  ~month, ~chamber,
  "04",     "2003", "05",   "Ayuntamiento",
  "04",     "2007", "05",   "Ayuntamiento",
  "04",     "2011", "05",   "Ayuntamiento",
  "04",     "2015", "05",   "Ayuntamiento",
  "04",     "2019", "05",   "Ayuntamiento",
  "04",     "2023", "05",   "Ayuntamiento"
)

# File 04 in MUNICIPAL elections is not the same layout as in general elections:
# it carries the municipality, and 2003 stores the whole name in a single field
# and has no sex column.
cols_04_local <- fwf_cols(
  province_code  = c( 10,  11),
  muni_code      = c( 13,  15),   # 12 is a control digit
  party_code     = c( 16,  21),
  list_position  = c( 22,  24),
  candidate_type = c( 25,  25),   # T = titular, S = substitute
  first_name     = c( 26,  50),
  surname_1      = c( 51,  75),
  surname_2      = c( 76, 100),
  sex            = c(101, 101),   # M / F
  elected        = c(120, 120)    # S = elected
)

cols_04_local_2003 <- fwf_cols(
  province_code  = c( 10,  11),
  muni_code      = c( 13,  15),
  party_code     = c( 16,  21),
  list_position  = c( 22,  24),
  candidate_type = c( 25,  25),
  full_name_cand = c( 26, 100),   # name and both surnames in one field
  elected        = c(120, 120)
)

# ------------------------------------------------------- download and read
local_candidates <- list()

for (i in seq_len(nrow(elections_local))) {

  process <- elections_local$process[i]
  year    <- elections_local$year[i]
  month   <- elections_local$month[i]
  chamber <- elections_local$chamber[i]

  folder   <- str_c(process, year, month, "_TOTA")
  zip_file <- str_c("datos_pp_locales/infoelectoral/", folder, ".zip")
  unzip_to <- tempfile()

  if (!file.exists(zip_file)) {
    message("Downloading ", chamber, " ", year, "-", month, " ...")
    download.file(str_c(PORTAL, "/", folder, ".zip"), zip_file, mode = "wb", quiet = TRUE)
  }
  unzip(zip_file, exdir = unzip_to)

  path_03 <- list.files(unzip_to, "^03.*[.]DAT$", full.names = TRUE,
                        recursive = TRUE, ignore.case = TRUE)[1]
  path_04 <- list.files(unzip_to, "^04.*[.]DAT$", full.names = TRUE,
                        recursive = TRUE, ignore.case = TRUE)[1]

  parties <- read_fwf(path_03, cols_03, locale = latin1,
                      col_types = cols(.default = "c")) |>
    mutate(across(everything(), str_trim)) |>
    mutate(
      abbr_ascii = abbreviation |> str_to_upper() |>
        stringi::stri_trans_general("Latin-ASCII"),
      name_ascii = full_name |> str_to_upper() |>
        stringi::stri_trans_general("Latin-ASCII")
    ) |>
    filter(
      str_detect(abbr_ascii, "^(PP|P\\.P\\.|PPC|PPDEG|PPCV|PPN|PPG|PP-|PP\\+)") |
        str_detect(name_ascii, "PARTIDO POPULAR|PARTIT POPULAR"),
      # formations that merely contain "popular" but are not the PP
      !str_detect(name_ascii,
                  str_c("UNIDAD POPULAR|FRENTE POPULAR|UNION POPULAR|",
                        "CANDIDATURA POPULAR|ALTERNATIVA POPULAR|ASAMBLEA POPULAR|",
                        "AGRUPACION POPULAR|IZQUIERDA|SOCIALISTA|NACIONALISTA|",
                        "COMUNISTA|ANDALUCISTA"))
    ) |>
    select(party_code, abbreviation, full_name)

  # 2003 has no sex column: decide the layout by probing column 101
  probe      <- read_lines(path_04, n_max = 200, locale = latin1)
  has_sex    <- any(str_sub(probe, 101, 101) %in% c("M", "F"))
  cols_local <- if (has_sex) cols_04_local else cols_04_local_2003

  local_candidates[[i]] <- read_fwf(path_04, cols_local, locale = latin1,
                                    col_types = cols(.default = "c")) |>
    mutate(across(everything(), str_trim)) |>
    inner_join(parties, by = "party_code") |>
    mutate(
      chamber   = chamber,
      year      = year,
      month     = month,
      candidate = if (has_sex)
        str_squish(str_c(first_name, " ", surname_1, " ", surname_2))
      else str_squish(full_name_cand),
      sex       = if (has_sex) sex else NA_character_
    )
}

municipal_candidates_pp <- bind_rows(local_candidates) |>
  left_join(provinces, by = "province_code") |>
  transmute(
    election      = str_c("municipal_", year, "_", month),
    year          = as.integer(year),
    chamber,
    province,
    ine_code      = str_c(province_code, muni_code),
    list_position = as.integer(list_position),
    candidate,
    sex           = recode(sex, "M" = "Male", "F" = "Female",
                           .default = NA_character_),
    is_titular    = candidate_type == "T",
    is_elected    = elected == "S",
    abbreviation,
    level         = "Municipal",
    source        = "infoelectoral.interior.gob.es"
  ) |>
  arrange(year, province, ine_code, list_position)

# ------------------------------------------------------------- mayors (SIL)
# The SIL download endpoints always serve the CURRENT term, so earlier terms can
# only be read from the files already saved locally.
mayors_files <- tribble(
  ~path,                                          ~term,       ~header_rows,
  "Alcaldes_Mandato_2015_2019.xlsx",              "2015-2019", 7,
  "Alcaldes_Mandato_2019_2023.xlsx",              "2019-2023", 7,
  "datos_pp_locales/sil/sil_alcaldes_2023.xlsx",  "2023-2027", 5
)

if (!file.exists("datos_pp_locales/sil/sil_alcaldes_2023.xlsx"))
  download.file("https://concejales.redsara.es/consulta/getAlcaldesLegislatura",
                "datos_pp_locales/sil/sil_alcaldes_2023.xlsx", mode = "wb", quiet = TRUE)

mayors_list <- list()

for (i in seq_len(nrow(mayors_files))) {

  raw <- readxl::read_excel(mayors_files$path[i], skip = mayors_files$header_rows[i],
                            col_types = "text")

  # the two layouts differ: the older files have no "Cargo" column and put the
  # full name in a single one
  if ("Cargo" %in% names(raw)) {
    mayors_list[[i]] <- raw |>
      filter(!is.na(`Código INE`)) |>
      transmute(term      = mayors_files$term[i],
                ine_code  = str_sub(`Código INE`, 1, 5),
                province  = Provincia,
                ccaa      = `Comunidad Autónoma`,
                candidate = str_squish(str_c(Nombre, " ", `1er Apellido`, " ",
                                             `2º Apellido`)),
                office    = Cargo,
                start     = `Fecha de Posesión`,
                party     = Partido)
  } else {
    mayors_list[[i]] <- raw |>
      select(ccaa = 4, province = 5, ine_code = 7, candidate = 8,
             party = 9, start = 10) |>
      filter(!is.na(ine_code)) |>
      transmute(term = mayors_files$term[i], ine_code = str_sub(ine_code, 1, 5),
                province, ccaa, candidate = str_squish(candidate),
                office = "Alcalde", start, party)
  }
}

municipal_mayors_pp <- bind_rows(mayors_list) |>
  filter(str_to_upper(str_trim(party)) %in% c("PP", "P.P.", "PARTIDO POPULAR")) |>
  mutate(level = "Municipal", abbreviation = "PP",
         source = "SIL - Ministerio de Politica Territorial") |>
  arrange(term, province, ine_code)

```


```{r CORRECTION MAZALY AGUILAR}
all_candidates <- all_candidates |>
  mutate(
    candidate = case_when(
      abbreviation == "VOX" & election == "general_2015_12" &
        province == "Madrid" & list_position == 6 ~
        "Mazaly Aguilar",

      abbreviation == "VOX" & election == "general_2015_12" &
        province == "Madrid" & list_position == 36 ~
        "Salvador Monedero (Independiente)",

      abbreviation == "VOX" & election == "general_2015_12" &
        province == "Madrid" & chamber == "Senate" & list_position == 3 ~
        "Doña Carmen Lomana (Independiente)",

      abbreviation == "VOX" & election == "general_2016_06" &
        province == "Madrid" & list_position == 4 ~
        "Mazaly Aguilar",

      abbreviation == "VOX" & election == "european_2014_05" &
        list_position == 13 ~
        "Mazaly Aguilar",

      abbreviation == "VOX" & election == "european_2019_05" &
        list_position == 2 ~
        "Mazaly Aguilar",

      abbreviation == "VOX" & election == "general_2023_07" &
        province == "Murcia" & chamber == "Senate" & list_position == 1 ~
        "José Francisco Garre Izquierdo",

      .default = candidate
    ),

    sex = case_when(
      abbreviation == "VOX" & election == "european_2014_05" &
        list_position == 12 ~ "Male",
      .default = sex
    )
  )
```

```{r ADD SEATS DISTRICT AND PARTY}
seats_district <- all_candidates |>
  filter(is_elected, is_titular = TRUE) |>
  count(election, chamber, province, name = "seats_district")

party_seats <- all_candidates |>
  filter(is_elected, is_titular) |>
  count(election, chamber, province, abbreviation, name = "seats_party")

all_candidates <- all_candidates |>
  left_join(seats_district,   by = c("election", "chamber", "province")) |>
  left_join(party_seats, by = c("election", "chamber", "province", "abbreviation")) |>
  mutate(seats_party = replace_na(seats_party, 0))
```


```{r NORMALIZE NAMES AND ELECTION DATE}
vox_candidates <- vox_candidates |>
  mutate(
    # crude person id: uppercase, no accents, no punctuation
    person_id = candidate |>
      str_to_upper() |>
      stringi::stri_trans_general("Latin-ASCII") |>
      str_remove_all("[^A-Z ]") |>
      str_squish()
    # Date?
  ) 

pp_candidates <- pp_candidates |>
  mutate(
    # crude person id: uppercase, no accents, no punctuation
    person_id = candidate |>
      str_to_upper() |>
      stringi::stri_trans_general("Latin-ASCII") |>
      str_remove_all("[^A-Z ]") |>
      str_squish()
    # Date?
  ) 

```

```{r}
inner_join(
  pp_candidates,
  vox_candidates,
  by = "person_id") |> filter(!is.na(person_id)) |> 
  distinct(person_id)

```


```{r NEWCOMER VARIABLE}
vox_candidates <- vox_candidates |> 
   group_by(person_id) |>
  mutate(
    appearance    = dense_rank(election_date),   # not row_number: someone can
    is_newcomer   = appearance == 1,             # run for Congress + Senate
    prev_position = lag(list_position, order_by = election_date),
    prev_elected  = lag(is_elected,    order_by = election_date)
  ) |>
  ungroup()
```


```{r} 

```






```{r ADD WHO WAS IN PP OR OTHER PARTIES BEFORE (NOT NEWCOMER)}

```

```{r}
library(readxl)

# --- 1. leer los dos ficheros de resultados provinciales ------------------
results <- list()

for (f in c("PROV_02_201606_1.xlsx", "PROV_02_201904_1.xlsx")) {

  raw <- read_excel(f, sheet = "Circunscripciones", col_names = FALSE)

  type_i  <- which(rowSums(raw == "Votos", na.rm = TRUE) > 0)[1]
  party_i <- type_i - 1

  party_row <- str_squish(as.character(unlist(raw[party_i, ])))
  type_row  <- str_squish(as.character(unlist(raw[type_i, ])))
  body      <- raw[(type_i + 1):nrow(raw), ]

  votes_cols <- which(type_row == "Votos"     & !is.na(party_row))
  seats_cols <- which(type_row == "Diputados" & !is.na(party_row))

  votes_long <- bind_cols(
      tibble(province = str_squish(as.character(body[[3]]))),
      set_names(body[votes_cols], party_row[votes_cols])
    ) |>
    pivot_longer(-province, names_to = "abbreviation", values_to = "votes")

  seats_long <- bind_cols(
      tibble(province = str_squish(as.character(body[[3]]))),
      set_names(body[seats_cols], party_row[seats_cols])
    ) |>
    pivot_longer(-province, names_to = "abbreviation", values_to = "seats")

  results[[f]] <- left_join(votes_long, seats_long,
                            by = c("province", "abbreviation")) |>
    mutate(
      votes    = as.numeric(votes),
      seats    = as.numeric(seats),
      province = province |>
        stringi::stri_trans_general("Latin-ASCII") |>
        str_squish(),
      base_election = if_else(str_detect(f, "201606"),
                              "general_2016_06", "general_2019_04")
    ) |>
    filter(votes > 0)
}

results <- bind_rows(results)

# --- 2. magnitud de cada circunscripcion ---------------------------------
magnitude <- results |>
  filter(base_election == "general_2019_04") |>
  group_by(province) |>
  summarise(magnitude = sum(seats), .groups = "drop")

# --- 3. escenarios de voto nacional de VOX -------------------------------
# marzo 2019: la referencia es 2016 (VOX partia de casi cero)
# octubre 2019: la referencia es abril 2019 (VOX ya tenia 10,3% real)
scenarios <- bind_rows(
  tibble(moment = "march_2019",   base_election = "general_2016_06",
         vox_nat = c(0.08, 0.10, 0.12, 0.14)),
  tibble(moment = "october_2019", base_election = "general_2019_04",
         vox_nat = c(0.08, 0.10, 0.12, 0.14))
)

# --- 4. proyeccion: swing uniforme, renormalizando el resto --------------
others <- results |>
  filter(abbreviation != "VOX") |>
  group_by(base_election, province) |>
  mutate(share_others = votes / sum(votes)) |>
  ungroup() |>
  select(base_election, province, abbreviation, share_others)

expected <- bind_rows(
  left_join(others, scenarios, by = "base_election",
            relationship = "many-to-many") |>
    mutate(share_exp = share_others * (1 - vox_nat)),
  left_join(distinct(others, base_election, province), scenarios,
            by = "base_election", relationship = "many-to-many") |>
    mutate(abbreviation = "VOX", share_exp = vox_nat)
)

# --- 5. D'Hondt -----------------------------------------------------------
seats_exp <- expected |>
  left_join(magnitude, by = "province") |>
  filter(share_exp >= 0.03) |>
  mutate(divisor = map(magnitude, seq_len)) |>
  unnest(divisor) |>
  mutate(quotient = share_exp / divisor) |>
  group_by(moment, vox_nat, province) |>
  arrange(desc(quotient), .by_group = TRUE) |>
  filter(row_number() <= magnitude) |>
  count(moment, vox_nat, province, abbreviation, name = "seats") |>
  ungroup()

vox_expected <- seats_exp |>
  filter(abbreviation == "VOX") |>
  select(-abbreviation) |>
  pivot_wider(names_from = vox_nat, values_from = seats, values_fill = 0) |>
  arrange(moment, province)

vox_expected

# --- 6. comprobaciones ----------------------------------------------------
count(results, base_election)
n_distinct(results$province)          # deben ser 52
sum(magnitude$magnitude)              # debe dar 350
anti_join(magnitude, distinct(all_candidates, province), by = "province")
```

```{r}
library(readxl)

# --- 1. leer los dos ficheros de resultados provinciales ------------------
results <- list()

for (f in c("PROV_02_201606_1.xlsx", "PROV_02_201904_1.xlsx")) {

  m <- as.matrix(read_excel(f, sheet = "Circunscripciones",
                            col_names = FALSE, .name_repair = "minimal"))
  mode(m) <- "character"

  hits    <- which(m == "Votos", arr.ind = TRUE)
  type_i  <- min(hits[, "row"])
  party_i <- type_i - 1

  party_row <- str_squish(m[party_i, ])
  type_row  <- str_squish(m[type_i, ])

  votes_cols <- which(type_row == "Votos" & !is.na(party_row))
  seats_cols <- votes_cols + 1

  stopifnot(length(votes_cols) > 0,
            all(type_row[seats_cols] == "Diputados"))

  body_rows <- (type_i + 1):nrow(m)

  province <- m[body_rows, 3] |>
    stringi::stri_trans_general("Latin-ASCII") |>
    str_squish() |>
    str_replace("^Araba.*", "Araba")

  votes_m <- m[body_rows, votes_cols, drop = FALSE]
  seats_m <- m[body_rows, seats_cols, drop = FALSE]
  colnames(votes_m) <- party_row[votes_cols]
  colnames(seats_m) <- party_row[votes_cols]

  votes_long <- as_tibble(votes_m, .name_repair = "unique") |>
    mutate(province = province) |>
    pivot_longer(-province, names_to = "abbreviation", values_to = "votes")

  seats_long <- as_tibble(seats_m, .name_repair = "unique") |>
    mutate(province = province) |>
    pivot_longer(-province, names_to = "abbreviation", values_to = "seats")

  results[[f]] <- left_join(votes_long, seats_long,
                            by = c("province", "abbreviation")) |>
    mutate(
      votes = as.numeric(votes),
      seats = as.numeric(seats),
      base_election = if_else(str_detect(f, "201606"),
                              "general_2016_06", "general_2019_04")
    ) |>
    filter(!is.na(province), province != "", !is.na(votes), votes > 0) |>
      filter(!is.na(province), province != "", province != "Total",
           !is.na(votes), votes > 0)
}

results <- bind_rows(results)

stopifnot(n_distinct(results$province) == 52)

# --- 2. magnitud de cada circunscripcion ---------------------------------
magnitude <- results |>
  filter(base_election == "general_2019_04") |>
  group_by(province) |>
  summarise(magnitude = sum(seats, na.rm = TRUE), .groups = "drop")

stopifnot(sum(magnitude$magnitude) == 350)

# --- 3. escenarios de voto nacional de VOX -------------------------------
scenarios <- bind_rows(
  tibble(moment = "march_2019",   base_election = "general_2016_06",
         vox_nat = c(0.08, 0.10, 0.12, 0.14)),
  tibble(moment = "october_2019", base_election = "general_2019_04",
         vox_nat = c(0.08, 0.10, 0.12, 0.14))
)

# --- 4. proyeccion: swing uniforme, renormalizando el resto --------------
others <- results |>
  filter(abbreviation != "VOX") |>
  group_by(base_election, province) |>
  mutate(share_others = votes / sum(votes)) |>
  ungroup() |>
  select(base_election, province, abbreviation, share_others)

expected <- bind_rows(
  left_join(others, scenarios, by = "base_election",
            relationship = "many-to-many") |>
    mutate(share_exp = share_others * (1 - vox_nat)),
  left_join(distinct(others, base_election, province), scenarios,
            by = "base_election", relationship = "many-to-many") |>
    mutate(abbreviation = "VOX", share_exp = vox_nat)
) |>
  left_join(magnitude, by = "province")

stopifnot(!any(is.na(expected$magnitude)))

# --- 5. D'Hondt -----------------------------------------------------------
seats_exp <- expected |>
  filter(share_exp >= 0.03) |>
  mutate(divisor = map(magnitude, seq_len)) |>
  unnest(divisor) |>
  mutate(quotient = share_exp / divisor) |>
  group_by(moment, vox_nat, province) |>
  arrange(desc(quotient), .by_group = TRUE) |>
  filter(row_number() <= magnitude) |>
  count(moment, vox_nat, province, abbreviation, name = "seats") |>
  ungroup()

vox_expected <- seats_exp |>
  filter(abbreviation == "VOX") |>
  select(-abbreviation) |>
  pivot_wider(names_from = vox_nat, values_from = seats, values_fill = 0) |>
  arrange(moment, province)

vox_expected |> 
  arrange(province)
```


```{r}
results |>
  filter(base_election == "general_2019_04") |>
  distinct(province) |>
  anti_join(filter(results, base_election == "general_2016_06") |> distinct(province),
            by = "province")
```



```{r}
# --- expectativa media de escaños por provincia y momento -----------------
expectation <- seats_exp |>
  filter(abbreviation == "VOX") |>
  group_by(moment, province) |>
  summarise(exp_seats = mean(seats), .groups = "drop") |>
  bind_rows(                                   # provincias con 0 en todos los escenarios
    crossing(moment = c("march_2019", "october_2019"),
             province = magnitude$province) |>
      anti_join(filter(seats_exp, abbreviation == "VOX"),
                by = c("moment", "province")) |>
      mutate(exp_seats = 0)
  ) |>
  pivot_wider(names_from = moment, values_from = exp_seats,
              names_prefix = "exp_") |>
  mutate(exp_gain = exp_october_2019 - exp_march_2019)

# --- listas de abril y noviembre, solo Congreso y titulares ---------------
lists_2019 <- vox_candidates |>
  filter(chamber == "Congress", is_titular,
         election %in% c("general_2019_04", "general_2019_11")) |>
  mutate(province = str_replace(province, "^Araba.*", "Araba"))

# --- rotación por provincia ----------------------------------------------
churn <- lists_2019 |>
  group_by(province, person_id) |>
  summarise(in_april = any(election == "general_2019_04"),
            in_nov   = any(election == "general_2019_11"),
            .groups  = "drop_last") |>
  summarise(
    n_april   = sum(in_april),
    n_nov     = sum(in_nov),
    stayed    = sum(in_april & in_nov),
    dropped   = sum(in_april & !in_nov),
    added     = sum(!in_april & in_nov),
    churn_rate = dropped / n_april,
    .groups   = "drop"
  ) |>
  left_join(expectation, by = "province")

churn |> arrange(desc(exp_october_2019))

# --- total de nombres que cambiaron --------------------------------------
churn |> summarise(across(c(n_april, stayed, dropped, added), sum))

# --- rotación solo en los puestos
```


```{r}
# lo que sabían en octubre: el resultado REAL de abril, no una proyección
seats_april <- vox_candidates |>
  filter(election == "general_2019_04", chamber == "Congress",
         is_titular, is_elected) |>
  mutate(province = str_replace(province, "^Araba.*", "Araba")) |>
  count(province, name = "seats_april")

bands <- expectation |>
  left_join(seats_april, by = "province") |>
  mutate(seats_april = replace_na(seats_april, 0))

# ¿quién es nuevo en la lista de noviembre?
november <- vox_candidates |>
  filter(election == "general_2019_11", chamber == "Congress", is_titular) |>
  mutate(province = str_replace(province, "^Araba.*", "Araba"))

april_ids <- vox_candidates |>
  filter(election == "general_2019_04", chamber == "Congress", is_titular) |>
  mutate(province = str_replace(province, "^Araba.*", "Araba")) |>
  distinct(province, person_id)

test <- november |>
  left_join(bands, by = "province") |>
  left_join(mutate(april_ids, was_in_april = TRUE),
            by = c("province", "person_id")) |>
  mutate(
    is_new = is.na(was_in_april),
    band = case_when(
      list_position <= seats_april       ~ "1_ya ganable en abril",
      list_position <= exp_october_2019  ~ "2_nuevo ganable",
      TRUE                               ~ "3_no ganable"
    )
  )

test |>
  group_by(band) |>
  summarise(n = n(), share_new = mean(is_new))

# solo en provincias donde la expectativa creció
test |>
  filter(exp_october_2019 > seats_april) |>
  group_by(band) |>
  summarise(n = n(), share_new = mean(is_new))

# posición exacta donde entran los nuevos, respecto al corte de abril
test |>
  mutate(dist_cutoff = list_position - seats_april) |>
  filter(dist_cutoff >= -3, dist_cutoff <= 6) |>
  group_by(dist_cutoff) |>
  summarise(n = n(), share_new = mean(is_new))
```


```{r}
test |>
  filter(!is_new, exp_october_2019 > seats_april) |>
  left_join(
    vox_candidates |>
      filter(election == "general_2019_04", chamber == "Congress", is_titular) |>
      mutate(province = str_replace(province, "^Araba.*", "Araba")) |>
      select(province, person_id, pos_april = list_position),
    by = c("province", "person_id")
  ) |>
  mutate(move = pos_april - list_position) |>
  group_by(band) |>
  summarise(n = n(), mean_move = mean(move, na.rm = TRUE),
            promoted = mean(move > 0, na.rm = TRUE))
```

```{r}
test |>
  group_by(province) |>
  summarise(identica = all(!is_new), n = n()) |>
  count(identica)
```

```{r}
test |>
  group_by(province) |>
  summarise(first_new = min(list_position[is_new]),
            seats_april = first(seats_april),
            .groups = "drop") |>
  count(first_new > seats_april)
```
```{r}
test |>
  group_by(province) |>
  summarise(first_new = min(list_position[is_new]),
            seats_april = first(seats_april), .groups = "drop") |>
  filter(is.na(first_new) | is.na(seats_april) | first_new <= seats_april) |>
  left_join(select(test, province, list_position, candidate, is_new),
            by = "province") |>
  filter(is_new) |>
  arrange(province, list_position) |> 
  view()
  
```
 
 
```{r}

```
 
