---
title: "TFM_CODE"
author: "Noah Emanuel Fürup"
format: html
editor: visual
---

##Libraries

```{r LIBRARIES}
library(tidyverse)
library(readxl)
library(sf)
library(rnaturalearth)
library(countrycode)
```

##Charts of the Theoretical Framework

ARREGLAR FRANCIA, NORUEGA que faltan

```{r ELECTORAL SYSTEM MAP}

electoral_system_database <-
  read_excel("idea_export_electoral_system_design_database.xlsx") |> 
  group_by(ISO3) |> 
   slice_max(Year) |> 
  ungroup() |> 
  select(ISO3, Country, Year,
         family = `Electoral system family`,
         system = `Electoral system for national legislature`)
 
europe <- 
  ne_countries(scale = "medium", continent = "Europe", returnclass = "sf") |>
  left_join(electoral_system_database, by = c("iso_a3" = "ISO3"))

ggplot(europe) +
  geom_sf(aes(fill = family), colour = "white", linewidth = 0.2) +
  coord_sf(xlim = c(-25, 45), ylim = c(34, 72)) +
  scale_fill_brewer(palette = "Set2", na.value = "grey90", name = "Electoral system family") +
  labs(title = "Electoral systems in Europe",
       caption = "Source: IDEA Electoral System Design Database") +
  theme_void()

```

```{r share of each electoral family}
europe |>
  st_drop_geometry() |>
  count(family) |>
  mutate(pct = round(100 * n / sum(n), 1)) |>
  arrange(desc(n))
```

```{r TABLE OF FAR RIGHT PARTIES BEFORE VOX}

```

```{r TABLE OF VOX'S FOUNDERS}

```

# Data Extraction

## General elections

Extraction of general elections candidates: european and general elections via github.com/JaimeObregon/infoelectoral

```{r SPANISH LETERS}
latin1 <- locale(encoding = "ISO-8859-1")
```

```{r GITHUB AND INTERIOR MINISTERY PORTAL}
GITHUB <- "https://raw.githubusercontent.com/JaimeObregon/infoelectoral/main/files"
PORTAL <- "https://infoelectoral.interior.gob.es/estaticos/docxl/apliextr"
```

The github has the files in .DAT, a file type that doesn't separate in ";" or "\|". In order to separate each column, we need to map where the separation between each variable occurs. I asked Claude to give me the exact character positions where the separation is. I used fwf_positions of ReadR to write down the column names, given the positions where the variables lie in the .DAT file.

EXPLICAR COMO FUNCIONAN LO DE 02197706_TOTA y decir que el github copia la estructura del ministerio

```{r FIXED POSITIONS}
# positions of each variable in the .DAT files
cols_parties <- fwf_cols(
  party_code   = c(9, 14),
  abbreviation = c(15, 64)
)

cols_candidates <- fwf_cols(
  province_code  = c(10, 11),
  party_code     = c(16, 21),
  list_position  = c(22, 24),
  candidate_type = c(25, 25),
  candidate      = c(26, 100),
  sex            = c(101, 101),
  elected        = c(120, 120)
)
```

```{r GENERAL AND EUROPEAN ELECTIONS TIBBLE (2014-2024)}
elections <- tribble(
  ~process, ~year,  ~month, ~chamber,              ~mirror_folder,
  "02",     "1977", "06",   "Congress",            "congreso",
  "02",     "1979", "03",   "Congress",            "congreso",
  "02",     "1982", "10",   "Congress",            "congreso",
  "02",     "1986", "06",   "Congress",            "congreso",
  "02",     "1989", "10",   "Congress",            "congreso",
  "02",     "1993", "06",   "Congress",            "congreso",
  "02",     "1996", "03",   "Congress",            "congreso",
  "02",     "2000", "03",   "Congress",            "congreso",
  "02",     "2004", "03",   "Congress",            "congreso",
  "02",     "2008", "03",   "Congress",            "congreso",
  "02",     "2011", "11",   "Congress",            "congreso",
  "02",     "2015", "12",   "Congress",            "congreso",
  "02",     "2016", "06",   "Congress",            "congreso",
  "02",     "2019", "04",   "Congress",            "congreso",
  "02",     "2019", "11",   "Congress",            "congreso",
  "02",     "2023", "07",   "Congress",            "congreso",
  "03",     "1977", "06",   "Senate",              "senado",
  "03",     "1979", "03",   "Senate",              "senado",
  "03",     "1982", "10",   "Senate",              "senado",
  "03",     "1986", "06",   "Senate",              "senado",
  "03",     "1989", "10",   "Senate",              "senado",
  "03",     "1993", "06",   "Senate",              "senado",
  "03",     "1996", "03",   "Senate",              "senado",
  "03",     "2000", "03",   "Senate",              "senado",
  "03",     "2004", "03",   "Senate",              "senado",
  "03",     "2008", "03",   "Senate",              "senado",
  "03",     "2011", "11",   "Senate",              "senado",
  "03",     "2015", "12",   "Senate",              "senado",
  "03",     "2016", "06",   "Senate",              "senado",
  "03",     "2019", "04",   "Senate",              "senado",
  "03",     "2019", "11",   "Senate",              "senado",
  "03",     "2023", "07",   "Senate",              "senado",
  "07",     "2014", "05",   "European Parliament", "europeas",
  "07",     "2019", "05",   "European Parliament", "europeas",
  "07",     "2024", "06",   "European Parliament", "europeas"
)
```

CHEQUEAR SI SE PUEDE MEJORAR CODIGO DE DESCARGA DE DATOS GENERALS

types of processes: 02 = congress, 03 = Senate, 07 = European Parliament

```{r ADD SOURCE AND PATH OF EACH ELECTION}
elections <-
  elections |>
  mutate(
    folder    = str_c(process, year, month, "_TOTA"),
    code      = str_c(process, str_sub(year, 3, 4), month),
    on_github = !(year %in% c("2023", "2024") |
                    (chamber == "Congress" & year %in% c("1989", "1993"))),
    path      = if_else(on_github,
                        str_c(GITHUB, "/", mirror_folder, "/", folder, "/"),
                        str_c(tempdir(), "/", folder, "/"))
  )

```

```{r DOWNLOAD NON-GITHUB DATA TO TEMPORARY FOLDER}
for (folder in elections$folder[!elections$on_github]) {
  zip_file <- tempfile(fileext = ".zip")
  download.file(str_c(PORTAL, "/", folder, ".zip"), zip_file, mode = "wb",
                method = "libcurl", headers = c("User-Agent" = "Mozilla/5.0"))
  unzip(zip_file, exdir = str_c(tempdir(), "/", folder), junkpaths = TRUE)
}
```

```{r READ CANDIDATES FROM PATH}

# candidates of every election (file 04)
candidates <- elections |>
  rowwise() |>
  reframe(
    folder, chamber, year, month,
    read_fwf(str_c(path, "04", code, ".DAT"), cols_candidates,
             locale = latin1, col_types = cols(.default = "c"))
  )

candidates
```

```{r JOIN PARTIES TO CANDIDATES}
# parties of every election (file 03)
parties <- elections |>
  rowwise() |>
  reframe(
    folder,
    read_fwf(str_c(path, "03", code, ".DAT"), cols_parties,
             locale = latin1, col_types = cols(.default = "c"))
  )

all_candidates <-
  candidates |> 
  left_join(parties,   by = c("folder", "party_code"))

all_candidates
```

```{r JOIN PROVINCE TO CANDIDATES}
provinces <- tribble(
  ~province_code, ~province,
  "01", "Araba/Alava",    "02", "Albacete",      "03", "Alicante",
  "04", "Almeria",        "05", "Avila",         "06", "Badajoz",
  "07", "Illes Balears",  "08", "Barcelona",     "09", "Burgos",
  "10", "Caceres",        "11", "Cadiz",         "12", "Castellon",
  "13", "Ciudad Real",    "14", "Cordoba",       "15", "A Coruna",
  "16", "Cuenca",         "17", "Girona",        "18", "Granada",
  "19", "Guadalajara",    "20", "Gipuzkoa",      "21", "Huelva",
  "22", "Huesca",         "23", "Jaen",          "24", "Leon",
  "25", "Lleida",         "26", "La Rioja",      "27", "Lugo",
  "28", "Madrid",         "29", "Malaga",        "30", "Murcia",
  "31", "Navarra",        "32", "Ourense",       "33", "Asturias",
  "34", "Palencia",       "35", "Las Palmas",    "36", "Pontevedra",
  "37", "Salamanca",      "38", "S.C. Tenerife", "39", "Cantabria",
  "40", "Segovia",        "41", "Sevilla",       "42", "Soria",
  "43", "Tarragona",      "44", "Teruel",        "45", "Toledo",
  "46", "Valencia",       "47", "Valladolid",    "48", "Bizkaia",
  "49", "Zamora",         "50", "Zaragoza",      "51", "Ceuta",
  "52", "Melilla",        "99", "Single national constituency"
)

 all_candidates <-
  all_candidates |>
  left_join(provinces, by = "province_code")
 
 all_candidates 
```

```{r MUTATE AND SELECT VARIABLES OF ALL_CANDIDATES}

all_candidates <-
  all_candidates |> 
  mutate(
    election      = str_c(if_else(chamber == "European Parliament", "european", "general"),
                          "_", year, "_", month),
    year          = as.integer(year),
    chamber,
    list_position = as.integer(list_position),
    candidate     = candidate |>
           str_to_upper() |>
           stringi::stri_trans_general("Latin-ASCII") |>
           str_remove_all("[^A-Z ]") |>
           str_squish(),
    sex           = case_match(sex, "M" ~ "Male", "F" ~ "Female"),
    is_titular    = candidate_type == "T",
    is_elected    = elected == "S",
    abbreviation,
    level         = "National",
    .keep = "none"
  )
all_candidates
```

```{r TAKE DON/ DOÑA OUT}
all_candidates <- 
  all_candidates |>
  mutate(candidate = str_remove(candidate,
                                regex("^(don|do[ñn]a|d[.ª])\\s+")))
```

```{r FILTER PP, CS AND VOX}
pp_candidates <- all_candidates |> 
  filter(str_detect(abbreviation, "PP"))

vox_candidates <- all_candidates |> 
  filter(str_detect(abbreviation, "VOX"))

cs_candidates <- all_candidates |> 
  filter(str_detect(abbreviation, "C's"))
```

## Autonomical Elections

Claude Code For Autonomical elections: two attempts to confirm it is reproducible.

Prompt 1:

"Quiero que busques todas las listas de todas las elecciones autonómicas que han habido en España desde 2015 hasta 2026 y que pongas en el archivo excel en el excel de CANDIDATOS_AUTONOMICAS. Tienes que ir a la web para recopilar las siguientes variables de cada eleccion: CANDIDATO, CCAA, ELECCION, CIRCUNSCRIPCION, ORDEN, ELEGIDO (TRUE/FALSE)"

Prompt 2:

"You're a Computational Social Scientist. I want you to search for every candidate list of all autonomical elections in Spain from 2015 to 2026 of the following parties: PP and VOX. You have to get into the web and recopilate the following variables for each election:

election: election (write it in like this: "autonomous_2015_05", where 2015 would be the year and 05 the month where the election took place) year: year of the election chamber: the chamber (such as Parlament of Andalucia, Parlament of Catalonia...) province: the province that was being represented by the candidate list_position: the position of the candidate within that list candidate: the candidate's full name is_titular: whether he was the titular of that seat or he was a substitution candidate (TRUE/FALSE), is_elected: whether he was elected (TRUE/FALSE) abbreviation: the abbreviation of the party's name

the final file should be a csv file."


```{r}

```


```{r READ ALL AUTONOMICAL CANDIDATES}
# =============================================================================
# PP and VOX candidates in the regional (autonomic) elections
#
# These do NOT come from infoelectoral: the Ministry of the Interior does not
# publish regional elections ("son competencia de cada Comunidad Autonoma"), so
# both files were built from the official proclamation bulletins collected by the
# Junta Electoral Central. That is why they are read from disk instead of a URL.
#
#   2000-2012 : PP only          (176 constituency-lists, from the JEC bulletins)
#   2015-2026 : PP and VOX
# =============================================================================

autonomical_pre2015 <- read_csv(
  "autonomicas_scrape/out_pre2015/pp_autonomicas_2000_2012.csv",
  show_col_types = FALSE
) |>
  transmute(
    election,
    year          = as.integer(year),
    chamber,
    province,
    list_position = as.integer(list_position),
    candidate     = str_squish(candidate),
    sex,
    is_titular,
    is_elected,                      # not resolved for this period: all NA
    abbreviation,
    level         = "Autonomical",
    source        = source_doc,
    list_completeness
  )

autonomical_2015_on <- read_csv(
  "candidatos_pp_vox_autonomicas_2015_2026.csv",
  show_col_types = FALSE
) |>
  transmute(
    election,
    year          = as.integer(year),
    chamber,
    province,
    list_position = as.integer(list_position),
    candidate     = str_squish(candidate),
    sex,
    is_titular,
    is_elected,
    abbreviation,
    level         = "Autonomical",
    source        = source_doc)

autonomical_candidates <- bind_rows(autonomical_pre2015, autonomical_2015_on) |>
  arrange(year, chamber, province, desc(is_titular), list_position)
```

```{r READ CIUDADANOS CANDIDATES AUTONOMICAL}
cs_candidates_autonomical <- read_csv("candidatos_cs_autonomicas_2015_2026.csv")
cs_candidates_autonomical
```

```{r VOX's POSITIONS IN GOVERNMENT}

```
